### 2019-02-25 docs/notebooks/results.Rmd
initial add of results.Rmd

---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.0'
      jupytext_version: 1.0.0
  kernelspec:
    display_name: Python 3
    language: python
    name: python3
---

# Attentional effects in convolutional neural nets without attentional mechanisms

## Introduction
Neural networks have in the past few years transitioned from being an academic area of artificial intellgience to being powerful tools for engineers working in areas such as image classification, speech-to-text, and automatic translation, to name but a few. In part because of this resurgence of research on neural networks, many in neuroscience and cognitive science have returned to using these models to understand the brain. Of course there are many questions about what aspects of the brain these models actually capture <cite data-cite="6027200/PAZZI897"></cite><cite data-cite="6027200/KEUFZ7BF"></cite>. The goal of the study summarized here is to investigate the extent to which one type of neural network architecture, the convolutional neural network, provides a good model of the visual system in the brain of humans and other primates.

### Connectionist models of cognition and deep neural networks
In the 1980s, artificial neural networks trained with backpropagation gave rise to [Connectionist](https://en.wikipedia.org/wiki/Connectionism) models of cognition. Similar neural network models lie at the heart of the recent rise of deep learning, in large part because of advances in methods for training algorithms, and advances in technology such as GPUs that allow for parallelized training. When benchmarked with scalar metrics like accuracy on computer vision tasks such as object recognition, these deep neural networks appear capable of human or super-human levels of skill. The first major gains in accuracy were seen on image classification tasks such as ImageNet <cite data-cite="6027200/WGZWMWX7"></cite>, and the deep neural networks that made these advances were built with what are known as convolutional layers. As the name implies, these layers perform a convolution-like operation by multiplying filters of learned weights with the input image, producing as outputs feature maps that can be further transformed by succeeding layers.

### Convolutional neural networks as models of the primate visual system
Many researchers have drawn parallels between the architecture of convolutional neural networks (CNNs) and the architecture of the visual system in the brain <cite data-cite="6027200/8ISULWGF"></cite>, in particular in humans and other primates where this system has been most thoroughly studied. Like CNNs, the visual system has a hierarchical structure, and is thought to function in part by performing transformations at each level of this hierarchy so that high-dimensional, low-level features are mapped into low-dimensional abstract representations. Several recent studies in the realm of cognitive neuroscience have measured representations in more directly tested the similarities between CNNs and the visual system, by comparing representations in both <cite data-cite="6027200/6D4LKJD9"></cite><cite data-cite="6027200/U77Y74SI"></cite>. Representations can be easily extracted from CNNs because they are "transparent box" models--i.e. we have access to the "activity" of the individual units from which these models are built when we "show" the CNNs a stimulus. In the same studies, brain activity was obtained with electrophysiological or functional imaging methods. Such studies have found in general that representations are similar. (The comparison depends on representational similarity methods <cite data-cite="6027200/IQ2Y895S"></cite><cite data-cite="6027200/EQTU2P3Z"></cite>, but the concern of this study is not the validity of those methods.)

### Does learning to perform some task with high accuracy constrain neural networks and brains to learn similar representations?
The authors conclude that one factor that may cause this representational similarity is that the goal-driven tasks provide a constraint on learning. In other words, these hierarchichal architectures learn similar representations because those representations provide a way to succesfully carry out behavior. Similar claims have been made elsewhere, e.g. that the brain and artificial networks both learn similar representations of location when performing navigation tasks <cite data-cite="6027200/LHVT6V6M"></cite><cite data-cite="6027200/SAZAIFBG"></cite>.

The impressive accuracy of deep convolutional neural networks performing image recognition tasks, taken in combination with the apparent correlations between their activity and activity in the primate visual system, raises two questions: (1) how accurately do CNNs perform other tasks that are though to rely on the visual system, and (2) if they perform a task accurately, is that enough evidence to conclude that they provide useful models of the visual system?

### Another type of constraint: controlled behavioral tasks in the lab
Importantly, the researchers in the studies just described constrain the "task"--passively viewing images--when measuring brain activity, with the goal of being able to directly compare brain activity to activity in neural networks. Specifically, the subjects are only shown the visual stimuli very briefly, typically about 100 ms for monkey electophysiology experiments and 300 ms for human neuroimaging experiments. There are a couple of reasons for this. The first is to allow for an apples-to-apples comparison between the brain and CNNs. Although the brain is highly recurrent, CNNs are purely **feedforward**. The researchers want to claim that they have only elicited  initial feedforward activity in the visual system by just briefly showing subjects the stimuli, without allowing feedback to influence activity, and by extension, the representation.
The second reason for the constraint of showing images only briefly is more historical. Classically, the visual system has been studied with what has been termed **passive vision** approach <cite data-cite="6027200/B49GIRUD"></cite>. This approach could be characterized as conceiving of the eye as unmoving, with a still image falling upon it that is then processed in parallel across the scene to extract information. These types of models have obviously been successful, and the incredible accuracy of CNNs might even be described as the culmination of this approach. The point here is not to debate about issues with passive vision models; most readers probably take it for granted that humans move their eyes, and anyone who has tried to walk through a crowd while texting on their phone will know that there are limitations on our ability to process the entire visual scene in parallel.

### How much can accuracy on a task tell us about cognition? Visual search tasks as a case study
The point is that the same constraint that was used to make apples-to-apples comparisons between CNNs and the brains, briefly showing stimuli, has also been used in visual search tasks in the laboratory <cite data-cite="6027200/AEJJ5749"></cite>, in part because of the passive vision approach with which the visual system has typically been studied. The fact that the same constraint was applied to visual search tasks lets us use the same logic as studies which conclude that goal-driven tasks constrain learning to produce useful representations. We can take advantage of this shared constraint to invert the logic and state a new hypothesis: if a CNN has learned the same representation as the brain, then its behavior when performing tasks should match our behavior. In other words, we are asking whether we can extend a claim about CNNs, that they provide good models of the visual system, by examining how they behave when performing a visual search task. There are several objections that could be made to doing so, including the question of whether the task we'll look at is really any different from image recognition. But let's pretend for a moment that the question is worth asking. Why is it worth doing so? To consider more generally how much we can learn about cognition by measuring the behavior of our model with a scalar metric applied to a highly constrained task.

What is that task? In the "classical" form of the visual search paradigm <cite data-cite="6027200/AEJJ5749"></cite>, experimenters present a stimulus to a subject made up of discrete items, i.e. objects. The subject is asked to report whether a target is present among distractors-- the experimenter sets the probability that the target will be present on any given trial. Experimenters measure the reaction time and/or accuracy of the subjects. Often, the results are presented as a plot of reaction time as a function of set size (the number of discrete items in stimulus). By fitting a line to the plot of reaction times as a function of set size and taking the slope of that line, experimenters produced estimates of processing speed, i.e. "seconds per item". Slopes are almost horizontal when subjects are shown visual search stimuli where targets differ from distractors in some clearly salient feature, e.g. a red target among green distractors, and slopes become much steeper when targets differ from distractors in some harder-to-distinguish conjunction of features, e.g., a digital 2 among digital 5s. Some examples of these stimuli are shown below, generated with https://github.com/NickleDave/searchstims.

<table><tr>
<td> <img src="../static/redvert_v_greenvert_set_size_4_target_present_55.png" alt="efficient" style="width: 250px;"/> </td>
<td> <img src="../static/two_v_five_set_size_6_target_present_78.png" alt="inefficient" style="width: 250px;"/> </td>
</tr>
<tr>
<td> <b>Feature</b> search $\longrightarrow$ almost horizontal slopes </td>
<td> <b>Spatial configuration</b> search $\longrightarrow$ steep slopes </td>
</tr>
</table>

Below is a figure explaining the basic visual search paradigm, reproduced from [Wolfe Horowitz 2017](http://search.bwh.harvard.edu/new/pubs/FiveFactors_Wolfe-Horowitz_2017.pdf)

<img src="../static/Fig-2-Wolfe-Horowitz-2017.png" alt="inefficient" style="width: 900px;"/>

This use of search slope as metric led to one of the most influential early models of visual search: feature integration theory <cite data-cite="6027200/KYF4NSZP"></cite>. Essentially, feature integration theory posits that the brain can search the visual field in parallel when a single feature distinguishes the target from distractors, but that a conjunction of multiple features requires serial processing. However, Wolfe <cite data-cite="6027200/LJEPEDWB"></cite>(1998) showed that there is no clear cut segregation of slopes into one type or the other, and that in fact the distributions of slope overlap so much it is not possible to distinguish them. He has proposed talking about visual search in terms of a spectrum of of speeds that depend in some way on features: when a salient feature like color is what discriminates the target from the distractor, this allows for **efficient** search, like in the visual search stimulus on the left. When a complex conjunction of features distinguishes the target from the distractor, as in the visual search stimulus on the right, this results in **inefficient** search.

### Accuracy as a measure of attentional effects
Although these classic visual search tasks were originally designed to answer questions about how features were processed across the visual field, they went on to became the basis for most modern theories of selective visual attention. As described above, the feature integration theory (FIT) held that some features could be processed in parallel, in a pre-attentional way, but a conjunction of features required directing attention to it, and that items consisting of such features were processed in a serial way. Given this theory about the algorithms underlying attention, reaction times become a proxy for attention; they are what we measure to see if our theories about attention are correct.  
The approach of measuring reaction times is not always used. In some cases, the stimulus is shown only briefly to the subject, and then after a variable length of time a masking stimulus such as white noise is shown. During the period between the stimulus with a set of items and the masking stimulus, the "*internal representation* of that stimulus can be searched \[by the subject\] until a mask appears several hundred milliseconds later" <cite data-cite="6027200/AEJJ5749"></cite> (emphasis added). When accuracy is measured as a function of set size, they have an inverse linear relationship, instead of the linear relationship seen between reaction times and set size. When the task is changed as just described, and accuracy is used as the measure, it fulfills the same function as reaction times: we use accuracy to tell us whether attentional processes behaves as we think it does. **But of course CNNs have no built-in mechanism of attention; in effect, they process the entire visual field in parallel, in a single pass, instantaneously**. Therefore, if we see attention-like effects when measuring the accuracy of CNNs, we must either conclude either one of two things: (1) we have not chosen a task and/or metric that gives us insight into the processing going on. Maybe a single scalar metric like accuracy is not actually telling us much about the behavior used to achieve the task, or the internal process of the algorithm that is performing the task. Or (2) we have chosen a good task and/or metric, but there's some other explanation that doesn't require attention, because an algorithm without attention can produce the same behavior that we produce. Maybe attention-like effects are just what happens when brains and CNNs solve this task, because targets are hard to discriminate from distractors, regardless of the features.

## Results
In an initial set of simulations, we trained a CNN to "perform" the visual search task. We used the AlexNet architecture, which has previously been used in studies comparing CNNs to the visual system <cite data-cite="6027200/U77Y74SI"></cite><cite data-cite="6027200/ZT5GRNIN"></cite>. To train AlexNet, we followed a fine-tuning approach, where we re-used weights from training this CNN on the ImageNet dataset. During fine-tuning, the weights in the convolutional layers were frozen, but randomly initialized weights in the final fully-connected layers were update with a very small learning rate. The original softmax layer with 1000 outputs representing ImageNet classes was replaced with a 2-unit softmax layer, where one unit represented a response of "target present" and the other a response of "target absent". This approach is similar to that taken by a previous study <cite data-cite="6027200/9CAW6DFJ"></cite>; we discuss differences in the details of fine-tuning in the methods below. To generate visual search stimuli that could be fed into a neural network, we used a simple Python package (<https://github.com/NickleDave/searchstims>) built on top of the [PyGame](https://www.pygame.org) library . The search displays were divided up into a 5x5 grid of cells, and items were plotted in a randomly chosen subset of cells (as described in <cite data-cite="6027200/QCQFYFDZ"></cite> and shown at <http://search.bwh.harvard.edu/new/data_set_files.html>). The library applied jitter to the items, so that two images could be different even if both images had the same set size and target present condition, and the items were in the exact same cells.  This is typically done with visual search stimuli used in psychophysics experiments, but an added benefit was that this allowed us to create a large set of unique training images (on the order of ten thousand) for each set size and condition, in effect resulting in a sort of data augmentation, as is often used when training neural networks.

### AlexNet shows attentional effects without any attention mechanisms in spite of processing images in parallel
After training AlexNet to perform this task, we found to our surprise that the results qualitatively resemble those seen for human subjects. We plotted accuracy as a function of set size, and saw that slopes of these lines were close to horizontal when AlexNet performed the task given feature search stimuli, where the target was a red vertical rectangle among distractors that were green vertical rectangles. In contrast, the slopes were steep when AlexNet performed the task given spatial conjunction search stimuli, where the target was a digital 2 and the distractors were digital 5s.

```{python}
# %matplotlib inline
import os

from searchnets.plots.figures import ftr_v_spt_conj, mn_slope_by_epoch

TEST_RESULTS_DIR = '../../data/experiment_1/results'
eff_400_results = os.path.join(TEST_RESULTS_DIR,
                            'config_efficient_400_epochs_models/test_alexnet_output.gz')
ineff_400_results = os.path.join(TEST_RESULTS_DIR,
                             'config_inefficient_400_epochs_models/test_alexnet_output.gz')

ftr_v_spt_conj(eff_400_results,
               ineff_400_results,
               epochs=400)
```

**Figure 1. Accuracy of AlexNet when reporting whether a target is present or absent, given efficient or inefficent visual search stimuli** X axis is set size, i.e., number of items in the visual search stimulus shown to the network. Y axis is mean accuracy for each set size, as measured on a previously unseen test set. The test set contained ~3000 images, balanced so that there were equal numbers of images for each set size.  Each line represents one trained network; there are five replicates total.

### Higher accuracy correlates with stronger attentional effects
Given the surprising result that CNNs showed attention-like effects without attention mechanisms and in spite of processing their inputs in a parallel, feedforward fashion, we next asked whether these effects might depend on training. Again to our surprise, we saw that slopes for the feature search got closer to zero as we increased the number of epochs of training, while the slopes for spatial conjunction search got steeper. This was in spite of the fact that training accuracy continued to increase up until about 800 epochs, as measured on a validation set.

```{python}
TEST_RESULTS_DIR = '../../data/experiment_2/results'
epochs_list = [10, 50, 200, 800]
eff_results_list = []
ineff_results_list = []
for epochs in epochs_list:
    eff_results = os.path.join(TEST_RESULTS_DIR, f'config_efficient_{epochs}_epochs_models/test_alexnet_output.gz')
    eff_results_list.append(eff_results)
    ineff_results = os.path.join(TEST_RESULTS_DIR, f'config_inefficient_{epochs}_epochs_models/test_alexnet_output.gz')
    ineff_results_list.append(ineff_results)
    ftr_v_spt_conj(eff_results,
                   ineff_results,
                   epochs=epochs)
```

To quantify this result, we fit slopes to the accuracy vs. set size plots for each trained replicate of AlexNet, for each number of training epochs. As shown below, the slope for feature search decreased as the number of training epochs increased, while the slpe for spatial conjunction search increased. We took the mean slope for each type of search stimulus and subtracted them to get a measure of the difference, and saw that indeed this difference did increase as the number of training epochs increased.

```{python}
mn_slope_by_epoch(eff_results_list, ineff_results_list, epochs_list)
```

## Methods
Alexnet training:
• based on Poder 2017
https://arxiv.org/pdf/1707.09775.pdf
• Stochastic gradient descent
• Learning rate for fully-connected layers: 1e-20
• 10 epochs (as in Poder 2017), and 400 epochs

## References
<div class="cite2c-biblio"></div>

### 2019-02-26 docs/notebooks/results.Rmd
remove a bunch of text from results.ipynb; add citations in Results.Rmd

@@
 
-# Attentional effects in convolutional neural nets without attentional mechanisms
+# Convolutional Neural Networks Show Attention-like Effects When Performing Visual Search Tasks
+
+## Summary
+Convolutional neural networks (CNNs) represent the state of the art in image recognition.
+Many mechanisms proposed to improve the performance of neural networks have been given the name "attention" [@xuShowAttendTell2015].
+This may be surpising to scientists studying the visual system, who typically describe attention as a result of the limitated resources available for processing sensory inputs.
+These scientists have developed behavioral tasks to investigate visual attention.
+Such tasks may provide an alternate way of benchmarking existing neural network architectures, comparing their output with human behavior, and measuring how proposed attentional mechanisms change the output of neural networks.
+Here we adapt a classic form of visual search task so it can be performed by neural networks.
+We measure the accuracy of two CNNs, AlexNet and VGG16, as they report whether a target is present among distractors.
+These architectures were chosen in part because they have been used in several studies that report similarities between activations in the hidden layers of the network and activity in the visual system of the brain when presented with the same visual stimuli.
+Surprisingly, we find that both networks show attention-like effects when performing the task.
 
 ## Introduction
-Neural networks have in the past few years transitioned from being an academic area of artificial intellgience to being powerful tools for engineers working in areas such as image classification, speech-to-text, and automatic translation, to name but a few.
-In part because of this resurgence of research on neural networks, many in neuroscience and cognitive science have returned to using these models to understand the brain.
-Of course there are many questions about what aspects of the brain these models actually capture <cite data-cite="6027200/PAZZI897"></cite><cite data-cite="6027200/KEUFZ7BF"></cite>.
-The goal of the study summarized here is to investigate the extent to which one type of neural network architecture, the convolutional neural network, provides a good model of the visual system in the brain of humans and other primates.
-
-### Connectionist models of cognition and deep neural networks
-In the 1980s, artificial neural networks trained with backpropagation gave rise to [Connectionist](https://en.wikipedia.org/wiki/Connectionism) models of cognition.
-Similar neural network models lie at the heart of the recent rise of deep learning, in large part because of advances in methods for training algorithms, and advances in technology such as GPUs that allow for parallelized training.
-When benchmarked with scalar metrics like accuracy on computer vision tasks such as object recognition, these deep neural networks appear capable of human or super-human levels of skill.
-The first major gains in accuracy were seen on image classification tasks such as ImageNet <cite data-cite="6027200/WGZWMWX7"></cite>, and the deep neural networks that made these advances were built with what are known as convolutional layers.
-As the name implies, these layers perform a convolution-like operation by multiplying filters of learned weights with the input image, producing as outputs feature maps that can be further transformed by succeeding layers.
-
-### Convolutional neural networks as models of the primate visual system
-Many researchers have drawn parallels between the architecture of convolutional neural networks (CNNs) and the architecture of the visual system in the brain <cite data-cite="6027200/8ISULWGF"></cite>, in particular in humans and other primates where this system has been most thoroughly studied.
-Like CNNs, the visual system has a hierarchical structure, and is thought to function in part by performing transformations at each level of this hierarchy so that high-dimensional, low-level features are mapped into low-dimensional abstract representations.
-Several recent studies in the realm of cognitive neuroscience have measured representations in more directly tested the similarities between CNNs and the visual system, by comparing representations in both <cite data-cite="6027200/6D4LKJD9"></cite><cite data-cite="6027200/U77Y74SI"></cite>.
-Representations can be easily extracted from CNNs because they are "transparent box" models--i.e.
-we have access to the "activity" of the individual units from which these models are built when we "show" the CNNs a stimulus.
-In the same studies, brain activity was obtained with electrophysiological or functional imaging methods.
-Such studies have found in general that representations are similar.
-(The comparison depends on representational similarity methods <cite data-cite="6027200/IQ2Y895S"></cite><cite data-cite="6027200/EQTU2P3Z"></cite>, but the concern of this study is not the validity of those methods.)
-
-### Does learning to perform some task with high accuracy constrain neural networks and brains to learn similar representations?
-The authors conclude that one factor that may cause this representational similarity is that the goal-driven tasks provide a constraint on learning.
-In other words, these hierarchichal architectures learn similar representations because those representations provide a way to succesfully carry out behavior.
-Similar claims have been made elsewhere, e.g.
-that the brain and artificial networks both learn similar representations of location when performing navigation tasks <cite data-cite="6027200/LHVT6V6M"></cite><cite data-cite="6027200/SAZAIFBG"></cite>.
-
-The impressive accuracy of deep convolutional neural networks performing image recognition tasks, taken in combination with the apparent correlations between their activity and activity in the primate visual system, raises two questions: (1) how accurately do CNNs perform other tasks that are though to rely on the visual system, and (2) if they perform a task accurately, is that enough evidence to conclude that they provide useful models of the visual system?
-
-### Another type of constraint: controlled behavioral tasks in the lab
-Importantly, the researchers in the studies just described constrain the "task"--passively viewing images--when measuring brain activity, with the goal of being able to directly compare brain activity to activity in neural networks.
-Specifically, the subjects are only shown the visual stimuli very briefly, typically about 100 ms for monkey electophysiology experiments and 300 ms for human neuroimaging experiments.
-There are a couple of reasons for this.
-The first is to allow for an apples-to-apples comparison between the brain and CNNs.
-Although the brain is highly recurrent, CNNs are purely **feedforward**.
-The researchers want to claim that they have only elicited  initial feedforward activity in the visual system by just briefly showing subjects the stimuli, without allowing feedback to influence activity, and by extension, the representation.
-The second reason for the constraint of showing images only briefly is more historical.
-Classically, the visual system has been studied with what has been termed **passive vision** approach <cite data-cite="6027200/B49GIRUD"></cite>.
-This approach could be characterized as conceiving of the eye as unmoving, with a still image falling upon it that is then processed in parallel across the scene to extract information.
-These types of models have obviously been successful, and the incredible accuracy of CNNs might even be described as the culmination of this approach.
-The point here is not to debate about issues with passive vision models; most readers probably take it for granted that humans move their eyes, and anyone who has tried to walk through a crowd while texting on their phone will know that there are limitations on our ability to process the entire visual scene in parallel.
-
-### How much can accuracy on a task tell us about cognition? Visual search tasks as a case study
-The point is that the same constraint that was used to make apples-to-apples comparisons between CNNs and the brains, briefly showing stimuli, has also been used in visual search tasks in the laboratory <cite data-cite="6027200/AEJJ5749"></cite>, in part because of the passive vision approach with which the visual system has typically been studied.
-The fact that the same constraint was applied to visual search tasks lets us use the same logic as studies which conclude that goal-driven tasks constrain learning to produce useful representations.
-We can take advantage of this shared constraint to invert the logic and state a new hypothesis: if a CNN has learned the same representation as the brain, then its behavior when performing tasks should match our behavior.
-In other words, we are asking whether we can extend a claim about CNNs, that they provide good models of the visual system, by examining how they behave when performing a visual search task.
-There are several objections that could be made to doing so, including the question of whether the task we'll look at is really any different from image recognition.
-But let's pretend for a moment that the question is worth asking.
-Why is it worth doing so?
-To consider more generally how much we can learn about cognition by measuring the behavior of our model with a scalar metric applied to a highly constrained task.
-
-What is that task?
-In the "classical" form of the visual search paradigm <cite data-cite="6027200/AEJJ5749"></cite>, experimenters present a stimulus to a subject made up of discrete items, i.e.
+### Visual search tasks for investigating attention
+In the "classical" form of the visual search paradigm [@wolfeVisualSearch1998], experimenters present a subject with a search display on which there are discrete items, i.e.
 objects.
-The subject is asked to report whether a target is present among distractors-- the experimenter sets the probability that the target will be present on any given trial.
+The subject is asked to report whether a target is present among distractors.
+The experimenter sets the probability that the target will be present on any given trial.
 Experimenters measure the reaction time and/or accuracy of the subjects.
@@
 "seconds per item".
-Slopes are almost horizontal when subjects are shown visual search stimuli where targets differ from distractors in some clearly salient feature, e.g.
-a red target among green distractors, and slopes become much steeper when targets differ from distractors in some harder-to-distinguish conjunction of features, e.g., a digital 2 among digital 5s.
+Slopes are almost horizontal when subjects are shown visual search stimuli where targets differ from distractors in some clearly salient **feature**, e.g.
+a red target among green distractors, and slopes become much steeper when targets differ from distractors in some harder-to-distinguish **spatial conjunction** of features, e.g., a digital 2 among digital 5s.
 Some examples of these stimuli are shown below, generated with https://github.com/NickleDave/searchstims.
@@
 
-This use of search slope as metric led to one of the most influential early models of visual search: feature integration theory <cite data-cite="6027200/KYF4NSZP"></cite>.
+This use of search slope as metric led to one of the most influential early models of visual search: feature integration theory [@treismanFeatureintegrationAttention1980].
 Essentially, feature integration theory posits that the brain can search the visual field in parallel when a single feature distinguishes the target from distractors, but that a conjunction of multiple features requires serial processing.
-However, Wolfe <cite data-cite="6027200/LJEPEDWB"></cite>(1998) showed that there is no clear cut segregation of slopes into one type or the other, and that in fact the distributions of slope overlap so much it is not possible to distinguish them.
-He has proposed talking about visual search in terms of a spectrum of of speeds that depend in some way on features: when a salient feature like color is what discriminates the target from the distractor, this allows for **efficient** search, like in the visual search stimulus on the left.
-When a complex conjunction of features distinguishes the target from the distractor, as in the visual search stimulus on the right, this results in **inefficient** search.
+However, Wolfe [@wolfeWhatCanMillion1998] showed that there is no clear cut segregation of slopes into one type or the other, and that in fact the distributions of slope overlap so much it is not possible to distinguish them.
+He has proposed talking about visual search in terms of a spectrum of of speeds that depend in some way on features: when a salient feature like color is what discriminates the target from the distractor, this allows for efficient search, like in the visual search stimulus on the left.
+When a complex conjunction of features distinguishes the target from the distractor, as in the visual search stimulus on the right, this results in inefficient search.
+In the results below, however, we continue to refer to the stimuli by the names "feature" and "spatial conjunction" search.
 
@@
 In some cases, the stimulus is shown only briefly to the subject, and then after a variable length of time a masking stimulus such as white noise is shown.
-During the period between the stimulus with a set of items and the masking stimulus, the "*internal representation* of that stimulus can be searched \[by the subject\] until a mask appears several hundred milliseconds later" <cite data-cite="6027200/AEJJ5749"></cite> (emphasis added).
-When accuracy is measured as a function of set size, they have an inverse linear relationship, instead of the linear relationship seen between reaction times and set size.
-When the task is changed as just described, and accuracy is used as the measure, it fulfills the same function as reaction times: we use accuracy to tell us whether attentional processes behaves as we think it does.
+During the period between the stimulus with a set of items and the masking stimulus, the "*internal representation* of that stimulus can be searched \[by the subject\] until a mask appears several hundred milliseconds later" [@wolfeVisualSearch1998] (emphasis added).
+When the experiment is carried out this way, accuracy is measured instead of reaction time [@bergenRapidDiscriminationVisual1983; @ecksteinLowerVisualSearch1998].
+Accuracy has an inverse linear relationship with set size in these experiments (as opposed to the linear relationship seen between reaction times and set size).
+In spite of this change, when accuracy is used as the measure, it fulfills the same function as reaction times: we use accuracy to tell us whether attentional processes behaves as we think they do.
 **But of course CNNs have no built-in mechanism of attention; in effect, they process the entire visual field in parallel, in a single pass, instantaneously**.
@@
 In an initial set of simulations, we trained a CNN to "perform" the visual search task.
-We used the AlexNet architecture, which has previously been used in studies comparing CNNs to the visual system <cite data-cite="6027200/U77Y74SI"></cite><cite data-cite="6027200/ZT5GRNIN"></cite>.
+We used the AlexNet architecture, which has previously been used in studies comparing CNNs to the visual system [@khaligh-razaviDeepSupervisedNot2014].
 To train AlexNet, we followed a fine-tuning approach, where we re-used weights from training this CNN on the ImageNet dataset.
@@
 The original softmax layer with 1000 outputs representing ImageNet classes was replaced with a 2-unit softmax layer, where one unit represented a response of "target present" and the other a response of "target absent".
-This approach is similar to that taken by a previous study <cite data-cite="6027200/9CAW6DFJ"></cite>; we discuss differences in the details of fine-tuning in the methods below.
+This approach is similar to that taken by a previous study [@poderCapacityLimitationsVisual2017]; we discuss differences in the details of fine-tuning in the methods below.
 To generate visual search stimuli that could be fed into a neural network, we used a simple Python package (<https://github.com/NickleDave/searchstims>) built on top of the [PyGame](https://www.pygame.org) library .
@@
 ## References
-<div class="cite2c-biblio"></div>
+
+## Acknowledgments

### 2019-02-26 docs/notebooks/paper.Rmd
text from results now in paper.Rmd & paper.ipynb

--- 2019-02-25 docs/notebooks/results.Rmd
@@
 
-# Attentional effects in convolutional neural nets without attentional mechanisms
+# Convolutional Neural Networks Show Attention-like Effects When Performing Visual Search Tasks
 
 ## Introduction
+Convolutional neural networks (CNNs) represent the state of the art in image recognition.
+Many mechanisms proposed to improve the performance on neural networks have been given the name "attention" [@xuShowAttendTell2015].
+This may be surpising to scientists studying the visual system, who typically describe attention as a result of the limitated resources available for processing sensory inputs.
+These scientists have developed
+
+## Convolutional neural networks as models of the visual system
 Neural networks have in the past few years transitioned from being an academic area of artificial intellgience to being powerful tools for engineers working in areas such as image classification, speech-to-text, and automatic translation, to name but a few.
 In part because of this resurgence of research on neural networks, many in neuroscience and cognitive science have returned to using these models to understand the brain.
-Of course there are many questions about what aspects of the brain these models actually capture <cite data-cite="6027200/PAZZI897"></cite><cite data-cite="6027200/KEUFZ7BF"></cite>.
-The goal of the study summarized here is to investigate the extent to which one type of neural network architecture, the convolutional neural network, provides a good model of the visual system in the brain of humans and other primates.
-
-### Connectionist models of cognition and deep neural networks
+Of course there are many questions about what aspects of the brain these models actually capture [@cichyDeepNeuralNetworks2019] and there are many suggestions on how to make their behavior closer to our own [@lakeBuildingMachinesThat2017; @marcusDeepLearningCritical2018].
 In the 1980s, artificial neural networks trained with backpropagation gave rise to [Connectionist](https://en.wikipedia.org/wiki/Connectionism) models of cognition.
@@
 
-### Convolutional neural networks as models of the primate visual system
 Many researchers have drawn parallels between the architecture of convolutional neural networks (CNNs) and the architecture of the visual system in the brain <cite data-cite="6027200/8ISULWGF"></cite>, in particular in humans and other primates where this system has been most thoroughly studied.
@@
 
-### How much can accuracy on a task tell us about cognition? Visual search tasks as a case study
-The point is that the same constraint that was used to make apples-to-apples comparisons between CNNs and the brains, briefly showing stimuli, has also been used in visual search tasks in the laboratory <cite data-cite="6027200/AEJJ5749"></cite>, in part because of the passive vision approach with which the visual system has typically been studied.
+### Visual search tasks for investigating attention
+The same constraint that was used to make apples-to-apples comparisons between CNNs and the brains, briefly showing stimuli, has also been used in visual search tasks in the laboratory <cite data-cite="6027200/AEJJ5749"></cite>, in part because of the passive vision approach with which the visual system has typically been studied.
 The fact that the same constraint was applied to visual search tasks lets us use the same logic as studies which conclude that goal-driven tasks constrain learning to produce useful representations.

### 2019-02-26 docs/notebooks/results.Rmd
rewrite 'Summary' as 'Abstract', flesh out. Possibly too strong

@@
 
-## Summary
-Convolutional neural networks (CNNs) represent the state of the art in image recognition.
+## Abstract
+Convolutional neural networks (CNNs) provide cutting edge tools for computer vision tasks like image classifcation and scene segmentation.
+However, they suffer from known flas.
 Many mechanisms proposed to improve the performance of neural networks have been given the name "attention" [@xuShowAttendTell2015].
 This may be surpising to scientists studying the visual system, who typically describe attention as a result of the limitated resources available for processing sensory inputs.
-These scientists have developed behavioral tasks to investigate visual attention.
-Such tasks may provide an alternate way of benchmarking existing neural network architectures, comparing their output with human behavior, and measuring how proposed attentional mechanisms change the output of neural networks.
+Attention in the brain has been primarily investiaged using visual search tasks.
+These tasks provide an alternate way of benchmarking existing neural network architectures, comparing their output with human behavior, and measuring how proposed attentional mechanisms change the output of neural networks.
+We are pursuing these alternate benchmarks for attentional mechanisms in the course of investigations of continual learning algorithms.
 Here we adapt a classic form of visual search task so it can be performed by neural networks.
 We measure the accuracy of two CNNs, AlexNet and VGG16, as they report whether a target is present among distractors.
-These architectures were chosen in part because they have been used in several studies that report similarities between activations in the hidden layers of the network and activity in the visual system of the brain when presented with the same visual stimuli.
+These architectures were chosen in part because they have been used in several studies that report similarities between activations in the hidden layers of the network and activity in the visual system of the brain, when presented with the same visual stimuli.
 Surprisingly, we find that both networks show attention-like effects when performing the task.
+When the target resembles the distractors, then a network's accuracy drops as the number of distractors increases.
+We do not find that these results can be explained by known factors to affect CNNs, such as edge effects.
+While it might seem like a given that target-distractor similarity impairs accuracy, qualtiatively similar findings in studies of human subjects have been used to conclude that the brain requires serial processing mechanisms.
+Given that we know that CNNs process their inputs effectively in parallel, we can conclusively attribute the impaired accuracy here to target-distractor similarity.
+Therefore, we suggest that any mechanism which can be reformulated as a learned weighting processed in parallel is not categorically different from current CNNs, and is not likely to improve performance, since it will just produce the kind of attention effects we describe here.
 

### 2019-02-27 docs/notebooks/results.Rmd

--- 2019-02-26 docs/notebooks/results.Rmd
@@
 
-## Summary
-Convolutional neural networks (CNNs) represent the state of the art in image recognition.
+## Abstract
+Convolutional neural networks (CNNs) provide cutting edge tools for computer vision tasks like image classification and scene segmentation.
+However, they suffer from known flaws.
 Many mechanisms proposed to improve the performance of neural networks have been given the name "attention" [@xuShowAttendTell2015].
-This may be surpising to scientists studying the visual system, who typically describe attention as a result of the limitated resources available for processing sensory inputs.
-These scientists have developed behavioral tasks to investigate visual attention.
-Such tasks may provide an alternate way of benchmarking existing neural network architectures, comparing their output with human behavior, and measuring how proposed attentional mechanisms change the output of neural networks.
+This may be surpising to scientists studying the visual system, who have traditionally thought of attention as an artifact of the limited resources available for processing sensory inputs.
+Attention in the brain has been primarily investiaged using visual search tasks.
+These tasks may provide an alternate way of benchmarking existing neural network architectures, comparing their output with human behavior, and measuring how any proposed attentional mechanisms change neural network outputs.
+We are pursuing these alternate benchmarks for attentional mechanisms in the course of our investigations of continual learning algorithms.
 Here we adapt a classic form of visual search task so it can be performed by neural networks.
-We measure the accuracy of two CNNs, AlexNet and VGG16, as they report whether a target is present among distractors.
-These architectures were chosen in part because they have been used in several studies that report similarities between activations in the hidden layers of the network and activity in the visual system of the brain when presented with the same visual stimuli.
-Surprisingly, we find that both networks show attention-like effects when performing the task.
+We measure the accuracy of two CNN architectures, AlexNet and VGG16, as they report whether a target is present among distractors.
+These architectures were chosen in part because they have been used in several studies that report similarities between activations in the hidden layers of the network and activity in the visual system of the brain, when both are presented with the same images.
+Surprisingly, we find that both AlexNet and VGG16 show attention-like effects when performing the task.
+When the target resembles the distractors, then a network's accuracy drops as the number of distractors increases.
+These dependency of accuracy on the number of items is often referred to as "set size effects".
+We do not find that these set size effects can be explained by known factors to affect CNNs, such as edge effects.
+It might seem obvious that target-distractor similarity would impair accuracy, but much of the visual attention literature boils down to a debate about the extent to which these effects can be attributed to cognitive processes, e.g.
+whether those processes are serial or parallel and noisy.
+Here, since we know that CNNs process their inputs in parallel, we can conclusively attribute the impaired accuracy here to target-distractor similarity.
+Therefore, we suggest that any mechanism which can be reformulated as a learned weighting processed in parallel is not categorically different from current CNNs.
+Such mechanisms may not ever improve performance of CNNs, given that the attentional effects they produce will be of the the kind we describe here.
 

### 2019-02-27 docs/notebooks/results.Rmd
update results.ipynb + .Rmd

@@
       format_version: '1.0'
-      jupytext_version: 1.0.0
+      jupytext_version: 1.0.1
   kernelspec:
@@
 TEST_RESULTS_DIR = '../../data/experiment_1/results'
-eff_400_results = os.path.join(TEST_RESULTS_DIR,
-                            'config_efficient_400_epochs_models/test_alexnet_output.gz')
-ineff_400_results = os.path.join(TEST_RESULTS_DIR,
-                             'config_inefficient_400_epochs_models/test_alexnet_output.gz')
+ftr_400_results = os.path.join(TEST_RESULTS_DIR,
+                            'feature_search_alexnet_models/test_alexnet_trained_400_epochs.gz')
+sptl_400_results = os.path.join(TEST_RESULTS_DIR,
+                             'spatial_config_search_alexnet_models/test_alexnet_trained_400_epochs.gz')
 
-ftr_v_spt_conj(eff_400_results,
-               ineff_400_results,
+ftr_v_spt_conj(ftr_400_results,
+               sptl_400_results,
                epochs=400)
@@
 ```{python}
-TEST_RESULTS_DIR = '../../data/experiment_2/results'
-epochs_list = [10, 50, 200, 800]
-eff_results_list = []
-ineff_results_list = []
+TEST_RESULTS_DIR = '../../data/experiment_1/results'
+epochs_list = [10, 50, 200, 400, 800]
+ftr_results_list = []
+sptl_results_list = []
 for epochs in epochs_list:
-    eff_results = os.path.join(TEST_RESULTS_DIR, f'config_efficient_{epochs}_epochs_models/test_alexnet_output.gz')
-    eff_results_list.append(eff_results)
-    ineff_results = os.path.join(TEST_RESULTS_DIR, f'config_inefficient_{epochs}_epochs_models/test_alexnet_output.gz')
-    ineff_results_list.append(ineff_results)
-    ftr_v_spt_conj(eff_results,
-                   ineff_results,
+    ftr_results = os.path.join(TEST_RESULTS_DIR, f'feature_search_alexnet_models/test_alexnet_trained_{epochs}_epochs.gz')
+    ftr_results_list.append(ftr_results)
+    sptl_results = os.path.join(TEST_RESULTS_DIR, f'spatial_config_search_alexnet_models/test_alexnet_trained_{epochs}_epochs.gz')
+    sptl_results_list.append(sptl_results)
+    ftr_v_spt_conj(ftr_results,
+                   sptl_results,
                    epochs=epochs)
@@
 set size plots for each trained replicate of AlexNet, for each number of training epochs.
-As shown below, the slope for feature search decreased as the number of training epochs increased, while the slpe for spatial conjunction search increased.
+As shown below, the slope for feature search decreased as the number of training epochs increased, while the slope for spatial conjunction search increased.
 We took the mean slope for each type of search stimulus and subtracted them to get a measure of the difference, and saw that indeed this difference did increase as the number of training epochs increased.
@@
 ```{python}
-mn_slope_by_epoch(eff_results_list, ineff_results_list, epochs_list)
+mn_slope_by_epoch(ftr_results_list, sptl_results_list, epochs_list)
 ```

### 2019-03-01 docs/notebooks/results.Rmd
update results notebooks with VGG16 figs

@@
                    epochs=epochs)
+alexnet_results = {
+    'ftr': ftr_results_list,
+    'sptl': sptl_results_list,
+}
 ```
@@
 ```{python}
-mn_slope_by_epoch(ftr_results_list, sptl_results_list, epochs_list)
+mn_slope_by_epoch(alexnet_results['ftr'], alexnet_results['sptl'], epochs_list)
+```
+
+### Other CNN architectures also show similar effects
+To test whether these effects resulted from the architecture of AlexNet or were more general, we repeated the simulations using the VGG16 architecture.
+
+```{python}
+TEST_RESULTS_DIR = '../../data/experiment_1/results'
+epochs_list = [10, 50, 200, 400, 800]
+ftr_results_list = []
+sptl_results_list = []
+for epochs in epochs_list:
+    ftr_results = os.path.join(TEST_RESULTS_DIR, f'feature_search_vgg16_models/test_VGG16_trained_{epochs}_epochs.gz')
+    ftr_results_list.append(ftr_results)
+    sptl_results = os.path.join(TEST_RESULTS_DIR, f'spatial_config_search_vgg16_models/test_VGG16_trained_{epochs}_epochs.gz')
+    sptl_results_list.append(sptl_results)
+vgg16_results = {
+    'ftr': ftr_results_list,
+    'sptl': sptl_results_list,
+}
+
+ftr_v_spt_conj(vgg16_results['ftr'][3],
+               vgg16_results['sptl'][3],
+               epochs=epochs_list[3])
+```
+
+We again saw this effect became stronger the more that the VGG16 architecture was trained.
+
+```{python}
+mn_slope_by_epoch(vgg16_results['ftr'], vgg16_results['sptl'], epochs_list)
 ```

### 2019-03-01 docs/notebooks/results.Rmd
update 'abstract' in results

@@
       format_version: '1.0'
-      jupytext_version: 1.0.1
+      jupytext_version: 1.0.0
   kernelspec:
@@
 Convolutional neural networks (CNNs) provide cutting edge tools for computer vision tasks like image classification and scene segmentation.
-However, they suffer from known flaws.
-Many mechanisms proposed to improve the performance of neural networks have been given the name "attention" [@xuShowAttendTell2015].
-This may be surpising to scientists studying the visual system, who have traditionally thought of attention as an artifact of the limited resources available for processing sensory inputs.
-Attention in the brain has been primarily investiaged using visual search tasks.
+Many mechanisms proposed to improve the performance of neural networks have been given the name "attention".
+Scientists studying the visual system have traditionally thought of attention as an artifact of the limited resources available for processing sensory inputs.
+Attention in the brain has been primarily investigated using visual search tasks.
 These tasks may provide an alternate way of benchmarking existing neural network architectures, comparing their output with human behavior, and measuring how any proposed attentional mechanisms change neural network outputs.
@@
 We measure the accuracy of two CNN architectures, AlexNet and VGG16, as they report whether a target is present among distractors.
-These architectures were chosen in part because they have been used in several studies that report similarities between activations in the hidden layers of the network and activity in the visual system of the brain, when both are presented with the same images.
-Surprisingly, we find that both AlexNet and VGG16 show attention-like effects when performing the task.
+Surprisingly, we find that both AlexNet and VGG16 show attention-like effects when performing the visual search task.
 When the target resembles the distractors, then a network's accuracy drops as the number of distractors increases.
-These dependency of accuracy on the number of items is often referred to as "set size effects".
+This dependency of accuracy on the number of items is often referred to as "set size effects".
 We do not find that these set size effects can be explained by known factors to affect CNNs, such as edge effects.
-It might seem obvious that target-distractor similarity would impair accuracy, but much of the visual attention literature boils down to a debate about the extent to which these effects can be attributed to cognitive processes, e.g.
-whether those processes are serial or parallel and noisy.
-Here, since we know that CNNs process their inputs in parallel, we can conclusively attribute the impaired accuracy here to target-distractor similarity.
-Therefore, we suggest that any mechanism which can be reformulated as a learned weighting processed in parallel is not categorically different from current CNNs.
-Such mechanisms may not ever improve performance of CNNs, given that the attentional effects they produce will be of the the kind we describe here.
+We then show that these attentional effects persist even when controlling for how discriminable the targets are from the distractors, by using an analysis of visual search tasks based on signal detection theory.
+Hence these "attentional" effects can be attributed to the decision criteria of the network.
+In the discussion, we consider how this type of analysis can be extended to stimuli besides visual search criteria.
 

### 2019-03-01 src/searchnets/utils/general.py
add make_targz and targz_dirs functions to searchstims.utils.general

@@ -1,4 +1,6 @@
 import csv
+import tarfile
+import os
 
 import pandas
 from statsmodels.formula.api import ols
@@ -50,3 +52,33 @@ def anova(csv_fname):
     model = ols('acc ~ C(stim_type) + epochs + set_size', data).fit()
     print(model.summary())
     return model
+
+
+def make_targz(output_filename, source_dir):
+    """make a .tar.gz file from a directory, i.e. a gzip-compressed archive"""
+    with tarfile.open(name=output_filename, mode="w:gz") as tar:
+        tar.add(source_dir, arcname=os.path.basename(source_dir))
+
+
+def targz_dirs(dir_names=('checkpoints', 'configs', 'data_prepd_for_nets', 'results', 'visual_search_stimuli'),
+               path='.'):
+    """makes a .tar.gz archive for a list of directories, for uploading to figshare.
+
+    Parameters
+    ----------
+    dir_names : list
+        of directory names.
+        Default is ['checkpoints', 'configs', 'data_prepd_for_nets', 'results', 'visual_search_stimuli'].
+    path : str
+        Path to parent directory where all the directories in dir_names can be found.
+        Default is '.', i.e. current working directory.
+
+    Notes
+    -----
+    Basically just a wrapper around make_targz for convenience;
+    assumes each .tar.gz output should have the name 'directory_name.tar.gz'
+    """
+    for dir_name in dir_names:
+        output_filename = dir_name + '.tar.gz'
+        print(f'making {output_filename}')
+        make_targz(output_filename=output_filename, source_dir=os.path.join(path, dir_name))

### 2019-03-02 docs/notebooks/results.Rmd

@@
 Convolutional neural networks (CNNs) provide cutting edge tools for computer vision tasks like image classification and scene segmentation.
-Many mechanisms proposed to improve the performance of neural networks have been given the name "attention".
-Scientists studying the visual system have traditionally thought of attention as an artifact of the limited resources available for processing sensory inputs.
-Attention in the brain has been primarily investigated using visual search tasks.
-These tasks may provide an alternate way of benchmarking existing neural network architectures, comparing their output with human behavior, and measuring how any proposed attentional mechanisms change neural network outputs.
-We are pursuing these alternate benchmarks for attentional mechanisms in the course of our investigations of continual learning algorithms.
-Here we adapt a classic form of visual search task so it can be performed by neural networks.
-We measure the accuracy of two CNN architectures, AlexNet and VGG16, as they report whether a target is present among distractors.
+Many mechanisms proposed to improve the accuracy of CNNs been given the name "attention", although it is becoming clear that accuracy alone does not explain how CNNs perform in real-world situations.
+Attention has a long history of being studied by cognitive scientists and neuroscientists in the context of vision.
+The CNNs developed in the past few years are also increasingly being used as models of the visual system in the brain.
+Here we study whether models of attention in the brain, and visual search tasks used to investigate attention, provide an alternate way of benchmarking existing CNNs.
+We are pursuing these alternate benchmarks in the course of our investigations of attention-like mechanisms for continual learning algorithms.
+We adapt a classic form of visual search task so it can be performed by neural networks, and measure the accuracy of two CNN architectures, AlexNet and VGG16, as they report whether a target is present among distractors.
 Surprisingly, we find that both AlexNet and VGG16 show attention-like effects when performing the visual search task.
@@
 We do not find that these set size effects can be explained by known factors to affect CNNs, such as edge effects.
-We then show that these attentional effects persist even when controlling for how discriminable the targets are from the distractors, by using an analysis of visual search tasks based on signal detection theory.
-Hence these "attentional" effects can be attributed to the decision criteria of the network.
-In the discussion, we consider how this type of analysis can be extended to stimuli besides visual search criteria.
+We then adapt an analysis of visual search tasks based on signal detection theory to show that set size effects persist even when controlling for how discriminable the targets are from the distractors.
+This analysis suggests the "attentional" effects can be attributed to the decision criteria used by the network.
+We consider how this type of analysis can be extended to images besides visual search stimuli.
+Then we discuss the implications of our results for comparing CNN output with human behavior, and measuring how any proposed attentional mechanisms change neural network outputs.
 

### 2019-03-02 src/searchnets/train.py
searchnets.train makes 'trained_x_epochs' subdirectories in MODEL_SAVE_PATH

@@ -105,6 +105,8 @@ def train(config):
     np.random.seed(random_seed)  # for shuffling in batch_generator
     tf.random.set_random_seed(random_seed)
 
+    model_save_path = config['TRAIN']['MODEL_SAVE_PATH']
+
     for epochs in epochs_list:
         print(f'training {net_name} model for {epochs} epochs')
         for net_number in range(number_nets_to_train):
@@ -193,10 +195,11 @@ def train(config):
                     else:
                         print()
 
-                savepath = os.path.join(config['TRAIN']['MODEL_SAVE_PATH'],
-                                        'net_number_{}'.format(net_number))
+                savepath = os.path.join(model_save_path,
+                                        f'trained_{epochs}_epochs',
+                                        f'net_number_{net_number}')
                 if not os.path.isdir(savepath):
-                    os.makedirs(savepath)
+                    os.makedirs(savepath, exist_ok=True)
                 print(f'Saving model in {savepath}')
                 ckpt_name = os.path.join(savepath, f'{net_name}-model.ckpt')
                 saver.save(sess, ckpt_name, global_step=epochs)

### 2019-03-02 src/searchnets/test.py
searchnets.test knows to look for 'trained_x_epochs' subdirectories

@@ -55,6 +55,8 @@ def test(config):
         raise TypeError("'EPOCHS' option in 'TRAIN' section of config.ini file parsed "
                         f"as invalid type: {type(epochs_list)}")
 
+    model_save_path = config['TRAIN']['MODEL_SAVE_PATH']
+
     for epochs in epochs_list:
         print(f'measuring accuracy on test set for {net_name} model trained for {epochs} epochs')
 
@@ -86,10 +88,11 @@ def test(config):
                 }
 
                 saver = tf.train.Saver()
-                savepath = os.path.join(config['TRAIN']['MODEL_SAVE_PATH'],
-                                        'net_number_{}'.format(net_number))
-                print('Loading model from %s' % savepath)
-                ckpt_path = os.path.join(savepath, f'{net_name}-model.ckpt-{epochs}')
+                restore_path = os.path.join(model_save_path,
+                                        f'trained_{epochs}_epochs',
+                                        f'net_number_{net_number}')
+                print(f'Loading model from {restore_path}')
+                ckpt_path = os.path.join(restore_path, f'{net_name}-model.ckpt-{epochs}')
                 saver.restore(sess, ckpt_path)
 
                 y_pred_all = []
@@ -116,8 +119,8 @@ def test(config):
                 acc_per_set_size_per_model.append(acc_per_set_size)
                 # and insert into dictionary where model name is key
                 # and list of accuracies per set size is the "value"
-                acc_per_set_size_model_dict[savepath] = acc_per_set_size
-                predictions_per_model_dict[savepath] = y_pred_all
+                acc_per_set_size_model_dict[restore_path] = acc_per_set_size
+                predictions_per_model_dict[restore_path] = y_pred_all
 
         acc_per_set_size_per_model = np.asarray(acc_per_set_size_per_model)
         acc_per_set_size_per_model = np.squeeze(acc_per_set_size_per_model)

### 2019-03-03 data/runall.sh
add /data/runall.sh

@@ -1,0 +1,15 @@
+#!/usr/bin/env bash
+
+source ~/anaconda3/bin/activate searchnets
+
+# create visual search stimuli used to train neural nets
+searchstims ./configs/searchstims/config_feature_alexnet.ini
+searchstims ./configs/searchstims/config_spatial_config_alexnet.ini
+searchstims ./configs/searchstims/config_feature_vgg16.ini
+searchstims ./configs/searchstims/config_spatial_config_vgg16.ini
+
+# prepare data sets for training neural nets, then train, then test accuracy of trained nets
+searchnets all ./configs/searchnets_feature_search_alexnet.ini
+searchnets all ./configs/searchnets_feature_search_vgg16.ini
+searchnets all ./configs/searchnets_spatial_config_search_alexnet.ini
+searchnets all ./configs/searchnets_spatial_config_search_vgg16.ini

### 2019-03-03 src/searchnets/utils/general.py
remove csv + anova functions from searchnets/utils/general -- not using

@@ -1,57 +1,5 @@
-import csv
 import tarfile
 import os
-
-import pandas
-from statsmodels.formula.api import ols
-
-
-def make_csv(eff_accs_10_epochs, ineff_accs_10_epochs, eff_accs_400_epochs, ineff_accs_400_epochs,
-             set_sizes=(1, 2, 4, 8), csv_fname='results.csv'):
-    """make csv from results
-
-    Parameters
-    ----------
-    eff_accs_10_epochs : numpy.ndarray
-    ineff_accs_10_epochs : numpy.ndarray
-    eff_accs_400_epochs : numpy.ndarray
-    ineff_accs_400_epochs : numpy.ndarray
-    set_sizes : list
-        of set sizes, in same order they appear in columns of array.
-        Default is [1, 2, 4, 8].
-    csv_fname : str
-        name of .csv file to save
-
-    Returns
-    -------
-    None
-    """
-    header = ["", "stim_type", "epochs", "replicate", "set_size", "acc"]
-    out = [header]
-
-    id = 1
-
-    mats = [eff_accs_10_epochs, ineff_accs_10_epochs, eff_accs_400_epochs, ineff_accs_400_epochs]
-    epochs = [10, 10, 400, 400]
-    stim_type = ['eff', 'ineff', 'eff', 'ineff']
-    for mat, epoch, stim in zip(mats, epochs, stim_type):
-        for replicate, mat_row in enumerate(mat):
-            for acc, set_size in zip(mat_row, set_sizes):
-                csv_row = [id, stim, epoch, replicate, set_size, acc]
-                out.append(csv_row)
-                id += 1
-
-    with open(csv_fname, 'w', newline='') as f:
-        writer = csv.writer(f)
-        writer.writerows(out)
-
-
-def anova(csv_fname):
-    """run anova on data in csv file generated with make_csv"""
-    data = pandas.read_csv(csv_fname)
-    model = ols('acc ~ C(stim_type) + epochs + set_size', data).fit()
-    print(model.summary())
-    return model
 
 
 def make_targz(output_filename, source_dir):

### 2019-03-03 docs/notebooks/results.Rmd
first add of psychometric approach to results.ipynb

--- 2019-03-01 docs/notebooks/results.Rmd
@@
 
+### A psychophysics analysis of visual search shows set size affects remain even when controlling for how discriminable targets are from distractors
+The Feature Integration Theory of visual search attributed set size effects to serial processing (Treisman).
+This conclusion made sense based on the result that reaction time increased when discriminating targets from distractors required considering more than one feature, e.g.
+both color and orientation.
+However, later work showed that reaction time effects disappear as targets and distractors become more easily discriminable along one feature dimension (Duncan paper).
+Later researchers, using methods from psychophysics and signal-detection theory, carefully controlled for the discriminability of targets and distractors (Palmer review).
+This work provided important alternative explanations for the results seen in visual search experiments: attentional effects can be explained in terms of completely parallel but noisy representations that are used by a decision rule (Palmer, Eckstein).
+This approach also explains errors in terms of low thresholds used by decision making criteria.
+We adapted these psychophysics methods to see if they offered additional insights into the attention-like effects we saw in CNNs.
+#### Psychometric curves
+First we generated 'feature' search stimuli--red vertical rectangle v. green vertical rectangle.  
+<img src="../static/psychometric_function_fits.png" alt="efficient" style="width: 500px;"/>
+#### Threshold v. set size
+We then used the psychometric curves to find what level of target-distractor discriminability gave rise to 75% accuracy, known as the "difference threshold".
+We found this difference threshold for each trained network and each set size.
+<img src="../static/threshold_v_setsize.png" alt="efficient" style="width: 600px;"/>  
+Here we see that the threshold goes up with set size, meaning more discriminable target-distractor pairs are required as set size increases.
+
+For a detailed walkthrough of how we did this, please see [this notebook](./psychometrics.ipynb).
+
 ## Methods

### 2019-03-03 docs/notebooks/results.Rmd
write a bit more in psychometric section, fix link to psychometric.ipynb

@@
 #### Psychometric curves
-First we generated 'feature' search stimuli--red vertical rectangle v. green vertical rectangle.  
+First we generated 'feature' search stimuli--red vertical rectangle v.
+green vertical rectangle--where the red and green were more or less discriminable.
+We did so by taking advantage of RGB color space: by adding more red to the green color, and vice versa, the colors of the target and distractor can be moved closer to each other from the two red and green verticies of the triangle that defines RGB space.
+(When the color has "all red" and "all green", it becomes yellow, so the stimuli we generated moved closer and closer to yellow.) We then used these stimuli to measure the accuracy of the neural nets we trained as the target and distractor became more and more similar.
 <img src="../static/psychometric_function_fits.png" alt="efficient" style="width: 500px;"/>
@@
 Here we see that the threshold goes up with set size, meaning more discriminable target-distractor pairs are required as set size increases.
+In signal detection theory terms, this result is consistent with any theory of attention where decisions are made with a low threshold.
 
-For a detailed walkthrough of how we did this, please see [this notebook](./psychometrics.ipynb).
+For a detailed walkthrough of how we did this, please see [this notebook](./psychometric.ipynb).
 

### 2019-03-03 src/searchnets/train.py
add comment in train.py about global_variables_initializer

@@ -164,6 +164,11 @@ def train(config):
                 y_data = np.array(training_set[1])
                 training_loss = []
 
+                # note that running global_variables_initializer() will initialize at random all the variables in the
+                # model that are in the `init_layer` list passed as an argument when the model was instantiated, **and**
+                # assign the pre-trained weights + biases to the other variables that are not in `init_layer`. This can
+                # be confusing if you are thinking of "initialize" as synonymous with "at random", but fear not, the
+                # pre-trained weights are in fact being loaded
                 sess.run(tf.global_variables_initializer())
 
                 for epoch in range(1, epochs + 1):

### 2019-03-03 docs/notebooks/results.Rmd
change abstract back in results.ipynb; somehow overwrote last edit

--- 2019-03-02 docs/notebooks/results.Rmd
@@
 Convolutional neural networks (CNNs) provide cutting edge tools for computer vision tasks like image classification and scene segmentation.
-Many mechanisms proposed to improve the accuracy of CNNs been given the name "attention", although it is becoming clear that accuracy alone does not explain how CNNs perform in real-world situations.
+Many mechanisms described as "attention" have been proposed to improve the accuracy of CNNs, although it is becoming clear that accuracy alone does not explain how CNNs perform in real-world situations.
 Attention has a long history of being studied by cognitive scientists and neuroscientists in the context of vision.
@@
 We adapt a classic form of visual search task so it can be performed by neural networks, and measure the accuracy of two CNN architectures, AlexNet and VGG16, as they report whether a target is present among distractors.
-Surprisingly, we find that both AlexNet and VGG16 show attention-like effects when performing the visual search task.
-When the target resembles the distractors, then a network's accuracy drops as the number of distractors increases.
+Surprisingly, we find that both AlexNet and VGG16 show attention-like effects when performing the visual search task: when the target resembles the distractors, then a network's accuracy drops as the number of distractors increases.
 This dependency of accuracy on the number of items is often referred to as "set size effects".
-We do not find that these set size effects can be explained by known factors to affect CNNs, such as edge effects.
+We do not find that these set size effects can be explained by factors known to affect CNNs, such as edge effects.
 We then adapt an analysis of visual search tasks based on signal detection theory to show that set size effects persist even when controlling for how discriminable the targets are from the distractors.
@@
 
+### A psychophysics analysis of visual search shows set size affects remain even when controlling for how discriminable targets are from distractors
+The Feature Integration Theory of visual search attributed set size effects to serial processing (Treisman).
+This conclusion made sense based on the result that reaction time increased when discriminating targets from distractors required considering more than one feature, e.g.
+both color and orientation.
+However, later work showed that reaction time effects disappear as targets and distractors become more easily discriminable along one feature dimension (Duncan paper).
+Later researchers, using methods from psychophysics and signal-detection theory, carefully controlled for the discriminability of targets and distractors (Palmer review).
+This work provided important alternative explanations for the results seen in visual search experiments: attentional effects can be explained in terms of completely parallel but noisy representations that are used by a decision rule (Palmer, Eckstein).
+This approach also explains errors in terms of low thresholds used by decision making criteria.
+We adapted these psychophysics methods to see if they offered additional insights into the attention-like effects we saw in CNNs.
+#### Psychometric curves
+First we generated 'feature' search stimuli--red vertical rectangle v.
+green vertical rectangle--where the red and green were more or less discriminable.
+We did so by taking advantage of RGB color space: by adding more red to the green color, and vice versa, the colors of the target and distractor can be moved closer to each other from the two red and green verticies of the triangle that defines RGB space.
+(When the color has "all red" and "all green", it becomes yellow, so the stimuli we generated moved closer and closer to yellow.) We then used these stimuli to measure the accuracy of the neural nets we trained as the target and distractor became more and more similar.
+<img src="../static/psychometric_function_fits.png" alt="efficient" style="width: 500px;"/>
+#### Threshold v. set size
+We then used the psychometric curves to find what level of target-distractor discriminability gave rise to 75% accuracy, known as the "difference threshold".
+We found this difference threshold for each trained network and each set size.
+<img src="../static/threshold_v_setsize.png" alt="efficient" style="width: 600px;"/>  
+Here we see that the threshold goes up with set size, meaning more discriminable target-distractor pairs are required as set size increases.
+In signal detection theory terms, this result is consistent with any theory of attention where decisions are made with a low threshold.
+
+For a detailed walkthrough of how we did this, please see [this notebook](./psychometric.ipynb).
+
 ## Methods

### 2019-03-03 docs/notebooks/results.Rmd
rewrite intro in notebooks/results.ipynb + .Rmd

@@
 ## Introduction
+
+### Convolutional neural networks and attentional mechanisms
+Neural networks have in the past few years transitioned from being an academic area of artificial intellgience to being powerful tools for engineers working in areas such as image classification, speech-to-text, and automatic translation, to name but a few.
+In the 1980s, artificial neural networks trained with backpropagation gave rise to [Connectionist](https://en.wikipedia.org/wiki/Connectionism) models of cognition.
+Similar neural network models lie at the heart of the recent rise of deep learning, in large part because of advances in methods for training networks, and advances in technology such as GPUs that allow for parallelized training.
+When benchmarked with scalar metrics like accuracy on computer vision tasks such as object recognition, these deep neural networks appear capable of human or super-human levels of skill.
+The first major gains in accuracy were seen on image classification tasks such as ImageNet <cite data-cite="6027200/WGZWMWX7"></cite>, and the deep neural networks that made these advances were built with what are known as convolutional layers.
+As the name implies, these layers perform a convolution-like operation by multiplying filters of learned weights with the input image, producing as outputs feature maps that can be further transformed by succeeding layers.
+Convolutional neural networks (CNNs) represent the state of the art in image recognition.
+Many mechanisms proposed to extend the abilities of CNNs have been given the name "attention", and these attention mechanisms typically involve some sort of recurrence [@baMultipleObjectRecognition2014; @mnihRecurrentModelsVisual2014; @xuShowAttendTell2015].
+These mechanisms are arguably analogous to what is called **overt** attention in psychology and neuroscience.
+In the context of vision, overt attention describes movement of the eyes or head so that the high-acuity fovea is placed on the focus of attention.
+In contrast, **covert** attention is any mechanism that occurs internally to select features relevant for a task while discarding irrelevant features.
+Covert attention can be considered one type of **dynamic routing** of information thought to occur in the brain.
+Models of dynamic routing, inspired by studies of visual attention [@bobierAttentionalRoutingCircuit2011; @bobierUnifyingMechanisticModel2014a; @olshausenMultiscaleDynamicRouting1995; @olshausenNeurobiologicalModelVisual1993] have also formed the basis of attention-like mechanisms in applied research on neural networks [@tangLearningGenerativeModels2014].
+The function of CNNs has not typically been considered within the context of this work on covert visual attention.
+
+### Convolutional neural networks as models of the visual system
+Recently, many in neuroscience and cognitive science have returned to using neural network models to understand the brain, in part because of their resurgence.
+More specifically, many researchers have drawn parallels between the architecture of CNNs and the architecture of the visual system in the brain <cite data-cite="6027200/8ISULWGF"></cite>, in particular in humans and other primates where this system has been most thoroughly studied.
+Like CNNs, the visual system has a hierarchical structure, and is thought to function in part by performing transformations at each level of this hierarchy so that high-dimensional, low-level features are mapped into low-dimensional abstract representations.
+Several recent studies in the realm of cognitive neuroscience have measured representations in more directly tested the similarities between CNNs and the visual system, by comparing representations in both <cite data-cite="6027200/6D4LKJD9"></cite><cite data-cite="6027200/U77Y74SI"></cite>.
+Representations can be easily extracted from CNNs because they are "transparent box" models--i.e.
+we have access to the "activity" of the individual units from which these models are built when we "show" the CNNs a stimulus.
+In the same studies, brain activity was obtained with electrophysiological or functional imaging methods.
+Such studies have found in general that representations are similar.
+(The comparison depends on representational similarity methods <cite data-cite="6027200/IQ2Y895S"></cite><cite data-cite="6027200/EQTU2P3Z"></cite>, but the concern of this study is not the validity of those methods.)
+
 ### Visual search tasks for investigating attention

### 2019-03-03 docs/notebooks/results.Rmd
fix up references in 'CNNs as visual system models' section of intro

@@
 When benchmarked with scalar metrics like accuracy on computer vision tasks such as object recognition, these deep neural networks appear capable of human or super-human levels of skill.
-The first major gains in accuracy were seen on image classification tasks such as ImageNet <cite data-cite="6027200/WGZWMWX7"></cite>, and the deep neural networks that made these advances were built with what are known as convolutional layers.
+The first major gains in accuracy were seen on image classification tasks such as ImageNet [@krizhevskyImagenetClassificationDeep2012; @simonyanVeryDeepConvolutional2014], and the deep neural networks that made these advances were built with what are known as convolutional layers.
 As the name implies, these layers perform a convolution-like operation by multiplying filters of learned weights with the input image, producing as outputs feature maps that can be further transformed by succeeding layers.
@@
 Recently, many in neuroscience and cognitive science have returned to using neural network models to understand the brain, in part because of their resurgence.
-More specifically, many researchers have drawn parallels between the architecture of CNNs and the architecture of the visual system in the brain <cite data-cite="6027200/8ISULWGF"></cite>, in particular in humans and other primates where this system has been most thoroughly studied.
+More specifically, many researchers have drawn parallels between the architecture of CNNs and the architecture of the visual system in the brain [@kriegeskorteDeepNeuralNetworks2015], in particular in humans and other primates where this system has been most thoroughly studied.
 Like CNNs, the visual system has a hierarchical structure, and is thought to function in part by performing transformations at each level of this hierarchy so that high-dimensional, low-level features are mapped into low-dimensional abstract representations.
-Several recent studies in the realm of cognitive neuroscience have measured representations in more directly tested the similarities between CNNs and the visual system, by comparing representations in both <cite data-cite="6027200/6D4LKJD9"></cite><cite data-cite="6027200/U77Y74SI"></cite>.
+Several recent studies in the realm of cognitive neuroscience have measured representations in more directly tested the similarities between CNNs and the visual system, by comparing representations in both [@khaligh-razaviDeepSupervisedNot2014; @yaminsPerformanceoptimizedHierarchicalModels2014; @yaminsUsingGoaldrivenDeep2016].
 Representations can be easily extracted from CNNs because they are "transparent box" models--i.e.
@@
 Such studies have found in general that representations are similar.
-(The comparison depends on representational similarity methods <cite data-cite="6027200/IQ2Y895S"></cite><cite data-cite="6027200/EQTU2P3Z"></cite>, but the concern of this study is not the validity of those methods.)
+(Of course this comparison depends on methods for measuring representational similarity [@bobadilla-suarezMeasuresNeuralSimilarity2018; @diedrichsenRepresentationalModelsCommon2017], but this study does not address those methods.)
 
@@
 To generate visual search stimuli that could be fed into a neural network, we used a simple Python package (<https://github.com/NickleDave/searchstims>) built on top of the [PyGame](https://www.pygame.org) library .
-The search displays were divided up into a 5x5 grid of cells, and items were plotted in a randomly chosen subset of cells (as described in <cite data-cite="6027200/QCQFYFDZ"></cite> and shown at <http://search.bwh.harvard.edu/new/data_set_files.html>).
+The search displays were divided up into a 5x5 grid of cells, and items were plotted in a randomly chosen subset of cells (as described in [@wolfeReactionTimeDistributions2010] and shown at <http://search.bwh.harvard.edu/new/data_set_files.html>).
 The library applied jitter to the items, so that two images could be different even if both images had the same set size and target present condition, and the items were in the exact same cells.

### 2019-03-03 docs/notebooks/results.Rmd
write overly verbose linker/topic sentence for 'visual search' in intro

@@
 ### Visual search tasks for investigating attention
+In spite of the many attentional mechanisms proposed for neural networks, and the many parallels and links between studies of vision in psychology, neuroscience, and artificial intelligence, little work has been done to relate the tasks and models used to investigate visual attention to neural networks.
+Here we study the whether visual search tasks used to investigate attention and the different models of visual attention in the brain provide alternate ways of understanding existing CNNs, comparing their output with human behavior, and benchmarking new attentional mechansims.
+We begin with background on visual search tasks.
 In the "classical" form of the visual search paradigm [@wolfeVisualSearch1998], experimenters present a subject with a search display on which there are discrete items, i.e.

### 2019-03-03 docs/notebooks/results.Rmd
better explain results in VGG16 section

@@
 To test whether these effects resulted from the architecture of AlexNet or were more general, we repeated the simulations using the VGG16 architecture.
+We again saw a clear difference in set size effects when comparing accuracy of networks performing either feature search or spatial conjunction search.
 
@@
 
-We again saw this effect became stronger the more that the VGG16 architecture was trained.
+We also again saw that the set size effect increased, as did difference in accuracy v.
+set size slopes for feature and spatial conjunction search, as the VGG16 architecture was trained for more epochs.
 

### 2019-03-03 docs/notebooks/results.Rmd
fix refs in psychometric results section

@@
 ### A psychophysics analysis of visual search shows set size affects remain even when controlling for how discriminable targets are from distractors
-The Feature Integration Theory of visual search attributed set size effects to serial processing (Treisman).
-This conclusion made sense based on the result that reaction time increased when discriminating targets from distractors required considering more than one feature, e.g.
+The Feature Integration Theory of visual search attributed set size effects to **serial** processing required to bind different types of features [@treismanFeatureintegrationAttention1980].
+This conclusion made sense, based on the result that reaction time increased when discriminating targets from distractors required subjects to consider more than one feature, e.g.
 both color and orientation.
-However, later work showed that reaction time effects disappear as targets and distractors become more easily discriminable along one feature dimension (Duncan paper).
-Later researchers, using methods from psychophysics and signal-detection theory, carefully controlled for the discriminability of targets and distractors (Palmer review).
-This work provided important alternative explanations for the results seen in visual search experiments: attentional effects can be explained in terms of completely parallel but noisy representations that are used by a decision rule (Palmer, Eckstein).
+However, later work showed that reaction time effects disappear as targets and distractors become more easily discriminable along one feature dimension [@duncanVisualSearchStimulus1989].
+Later researchers, using methods from psychophysics and signal-detection theory, carefully controlled for the discriminability of targets and distractors [@palmerPsychophysicsVisualSearch2000].
+This work provided important alternative explanations for the results seen in visual search experiments: attentional effects can be explained in terms of completely parallel but noisy representations that are used by a decision rule [@ecksteinLowerVisualSearch1998; @ecksteinSignalDetectionModel2000; @palmerMeasuringEffectAttention1993; @palmerSetsizeEffectsVisual1994].
 This approach also explains errors in terms of low thresholds used by decision making criteria.
 We adapted these psychophysics methods to see if they offered additional insights into the attention-like effects we saw in CNNs.
+
 #### Psychometric curves
@@
 <img src="../static/threshold_v_setsize.png" alt="efficient" style="width: 600px;"/>  
-Here we see that the threshold goes up with set size, meaning more discriminable target-distractor pairs are required as set size increases.
+Here we see that the difference threshold goes up with set size, meaning more discriminable target-distractor pairs are required as set size increases.
 In signal detection theory terms, this result is consistent with any theory of attention where decisions are made with a low threshold.
-
-For a detailed walkthrough of how we did this, please see [this notebook](./psychometric.ipynb).
+(For a detailed walkthrough of how we did this, please see [this notebook](./psychometric.ipynb).)
 

### 2019-03-03 docs/notebooks/results.Rmd
polish up 'vis search' section in intro, add ref to Põder

@@
 In spite of the many attentional mechanisms proposed for neural networks, and the many parallels and links between studies of vision in psychology, neuroscience, and artificial intelligence, little work has been done to relate the tasks and models used to investigate visual attention to neural networks.
-Here we study the whether visual search tasks used to investigate attention and the different models of visual attention in the brain provide alternate ways of understanding existing CNNs, comparing their output with human behavior, and benchmarking new attentional mechansims.
+We are aware of only one study [@poderCapacityLimitationsVisual2017], which led us to the results that we report below as we replicated it.
+Like that study, we measured the performance of an exist CNN on a visual search task.
+In addition, we test whether our results hold for other CNN architectures, and we apply analyses from signal detection theory-based models of attention to understand our results.
+Our ultimate goal is to provide new methods for comparing neural network output with human behavior and of benchmarking proposed attentional mechansims.
+First we describe visual search tasks, and then we describe different types of cognitive models used to explain results obtained with these tasks.
 We begin with background on visual search tasks.

### 2019-03-07 docs/notebooks/results.Rmd
update methods in results.ipynb / results.Rmd

@@
 ## Methods
-Alexnet training:  
-• based on Poder 2017  
-https://arxiv.org/pdf/1707.09775.pdf
-• Stochastic gradient descent  
-• Learning rate for fully-connected layers: 1e-20  
-• 10 epochs (as in Poder 2017), and 400 epochs
+Fine-tuning Alexnet and VGG16:
+* based on Põder 2017: https://arxiv.org/pdf/1707.09775.pdf
+* Stochastic gradient descent
+* Base learning rate: 1e-20
+* Rate for fully-connected layers: 0.0001
+* epochs: [10 (as in Poder 2017), 50, 200, 400, 800]
 

### 2019-03-15 docs/notebooks/results.Rmd
update acknowledgments in results
update results.Rmd

@@
 ## Acknowledgments
+- Research funded by the Lifelong Learning Machines program, DARPA/Microsystems Technology Office,  DARPA cooperative agreement HR0011-18-2-0019
+- David Nicholson was partially supported by the 2017 William K. and Katherine W. Estes Fund to F. Pestilli, R. Goldstone and L. Smith, Indiana University Bloomington.

### 2019-03-15 docs/notebooks/results.Rmd
update results.Rmd

--- 2019-03-03 docs/notebooks/results.Rmd
@@
 In spite of the many attentional mechanisms proposed for neural networks, and the many parallels and links between studies of vision in psychology, neuroscience, and artificial intelligence, little work has been done to relate the tasks and models used to investigate visual attention to neural networks.
-We are aware of only one study [@poderCapacityLimitationsVisual2017], which led us to the results that we report below as we replicated it.
-Like that study, we measured the performance of an exist CNN on a visual search task.
-In addition, we test whether our results hold for other CNN architectures, and we apply analyses from signal detection theory-based models of attention to understand our results.
-Our ultimate goal is to provide new methods for comparing neural network output with human behavior and of benchmarking proposed attentional mechansims.
-First we describe visual search tasks, and then we describe different types of cognitive models used to explain results obtained with these tasks.
+Here we study the whether visual search tasks used to investigate attention and the different models of visual attention in the brain provide alternate ways of understanding existing CNNs, comparing their output with human behavior, and benchmarking new attentional mechansims.
 We begin with background on visual search tasks.
@@
 ### A psychophysics analysis of visual search shows set size affects remain even when controlling for how discriminable targets are from distractors
-The Feature Integration Theory of visual search attributed set size effects to **serial** processing required to bind different types of features [@treismanFeatureintegrationAttention1980].
-This conclusion made sense, based on the result that reaction time increased when discriminating targets from distractors required subjects to consider more than one feature, e.g.
+The Feature Integration Theory of visual search attributed set size effects to serial processing (Treisman).
+This conclusion made sense based on the result that reaction time increased when discriminating targets from distractors required considering more than one feature, e.g.
 both color and orientation.
-However, later work showed that reaction time effects disappear as targets and distractors become more easily discriminable along one feature dimension [@duncanVisualSearchStimulus1989].
-Later researchers, using methods from psychophysics and signal-detection theory, carefully controlled for the discriminability of targets and distractors [@palmerPsychophysicsVisualSearch2000].
-This work provided important alternative explanations for the results seen in visual search experiments: attentional effects can be explained in terms of completely parallel but noisy representations that are used by a decision rule [@ecksteinLowerVisualSearch1998; @ecksteinSignalDetectionModel2000; @palmerMeasuringEffectAttention1993; @palmerSetsizeEffectsVisual1994].
+However, later work showed that reaction time effects disappear as targets and distractors become more easily discriminable along one feature dimension (Duncan paper).
+Later researchers, using methods from psychophysics and signal-detection theory, carefully controlled for the discriminability of targets and distractors (Palmer review).
+This work provided important alternative explanations for the results seen in visual search experiments: attentional effects can be explained in terms of completely parallel but noisy representations that are used by a decision rule (Palmer, Eckstein).
 This approach also explains errors in terms of low thresholds used by decision making criteria.
 We adapted these psychophysics methods to see if they offered additional insights into the attention-like effects we saw in CNNs.
-
 #### Psychometric curves
@@
 <img src="../static/threshold_v_setsize.png" alt="efficient" style="width: 600px;"/>  
-Here we see that the difference threshold goes up with set size, meaning more discriminable target-distractor pairs are required as set size increases.
+Here we see that the threshold goes up with set size, meaning more discriminable target-distractor pairs are required as set size increases.
 In signal detection theory terms, this result is consistent with any theory of attention where decisions are made with a low threshold.
-(For a detailed walkthrough of how we did this, please see [this notebook](./psychometric.ipynb).)
+
+For a detailed walkthrough of how we did this, please see [this notebook](./psychometric.ipynb).
 

### 2019-05-27 src/bin/searchstims_finetune_script.py
update finetune configs

@@ -43,14 +43,14 @@ for window_size in ALEXNET_SIZE, VGG16_SIZE:
         vgg16_zip = zip(keys, vals)
 
 OUTPUT_DIR = Path('data/visual_search_stimuli')
-TARGET_PRESENT = 20000
-TARGET_ABSENT = 20000
-SET_SIZES = [3, 6, 12, 18]
+TARGET_PRESENT = 12100
+TARGET_ABSENT = 12100
+SET_SIZES = [1, 2, 4, 8]
 
 
 def main():
     for cnn, zipped in zip(
-        ['alexnet', 'vgg16'],
+        ['alexnet', 'VGG16'],
             [alexnet_zip, vgg16_zip]):
         for key, val in zipped:
             json_filename = f'{cnn}_{key}.json'

### 2019-05-27 data/runall_VGG16_finetune.sh
update runall_VGG16_finetune.sh

@@ -1,6 +1,6 @@
 #!/usr/bin/env bash
 
 # prepare data sets for training neural nets, then train, then test accuracy of trained nets
-searchnets all ./configs/searchnets_vgg16_finetune_RVvGV.ini
-searchnets all ./configs/searchnets_vgg16_finetune_RVvRHGV.ini
-searchnets all ./configs/searchnets_vgg16_finetune_2_v_5.ini
+searchnets all ./configs/searchnets_VGG16_finetune_RVvGV.ini
+searchnets all ./configs/searchnets_VGG16_finetune_RVvRHGV.ini
+searchnets all ./configs/searchnets_VGG16_finetune_2_v_5.ini

### 2019-05-27 src/searchnets/data.py
remove extra newline at end of searchnets/data.py

@@ -307,4 +307,3 @@ def data(train_dir,
                      stim_type_vec_test=stim_type_vec_test,
                      )
     joblib.dump(data_dict, gz_filename)
-

### 2019-05-27 src/searchnets/nets/alexnet.py
remove two extra newlines at end of searchnets/nets/alexnet.py

@@ -164,5 +164,3 @@ class AlexNet:
         fc8 = self.fc(dropout7, 4096, self.num_classes, relu=False, name='fc8')
 
         self.output = fc8
-
-

### 2019-05-27 data/runall_VGG16_learncurve_2_v_5.sh
add runall bash scripts for alexenet and VGG16 learning curves

@@ -1,0 +1,5 @@
+#!/usr/bin/env bash
+
+# prepare data sets for training neural nets, then train, then test accuracy of trained nets
+searchnets data ./configs/searchnets_VGG16_learncurve_2_v_5.ini
+searchnets learncurve ./configs/searchnets_VGG16_learncurve_2_v_5.ini

### 2019-05-27 data/runall_alexnet_learncurve_2_v_5.sh

@@ -1,0 +1,5 @@
+#!/usr/bin/env bash
+
+# prepare data sets for training neural nets, then train, then test accuracy of trained nets
+searchnets data ./configs/searchnets_alexnet_learncurve_2_v_5.ini
+searchnets learncurve ./configs/searchnets_alexnet_learncurve_2_v_5.ini

### 2019-05-27 src/searchnets/learncurve.py
fix how learncurve.py determines set sizes in training data

@@ -26,8 +26,8 @@ def learncurve(gz_filename,
                test_results_save_path,
                dropout_rate=0.5,
                val_size=None):
-    """train convolutional neural networks to perform visual search task.
-    A fine tuning approach is used, as in Poder 2017 (https://arxiv.org/pdf/1707.09775.pdf)
+    """generate a learning curve, training convolutional neural networks
+    to perform visual search task with a range of sizes of training set.
 
     Parameters
     ----------
@@ -220,6 +220,17 @@ def learncurve(gz_filename,
     # -------------------- now measure accuracy on test set ------------------------------------------------------------
     x_test = data_dict['x_test']
     y_test = data_dict['y_test']
+    # make sure there's only one 'set' of set sizes
+    set_sizes_by_stim_type = data_dict['set_sizes_by_stim_stype']
+    set_sizes = []
+    for stim_type, set_sizes_this_stim in set_sizes_by_stim_type.items():
+        if set_sizes == []:
+            set_sizes = set_sizes_this_stim
+        else:
+            if set_sizes_this_stim != set_sizes:
+                raise ValueError('set sizes are not the same across visual search stimuli')
+            else:
+                continue
 
     for train_size in train_size_list:
         for epochs in epochs_list:
@@ -270,7 +281,7 @@ def learncurve(gz_filename,
                         y_pred = sess.run(predictions['labels'], feed_dict=feed)
                         y_pred_all.append(y_pred)
                     y_pred_all = np.concatenate(y_pred_all)
-                    set_sizes = data_dict['set_sizes']
+
                     set_size_vec_test = data_dict['set_size_vec_test']
                     acc_per_set_size = []
                     for set_size in set_sizes:

### 2019-05-27 src/searchnets/train.py
remove link to Poder 2017 in train.py docstring
since that is not the only approach being used anymore

@@ -70,7 +70,6 @@ def train(gz_filename,
           val_size=None,
           save_acc_by_set_size_by_epoch=False):
     """train convolutional neural networks to perform visual search task.
-    A fine tuning approach is used, as in Poder 2017 (https://arxiv.org/pdf/1707.09775.pdf)
 
     Parameters
     ----------

### 2019-05-27 src/searchnets/test.py
remove unused import of ast in test.py

@@ -1,4 +1,3 @@
-import ast
 import os
 
 import numpy as np

### 2019-05-27 src/searchnets/learncurve.py
learncurve feeds validation set in batches to avoid OOM

@@ -84,12 +84,11 @@ def learncurve(gz_filename,
     data_dict = joblib.load(gz_filename)
     x_train = data_dict['x_train']
     y_train = data_dict['y_train']
-    training_set = [x_train, y_train]
     if val_size:
-        val_set = [data_dict['x_val'],
-                   data_dict['y_val']]
+        x_val = data_dict['x_val']
+        y_val = data_dict['y_val']
     else:
-        val_set = None
+        x_val, y_val = None, None
 
     if type(epochs_list) is int:
         epochs_list = [epochs_list]
@@ -198,10 +197,23 @@ def learncurve(gz_filename,
                         training_loss.append(avg_loss / (i + 1))
                         print('Epoch %02d Training Avg. Loss: %7.3f' % (
                             epoch, avg_loss), end=' ')
-                        if val_set is not None:
-                            feed = {x: val_set[0],
-                                    y: val_set[1]}
-                            valid_acc = sess.run(accuracy, feed_dict=feed)
+                        if x_val is not None:
+                            batch_gen = batch_generator(x_val, y_val,
+                                                        batch_size=batch_size,
+                                                        shuffle=False)
+                            total = int(np.ceil(x_val.shape[0] / batch_size))
+                            pbar = tqdm(enumerate(batch_gen), total=total)
+
+                            valid_acc = []
+                            for i, (batch_x, batch_y) in pbar:
+                                pbar.set_description(f'batch {i} of {total}')
+                                feed = {x: batch_x,
+                                        y: batch_y,
+                                        rate: dropout_rate}
+
+                                valid_acc.append(sess.run(accuracy, feed_dict=feed))
+                            valid_acc = np.asarray(valid_acc).mean()
+
                             print(' Validation Acc: %7.3f' % valid_acc)
                         else:
                             print()

### 2019-05-27 src/searchnets/train.py
train feeds validation set in batches to avoid OOM error

@@ -264,10 +264,24 @@ def train(gz_filename,
 
                     training_loss.append(avg_loss / (i + 1))
                     print(f'Epoch {epoch + 1}, Training Avg. Loss: {avg_loss:7.3f}')
+
                     if val_set is not None:
-                        feed = {x: val_set[0],
-                                y: val_set[1]}
-                        valid_acc = sess.run(accuracy, feed_dict=feed)
+                        batch_gen = batch_generator(val_set[0], val_set[1],
+                                                    batch_size=batch_size,
+                                                    shuffle=False)
+                        total = int(np.ceil(val_set[0].shape[0] / batch_size))
+                        pbar = tqdm(enumerate(batch_gen), total=total)
+
+                        valid_acc = []
+                        for i, (batch_x, batch_y) in pbar:
+                            pbar.set_description(f'batch {i} of {total}')
+                            feed = {x: batch_x,
+                                    y: batch_y,
+                                    rate: dropout_rate}
+
+                            valid_acc.append(sess.run(accuracy, feed_dict=feed))
+                        valid_acc = np.asarray(valid_acc).mean()
+
                         print(' Validation Acc: %7.3f' % valid_acc)
                     else:
                         print()

### 2019-05-27 src/searchnets/plot/figures.py
add functionality to plot.figures.acc_v_set_size
- can pass axis as argument
  + if not passed, make a figure with a single axis instead
- can pass a bunch of boolean args that control plot behavior
  + set_xlabel, set_ylabel, set_ylims, add_legend, plot_mean
- can pass 'task name' which gets added to label used for legend

@@ -20,7 +20,11 @@ plt.rcParams['legend.fontsize'] = 16
 plt.rcParams['figure.titlesize'] = 20
 
 
-def acc_v_set_size(results, set_sizes=(1, 2, 4, 8), title=None, save_as=None, figsize=(10, 5)):
+def acc_v_set_size(results, set_sizes=(1, 2, 4, 8), ax=None,
+                   title=None, save_as=None, figsize=(10, 5),
+                   set_xlabel=False, set_ylabel=False, set_ylim=True,
+                   ylim=(0, 1.1), plot_mean=True, add_legend=False,
+                   task_name=None):
     """plot accuracy as a function of visual search task set size
     for models trained on a single task or dataset
 
@@ -31,13 +35,31 @@ def acc_v_set_size(results, set_sizes=(1, 2, 4, 8), title=None, save_as=None, fi
         on test set of visual search stimuli
     set_sizes : list
         of int, set sizes of visual search stimuli. Default is [1, 2, 4, 8].
+    ax : matplotlib.Axis
+        axis on which to plot figure. Default is None, in which case a new figure with
+        a single axis is created for the plot.
     title : str
         string to use as title of figure. Default is None.
     save_as : str
         path to directory where figure should be saved. Default is None, in which
         case figure is not saved.
     figsize : tuple
-        (width, height) in inches. Default is (10, 5).
+        (width, height) in inches. Default is (10, 5). Only used if ax is None and a new
+        figure is created.
+    set_xlabel : bool
+        if True, set the value of xlabel to "set size". Default is False.
+    set_ylabel : bool
+        if True, set the value of ylabel to "accuracy". Default is False.
+    set_ylim : bool
+        if True, set the y-axis limits to the value of ylim.
+    ylim : tuple
+        with two elements, limits for y-axis. Default is (0, 1.1).
+    plot_mean : bool
+        if True, find mean accuracy and plot as a separate solid line. Default is True.
+    add_legend : bool
+        if True, add legend to axis. Default is False.
+    task_name : str
+
 
     Returns
     -------
@@ -46,16 +68,32 @@ def acc_v_set_size(results, set_sizes=(1, 2, 4, 8), title=None, save_as=None, fi
     accs = joblib.load(results)['acc_per_set_size_per_model']
[…]

### 2019-05-27 docs/ccneuro2019/paper.tex
add docs/ccneuro2019
with apacite.bst and ccn.sty as downloaded from ccneuro site,
plus initial draft of paper.tex

% 
% Annual CCN conference
% Sample LaTeX Two-Page Summary -- Proceedings Format
% based on the prior cognitive science style file

% Original : Ashwin Ram (ashwin@cc.gatech.edu)       04/01/1994
% Modified : Johanna Moore (jmoore@cs.pitt.edu)      03/17/1995
% Modified : David Noelle (noelle@ucsd.edu)          03/15/1996
% Modified : Pat Langley (langley@cs.stanford.edu)   01/26/1997
% Latex2e corrections by Ramin Charles Nakisa        01/28/1997 
% Modified : Tina Eliassi-Rad (eliassi@cs.wisc.edu)  01/31/1998
% Modified : Trisha Yannuzzi (trisha@ircs.upenn.edu) 12/28/1999 (in process)
% Modified : Mary Ellen Foster (M.E.Foster@ed.ac.uk) 12/11/2000
% Modified : Ken Forbus                              01/23/2004
% Modified : Eli M. Silk (esilk@pitt.edu)            05/24/2005
% Modified : Niels Taatgen (taatgen@cmu.edu)        10/24/2006
% Modified : David Noelle (dnoelle@ucmerced.edu)     11/19/2014
% Modified : Konrad Kording (koerding@gmail.com) 2/15/2017

\documentclass[10pt,letterpaper]{article}

\usepackage{ccn}
\usepackage{pslatex}
\usepackage{apacite}
\usepackage{graphicx}

\graphicspath{ {./images/} }

\title{What is attention? A study of neural networks performing a visual search task}

\author{{\large \bf David Nicholson (dnicho4@emory.edu)} \\
  A Department, 1234 Example Street\\
A City, State 12345 A country
  \AND {\large \bf Astrid Prinz (AnotherPerson@this.planet.edu)} \\
  A Department, 1234 Example Street\\
A City, State 12345 A country}

\begin{document}

\maketitle

\section{Abstract}
{
\bf
In machine learning and artificial intelligence, attention mechanisms are proposed to improve classification accuracy. In cognitive science, however, attention mechanisms are often defined in terms of experimental manipulations that impair performance. For example, psychologists have long used a now classic form of visual search task to determine what factors increase reaction times or decrease accuracy. Here we investigate 
the ability of neural networks to perform the classic visual search task, with the goal of reconciling what appear to be different definitions of 
attention in cognitive science and in artificial intelligence. We first show that convolutional neural networks (CNNs) trained for image classification 
can show human-like attentional effects, replicating and extending a previous study. Then we demonstrate that, by changing how we train the networks 
and the statistics of the training data, these effects can largely be eliminated, i.e., that CNNs are capable of performing the task without explicit 
attention mechanisms. We go on to show that manipulating the discriminability of the target and distractors in the task can produce attentional effects 
in spite of the changes we made to training. We discuss the implications of these findings for the study of attention in neuroscience and AI.
}
\begin{quote}
\small
\textbf{Keywords:} 
attention;visual search
\end{quote}

\section{Introduction}
\subsection{Attention}
Several attention mechanisms are currently used in machine learning 
and artificial intelligence to improve classification accuracy and similar 
metrics. Commonly used attention mechanisms consist essentially of learned 
weightings, with some additionally using recurrence.
This focus on attention as a mechanism to improve accuracy or other metrics 
stands in contrast to studies of attention in psychology and neuroscience, 
where typically it is experimental manipulations that impair behavior which 
produce the results on which theories of the mechanisms of attention are built.

\subsection{visual search}
\textbf{set size effects}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\columnwidth]{fig1}
\end{center}
\caption{A classic form of visual search task.} 
\label{sample-figure}
\end{figure}

\section{Results}

\subsection{CNNs trained for image classification show set size effects}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\columnwidth]{fig2}
\end{center}
\caption{CNNs trained for image classification show set size effects.} 
\label{sample-figure}
\end{figure}

\subsection{Changing hyperparameters and balancing the dataset largely 
eliminates set size effects}

Because we defined set size with effects in terms of accuracy, it could 
be the case that our results do not arise because of a similarity between 
how CNNs and brains process images. Instead they could be an artifact of 
how we trained networks. As discussed below, we see this as a general issue 
for studies that compare cognitive science and artificial intelligence, and 
so here we report results that might otherwise be considered methodological 
troubleshooting.
To determine whether our results were an artifact of training, we considered aspects of 
training that can impair accuracy: the amount of training data, the hyperparameters 
used to train the network, and the statistics of the dataset. As shown in 
figure two, we found that all three factors contributed to the effect we saw.
We generated learning curves where we plotted accuracy on the training and 
test set as a function of the training set size. These learning curves revealed 
that the original size we chose for the training set was far from where 
accuracy on training and test set converged. This result indicates that the 
network can improve accuracy with increased training data. To gain insight 
about how accuracy depended on the learning rate, a key hyperparameter, and 
the balance of the data set, we plotted accuracy on the training set 
\emph{separately for each set size in the visual search stimuli}. As shown in 
figure 2b, these plots revealed (1) that accuracy had not yet reached some asymptotic 
value, and (2) that there was an inverse relationship between the set size of 
a visual search stimulus and the rate that its accuracy increased.
To test whether we could improve the learning rate, we used random search, and 
did find we were able to improve accuracy and decrease training time by 
abandoning the fine-tuning approach and instead using a typical learning rate 
on the fully-connected layers (and simply freezing the pre-trained weights in other layers).

Use standard APA citation format. Citations within the text should
include the author's last name and year. If the authors' names are
included in the sentence, place only the year in parentheses, as in
\citeA{NewellSimon1972a}, but otherwise place the entire reference in
parentheses with the authors and year separated by a comma
\cite{NewellSimon1972a}. List multiple references alphabetically and
separate them by semicolons
\cite{ChalnickBillman1988a,NewellSimon1972a}. Use the
``et~al.'' construction only after listing all the authors to a
publication in an earlier reference and for citations with four or
more authors.

\section{Discussion}

\subsection{Implications for comparative studies of neuroscience and 
artificial intelligence}
- problem for any

- e.g. Bengio study of gestalt fx

\subsection{Implications for the study of attention in the brain}
- if the task is possible to perform perfectly, who needs attention?
- brain optimized for something else besides classifying static images
- in this sense, consistent with findings that recurrent

\subsection{Footnotes}

Indicate footnotes with a number\footnote{Sample of the first
footnote.} in the text. Place the footnotes in 9~point type at the
bottom of the column on which they appear. Precede the footnote block
with a horizontal rule.\footnote{Sample of the second footnote.}

\subsection{Tables}

Number tables consecutively. Place the table number and title (in
10~point) above the table with one line space above the caption and
one line space below it, as in Table~\ref{sample-table}. You may float
tables to the top or bottom of a column, or set wide tables across
both columns.

\begin{table}[!ht]
\begin{center} 
\caption{Sample table title.} 
\label{sample-table} 
\vskip 0.12in
\begin{tabular}{ll} 
\hline
Error type    &  Example \\
\hline
Take smaller        &   63 - 44 = 21 \\
Always borrow~~~~   &   96 - 42 = 34 \\
0 - N = N           &   70 - 47 = 37 \\
0 - N = 0           &   70 - 47 = 30 \\
\hline
\end{tabular} 
\end{center} 
\end{table}

\subsection{Figures}

Make sure that the artwork can be printed well (e.g. dark colors) and that 
the figures make understanding the paper easy.
 Number figures sequentially, placing the figure
number and caption, in 10~point, after the figure with one line space
above the caption and one line space below it, as in
Figure~\ref{sample-figure}. If necessary, leave extra white space at
the bottom of the page to avoid splitting the figure and figure
caption. You may float figures to the top or bottom of a column, or
set wide figures across both columns.

\begin{figure}[ht]
\begin{center}
\includegraphics{fig1}
\end{center}
\caption{This is a figure.} 
\label{sample-figure}
\end{figure}

\section{Acknowledgments}

Place acknowledgments (including funding information) in a section at
the end of the paper.

\section{References Instructions}

Follow the APA Publication Manual for citation format, both within the
text and in the reference list, with the following exceptions: (a) do
not cite the page numbers of any book, including chapters in edited
volumes; (b) use the same format for unpublished references as for
published ones. Alphabetize references by the surnames of the authors,
with single author entries preceding multiple author entries. Order
references by the same authors by the year of publication, with the
earliest first.

Use a first level section heading, ``{\bf References}'', as shown
below. Use a hanging indent style, with the first line of the
reference flush against the left margin and subsequent lines indented
by 1/8~inch. Below are example references for a conference paper, book
chapter, journal article, dissertation, book, technical report, and
edited volume, respectively.

\nocite{ChalnickBillman1988a}
\nocite{Feigenbaum1963a}
\nocite{Hill1983a}
\nocite{OhlssonLangley1985a}
% \nocite{Lewis1978a}
\nocite{Matlock2001}
\nocite{NewellSimon1972a}
\nocite{ShragerLangley1990a}

\bibliographystyle{apacite}

\setlength{\bibleftmargin}{.125in}
\setlength{\bibindent}{-\bibleftmargin}

\bibliography{ccn_style}

\end{document}

### 2019-05-27 docs/ccneuro2019/paper.tex
fix graphicspath and image names in paper.tex
add linebreaks in docs/ccneuro2019/paper.tex abstract
to be consistent with 90 char margin width
change paper title

@@
 
-\graphicspath{ {./images/} }
+\graphicspath{ {../static/} }
 
-\title{What is attention? A study of neural networks performing a visual search task}
+\title{Do convolutional neural networks exhibit human-like attentional effects? 
+Conflicting results from a visual search task}
 
@@
 \bf
-In machine learning and artificial intelligence, attention mechanisms are proposed to improve classification accuracy.
-In cognitive science, however, attention mechanisms are often defined in terms of experimental manipulations that impair performance.
-For example, psychologists have long used a now classic form of visual search task to determine what factors increase reaction times or decrease accuracy.
-Here we investigate
-the ability of neural networks to perform the classic visual search task, with the goal of reconciling what appear to be different definitions of 
-attention in cognitive science and in artificial intelligence. We first show that convolutional neural networks (CNNs) trained for image classification 
-can show human-like attentional effects, replicating and extending a previous study. Then we demonstrate that, by changing how we train the networks 
-and the statistics of the training data, these effects can largely be eliminated, i.e., that CNNs are capable of performing the task without explicit 
-attention mechanisms. We go on to show that manipulating the discriminability of the target and distractors in the task can produce attentional effects 
-in spite of the changes we made to training. We discuss the implications of these findings for the study of attention in neuroscience and AI.
+In machine learning and artificial intelligence, attention mechanisms are proposed to 
+improve classification accuracy. In cognitive science, however, attention mechanisms are 
+often defined in terms of experimental manipulations that impair performance. For example,
+psychologists have long used a now classic form of visual search task to determine what 
+factors increase reaction times or decrease accuracy. Here we investigate the ability of 
+neural networks to perform the classic visual search task, with the goal of reconciling 
+what appear to be different definitions of attention in cognitive science and in 
+artificial intelligence. We first show that convolutional neural networks (CNNs) trained 
+for image classification can show human-like attentional effects, replicating and 
+extending a previous study. Then we demonstrate that, by changing how we train the 
+networks and the statistics of the training data, these effects can largely be 
+eliminated, i.e., that CNNs are capable of performing the task without explicit attention 
+mechanisms. We go on to show that manipulating the discriminability of the target and 
+distractors in the task can produce attentional effects in spite of the changes we made to
+training. We discuss the implications of these findings for the study of attention.
 }
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{fig1}
+\includegraphics[width=\columnwidth]{fig1/fig1.eps}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{fig2}
+\includegraphics[width=\columnwidth]{fig2/fig2.png}
 \end{center}

### 2019-05-27 docs/ccneuro2019/paper.tex
rewrite ccneuro paper abstract

@@
 \bf
-In machine learning and artificial intelligence, attention mechanisms are proposed to 
-improve classification accuracy. In cognitive science, however, attention mechanisms are 
-often defined in terms of experimental manipulations that impair performance. For example,
-psychologists have long used a now classic form of visual search task to determine what 
-factors increase reaction times or decrease accuracy. Here we investigate the ability of 
-neural networks to perform the classic visual search task, with the goal of reconciling 
-what appear to be different definitions of attention in cognitive science and in 
-artificial intelligence. We first show that convolutional neural networks (CNNs) trained 
-for image classification can show human-like attentional effects, replicating and 
-extending a previous study. Then we demonstrate that, by changing how we train the 
-networks and the statistics of the training data, these effects can largely be 
-eliminated, i.e., that CNNs are capable of performing the task without explicit attention 
-mechanisms. We go on to show that manipulating the discriminability of the target and 
-distractors in the task can produce attentional effects in spite of the changes we made to
-training. We discuss the implications of these findings for the study of attention.
+Selective attention mechanisms preferentially route some sensory inputs for further 
+processing, while necessarily discarding others. Researchers studying visual
+attention have often divided selective attention into overt and covert mechanisms. 
+Overt mechanisms involve orienting the eyes and/or head, while covert mechanisms 
+occur "centrally", in the brain, without overt orientation. Cognitive scientists have long
+used a new classic form of visual search task to investigate theoretical mechanisms of 
+selective covert visual attention. Here we build on previous studies that examine the 
+parallels between the visual system and convolutional neural networks (CNNs), by testing 
+whether CNNs exhibit human-like attentional effects while performing this visual search 
+task. Following a typical approach, we first show that CNNs trained 
+for image classification can show human-like attentional effects. Then we demonstrate 
+that, by changing how we train the networks and the statistics of the training data, 
+these effects can largely be eliminated, i.e., that CNNs are capable of performing the 
+task without explicit attention mechanisms. We go on to show that manipulating the 
+discriminability of the target and distractors in the task can produce attentional 
+effects in spite of the changes we made to training. Based on these results, we conclude 
+that CNNs do not show attentional effects.
 }

### 2019-05-27 docs/ccneuro2019/paper.tex
start to flesh out ccneuro paper introduction

@@
 and artificial intelligence to improve classification accuracy and similar 
-metrics. Commonly used attention mechanisms consist essentially of learned 
-weightings, with some additionally using recurrence.
-This focus on attention as a mechanism to improve accuracy or other metrics 
-stands in contrast to studies of attention in psychology and neuroscience, 
-where typically it is experimental manipulations that impair behavior which 
-produce the results on which theories of the mechanisms of attention are built.
+metrics. Commonly used attention mechanisms consist essentially of weightings 
+that are learned by updating the weights so as to minimize a loss function, 
+with some additionally using recurrence
+[@huSqueezeandExcitationNetworks2017; @vaswaniAttentionAllYou2017].
+This understanding of attention as a mechanism to improve accuracy or other metrics 
+stands in contrast to studies of attention in psychology and neuroscience. 
+In these fields, theories of attention mechanisms are often built on 
+results produced by experimental manipulations that impair behavior.
+Perhaps the pre-eminent example of this approach to understanding attention is a 
+a long-established form of visual search task [@wolfeVisualSearch1998]. In this behavioral
+paradigm, subjects view a set of discrete items on a flat background and report whether a 
+target is present or absent among distractors. Classically, results from this paradigm 
+have been used to support different theories of \bf{selective covert visual attention}, 
+a term that requires some defintion. The term "selective" is simply supposed to mean 
+some mechanism that selects some visual inputs for further processing while necessarily 
+discarding others. The word "covert" in this context means any mechanism that occurs 
+"centrally", i.e. inside the brain, without overt orientation, i.e. without eye or head 
+movements.
+Our goal here is not to resolve the debate about whether selective covert visual 
+attention is serial or parallel. Rather our goal is to better understand attention as 
+defined in cognitive science and reconcile that with the definition used in artificial 
+intelligence.
 

### 2019-05-28 docs/ccneuro2019/paper.tex
yet another rewrite of abstract in ccneuro/paper.tex

@@
 \bf
-Selective attention mechanisms preferentially route some sensory inputs for further 
-processing, while necessarily discarding others. Researchers studying visual
-attention have often divided selective attention into overt and covert mechanisms. 
-Overt mechanisms involve orienting the eyes and/or head, while covert mechanisms 
-occur "centrally", in the brain, without overt orientation. Cognitive scientists have long
-used a new classic form of visual search task to investigate theoretical mechanisms of 
-selective covert visual attention. Here we build on previous studies that examine the 
-parallels between the visual system and convolutional neural networks (CNNs), by testing 
-whether CNNs exhibit human-like attentional effects while performing this visual search 
-task. Following a typical approach, we first show that CNNs trained 
-for image classification can show human-like attentional effects. Then we demonstrate 
-that, by changing how we train the networks and the statistics of the training data, 
-these effects can largely be eliminated, i.e., that CNNs are capable of performing the 
-task without explicit attention mechanisms. We go on to show that manipulating the 
+Convolutional neural networks (CNNs) trained to perform tasks such as image classification
+with human-like accuracy acquire representations similar to those in the primate 
+visual system. Do CNNs that are pre-trained to use such representations conversely exhibit 
+human-like behavior? We test this with a visual search task that has long been used to 
+investigate attention mechanisms. We first show that CNNs using weights trained for image 
+classification can exhibit human-like attentional effects when performing this visual 
+search task. However, we then demonstrate that these effects can be 
+eliminated, without changing the lower-level learned representations, by modifying other 
+parameters of training and the balance of classes in the training data. One explanation 
+for this finding is that we are only training networks to discriminate single types of 
+visual search stimuli. We further test CNNs trained to perform multiple tasks.
+We go on to show that manipulating the 
 discriminability of the target and distractors in the task can produce attentional 

### 2019-05-28 docs/ccneuro2019/paper.tex
bullet-pointing intro in ccneuro/paper.tex

@@
 \subsection{Attention}
-Several attention mechanisms are currently used in machine learning 
-and artificial intelligence to improve classification accuracy and similar 
-metrics. Commonly used attention mechanisms consist essentially of weightings 
-that are learned by updating the weights so as to minimize a loss function, 
-with some additionally using recurrence
-[@huSqueezeandExcitationNetworks2017; @vaswaniAttentionAllYou2017].
-This understanding of attention as a mechanism to improve accuracy or other metrics 
-stands in contrast to studies of attention in psychology and neuroscience. 
+Multiple converging lines of evidence seem to suggest that training neural networks to 
+perform tasks with human-like abilities causes them to acquire brain-like representations.
+
+- If brain-like representations are both necessary and sufficient for human-like 
+intelligence, then shouldn't the converse also be true? Shouldn't neural networks that 
+have acquired brain-like representations behave like humans when asked to perform related 
+tasks? We test this
+
 In these fields, theories of attention mechanisms are often built on 
@@
 movements.
+- here we do not
+
 Our goal here is not to resolve the debate about whether selective covert visual 

### 2019-05-28 src/searchnets/learncurve/learncurve.py
make searchnets/learncurve subpackage, move learncurve.py into it

@@ -11,98 +11,8 @@ from .nets import VGG16
 from .train import batch_generator
 
 
-def learncurve(gz_filename,
-               net_name,
-               number_nets_to_train,
-               input_shape,
-               base_learning_rate,
-               new_learn_rate_layers,
-               new_layer_learning_rate,
-               epochs_list,
-               train_size_list,
-               batch_size,
-               random_seed,
-               model_save_path,
-               test_results_save_path,
-               dropout_rate=0.5,
-               val_size=None):
-    """generate a learning curve, training convolutional neural networks
-    to perform visual search task with a range of sizes of training set.
-
-    Parameters
-    ----------
-    gz_filename : str
-        name of .gz file containing prepared data sets
-    net_name : str
-        name of convolutional neural net architecture to train.
-        One of {'alexnet', 'VGG16'}
-    number_nets_to_train : int
-        number of training "replicates"
-    input_shape : tuple
-        with 3 elements, (rows, columns, channels)
-        should be (227, 227, 3) for AlexNet
-        and (224, 224, 3) for VGG16
-    new_learn_rate_layers : list
-        of layer names whose weights will be initialized randomly
-        and then trained with the 'new_layer_learning_rate'.
-    base_learning_rate : float
-        Applied to layers with weights loaded from training the
-        architecture on ImageNet. Should be a very small number
-        so the trained weights don't change much.
-    new_layer_learning_rate : float
-        Applied to `new_learn_rate_layers'. Should be larger than
-        `base_learning_rate` but still smaller than the usual
-        learning rate for a deep net trained with SGD,
-        e.g. 0.001 instead of 0.01
-    epochs_list : list
-        of training epochs. Replicates will be trained for each
-        value in this list. Can also just be one value, but a list
-        is useful if you want to test whether effects depend on
-        number of training epochs.
-    train_size_list : list
-        of number of samples in training set. Used to generate a learning curve
-        (where x axis is size of training set and y axis is accuracy)
-    batch_size : int
-        number of samples in a batch of training data
-    random_seed : int
-        to seed random number generator
-    model_save_path : str
-        path to directory where model checkpoints should be saved
-    test_results_save_path : str
-        path to directory where results from measuring accuracy on test set should be saved
-    dropout_rate : float
[…]

### 2019-05-28 src/searchnets/learncurve/train.py
add searchnets/learncurve/train.py

@@ -1,0 +1,141 @@
+import os
+
+import numpy as np
+import tensorflow as tf
+import joblib
+from tqdm import tqdm
+
+from searchnets.nets import AlexNet
+from searchnets.nets import VGG16
+from searchnets.train import batch_generator
+
+
+def train(x_train, y_train, train_size_list, epochs_list, net_name, number_nets_to_train, input_shape,
+          new_learn_rate_layers, batch_size, dropout_rate, model_save_path):
+    """train networks for a learning curve"""
+    all_train_inds = np.arange(x_train.shape[0])
+
+    for train_size in train_size_list:
+
+        train_inds = np.random.choice(all_train_inds, train_size)
+        X_data = np.array(x_train)[train_inds]
+        y_data = np.array(y_train)[train_inds]
+
+        for epochs in epochs_list:
+            print(
+                f'training {net_name} model with {train_size} samples for {epochs} epochs'
+            )
+            for net_number in range(number_nets_to_train):
+                print(f'training replicate {net_number + 1} of {number_nets_to_train}')
+                tf.reset_default_graph()
+                graph = tf.Graph()
+                with tf.Session(graph=graph) as sess:
+                    x = tf.placeholder(tf.float32, (None,) + input_shape, name='x')
+                    y = tf.placeholder(tf.int32, shape=[None], name='y')
+                    y_onehot = tf.one_hot(indices=y, depth=len(np.unique(y_train)),
+                                          dtype=tf.float32, name='y_onehot')
+                    rate = tf.placeholder_with_default(tf.constant(1.0, dtype=tf.float32), shape=(),
+                                                       name='dropout_rate')
+
+                    if net_name == 'alexnet':
+                        model = AlexNet(x, init_layer=new_learn_rate_layers, dropout_rate=rate)
+                    elif net_name == 'VGG16':
+                        model = VGG16(x, init_layer=new_learn_rate_layers, dropout_rate=rate)
+
+                    predictions = {
+                        'probabilities': tf.nn.softmax(model.output, name='probabilities'),
+                        'labels': tf.cast(tf.argmax(model.output, axis=1), tf.int32, name='labels')
+                    }
+
+                    cross_entropy_loss = tf.reduce_mean(
+                        tf.nn.softmax_cross_entropy_with_logits_v2(logits=model.output,
+                                                                   labels=y_onehot),
+                        name='cross_entropy_loss')
+
+                    var_list1 = []  # all layers before fully-connected
[…]

### 2019-05-28 src/searchnets/learncurve/test.py
add searchnets/learncurve/test.py

@@ -1,0 +1,128 @@
+import copy
+import csv
+import os
+
+import numpy as np
+import tensorflow as tf
+import joblib
+from tqdm import tqdm
+
+from searchnets.nets import AlexNet
+from searchnets.nets import VGG16
+
+CSV_FIELDNAMES = ['setname', 'train_size', 'epochs', 'net_number', 'err']
+
+
+def test(x_train, y_train, x_test, y_test, set_size_vec_train, set_size_vec_test,
+         net_name, number_nets_to_train, set_sizes, input_shape, new_learn_rate_layers, epochs_list, train_size_list,
+         batch_size, model_save_path, test_results_save_path):
+    """test networks trained for a learning curve"""
+    csv_rows = []
+    results_dict = {}
+
+    for setname, x_data, y_data, set_size_vec in zip(
+            ['train', 'test'],
+            [x_train, x_test],
+            [y_train, y_test],
+            [set_size_vec_train, set_size_vec_test]):
+
+        for train_size in train_size_list:
+            for epochs in epochs_list:
+                print(f'measuring accuracy on {setname} set for {net_name} model trained for {epochs} epochs'
+                      f'with training set of {train_size} samples')
+
+                err_per_set_size_per_model = []
+                err_per_set_size_model_dict = {}
+                predictions_per_model_dict = {}
+
+                for net_number in range(number_nets_to_train):
+                    tf.reset_default_graph()
+                    graph = tf.Graph()
+                    with tf.Session(graph=graph) as sess:
+                        x = tf.placeholder(tf.float32, (None,) + input_shape, name='x')
+                        rate = tf.placeholder_with_default(tf.constant(1.0, dtype=tf.float32), shape=(),
+                                                           name='dropout_rate')
+
+                        if net_name == 'alexnet':
+                            model = AlexNet(x, init_layer=new_learn_rate_layers, dropout_rate=rate)
+                        elif net_name == 'VGG16':
+                            model = VGG16(x, init_layer=new_learn_rate_layers, dropout_rate=rate)
+
+                        predictions = {
+                            'probabilities': tf.nn.softmax(model.output,  # last fully-connected
+                                                           name='probabilities'),
+                            'labels': tf.cast(tf.argmax(model.output, axis=1),
+                                              tf.int32,
+                                              name='labels')
+                        }
+
+                        saver = tf.train.Saver()
+                        restore_path = os.path.join(model_save_path,
[…]

### 2019-05-28 src/searchnets/learncurve/learncurve.py
rewrite learncurve.py to use learncurve.train and learncurve.test

@@ -1,243 +1,23 @@
-# Approach to training based on https://arxiv.org/pdf/1707.09775.pdf
-import os
-
 import numpy as np
 import tensorflow as tf
 import joblib
-from tqdm import tqdm
 
-from .nets import AlexNet
-from .nets import VGG16
-from .train import batch_generator
-
-
-def train(x_train, y_train, train_size_list, net_name, number_nets_to_train, input_shape,
-          new_learn_rate_layers, batch_size, dropout_rate, model_save_path):
-    all_train_inds = np.arange(x_train.shape[0])
-
-    for train_size in train_size_list:
-
-        train_inds = np.random.choice(all_train_inds, train_size)
-        X_data = np.array(x_train)[train_inds]
-        y_data = np.array(y_train)[train_inds]
-
-        for epochs in epochs_list:
-            print(
-                f'training {net_name} model with {train_size} samples for {epochs} epochs'
-            )
-            for net_number in range(number_nets_to_train):
-                print(f'training replicate {net_number + 1} of {number_nets_to_train}')
-                tf.reset_default_graph()
-                graph = tf.Graph()
-                with tf.Session(graph=graph) as sess:
-                    x = tf.placeholder(tf.float32, (None,) + input_shape, name='x')
-                    y = tf.placeholder(tf.int32, shape=[None], name='y')
-                    y_onehot = tf.one_hot(indices=y, depth=len(np.unique(y_train)),
-                                          dtype=tf.float32, name='y_onehot')
-                    rate = tf.placeholder_with_default(tf.constant(1.0, dtype=tf.float32), shape=(),
-                                                       name='dropout_rate')
-
-                    if net_name == 'alexnet':
-                        model = AlexNet(x, init_layer=new_learn_rate_layers, dropout_rate=rate)
-                    elif net_name == 'VGG16':
-                        model = VGG16(x, init_layer=new_learn_rate_layers, dropout_rate=rate)
-
-                    predictions = {
-                        'probabilities': tf.nn.softmax(model.output, name='probabilities'),
-                        'labels': tf.cast(tf.argmax(model.output, axis=1), tf.int32, name='labels')
-                    }
-
-                    cross_entropy_loss = tf.reduce_mean(
-                        tf.nn.softmax_cross_entropy_with_logits_v2(logits=model.output,
-                                                                   labels=y_onehot),
-                        name='cross_entropy_loss')
-
-                    var_list1 = []  # all layers before fully-connected
[…]

### 2019-05-28 src/searchnets/__init__.py
searchnets/__init__.py imports learncurve from .learncurve

@@ -3,4 +3,3 @@ from .train import train
 from .data import data
 from .test import test
 from .learncurve import learncurve
-

### 2019-05-29 src/searchnets/_learncurve/test.py
re-organize learncurve sub-package and learncurve module
- rename sub-pacakge to '_learncurve'
- move learncurve module up to top package level
- learncurve module imports _learncurve.train and _learncurve.test
- if user needs them individually, can access via _learncurve

@@ -1,4 +1,3 @@
-import copy
 import csv
 import os
 
@@ -10,7 +9,7 @@ from tqdm import tqdm
 from searchnets.nets import AlexNet
 from searchnets.nets import VGG16
 
-CSV_FIELDNAMES = ['setname', 'train_size', 'epochs', 'net_number', 'err']
+CSV_FIELDNAMES = ['setname', 'train_size', 'epochs', 'net_number', 'set_size', 'err']
 
 
 def test(x_train, y_train, x_test, y_test, set_size_vec_train, set_size_vec_test,
@@ -75,22 +74,29 @@ def test(x_train, y_train, x_test, y_test, set_size_vec_train, set_size_vec_test
                             y_pred = sess.run(predictions['labels'], feed_dict=feed)
                             y_pred_all.append(y_pred)
                         y_pred_all = np.concatenate(y_pred_all)
-                        acc = np.sum(y_pred_all == y_data) / y_data.shape
+                        acc = np.asscalar(
+                            np.sum(y_pred_all == y_data) / y_data.shape
+                        )
                         err = 1.0 - acc
-                        row = [setname, train_size,  epochs, net_number, err]
+                        row = [setname, train_size,  epochs, net_number, 'all', err]
+                        csv_rows.append(row)
 
                         err_per_set_size = []
                         for set_size in set_sizes:
                             # in line below, [0] at end because np.where returns a tuple
                             inds = np.where(set_size_vec == set_size)[0]
-                            acc_this_set_size = (np.sum(y_pred_all[inds] == y_data[inds]) /
-                                                 y_data[inds].shape)
+                            acc_this_set_size = np.asscalar(
+                                np.sum(y_pred_all[inds] == y_data[inds]) / y_data[inds].shape
+                            )
                             err_this_set_size = 1.0 - acc_this_set_size
                             err_per_set_size.append(err_this_set_size)
-                        row.extend(err_per_set_size)
+                            row = [setname, train_size,  epochs, net_number, set_size, err_this_set_size]
+                            csv_rows.append(row)
 
-                        csv_rows.append(row)
-
+                        # insert into dictionary where model name is key
+                        # and list of accuracies per set size is the "value"
+                        err_per_set_size_model_dict[restore_path] = err_per_set_size
+                        predictions_per_model_dict[restore_path] = y_pred_all
[…]

### 2019-05-29 src/searchnets/learncurve.py

@@ -2,8 +2,8 @@ import numpy as np
 import tensorflow as tf
 import joblib
 
-from .train import train
-from .test import test
+from ._learncurve import train
+from ._learncurve import test
 
 
 def learncurve(gz_filename,

### 2019-05-29 src/searchnets/_learncurve/__init__.py
_learncurve/__init__.py imports train from .train and test from .test

@@ -1,1 +1,2 @@
-from .learncurve import learncurve
+from .train import train
+from .test import test

### 2019-05-29 src/searchnets/__init__.py
do more futzing with learning curve module / sub-package
- decide to name sub-package learncurve
- and name *function* 'learning_curve' that's in learncurve *module*
- and finally import this *function* wherever I need it
- and user can still access learncurve.train/test without being weirded out by underscores

@@ -2,4 +2,4 @@ from . import config
 from .train import train
 from .data import data
 from .test import test
-from .learncurve import learncurve
+from . import learncurve

### 2019-05-29 src/searchnets/__init__.py
searchnets/__init__.py imports learning_curve from .learncurve

@@ -2,4 +2,5 @@ from . import config
 from .train import train
 from .data import data
 from .test import test
-from . import learncurve
+from .learncurve import learning_curve
+

### 2019-05-29 src/searchnets/__main__.py
searchnets/__main__.py imports and uses learning_curve from .learncurve

@@ -12,7 +12,7 @@ from .config import parse_config
 from .data import data
 from .train import train
 from .test import test
-from .learncurve import learncurve
+from .learncurve import learning_curve
 
 
 def _call_data(config):
@@ -62,21 +62,21 @@ def _call_test(config):
 def _call_learncurve(config):
     """helper function to call searchstims.learncurve
     to achieve Don't Repeat Yourself within cli function"""
-    learncurve(gz_filename=config.data.gz_filename,
-               net_name=config.train.net_name,
-               number_nets_to_train=config.train.number_nets_to_train,
-               input_shape=config.train.input_shape,
-               new_learn_rate_layers=config.train.new_learn_rate_layers,
-               base_learning_rate=config.train.base_learning_rate,
-               new_layer_learning_rate=config.train.new_layer_learning_rate,
-               epochs_list=config.train.epochs_list,
-               train_size_list=config.learncurve.train_size_list,
-               batch_size=config.train.batch_size,
-               random_seed=config.train.random_seed,
-               model_save_path=config.train.model_save_path,
-               test_results_save_path=config.test.test_results_save_path,
-               dropout_rate=config.train.dropout_rate,
-               val_size=config.data.val_size)
+    learning_curve(gz_filename=config.data.gz_filename,
+                   net_name=config.train.net_name,
+                   number_nets_to_train=config.train.number_nets_to_train,
+                   input_shape=config.train.input_shape,
+                   new_learn_rate_layers=config.train.new_learn_rate_layers,
+                   base_learning_rate=config.train.base_learning_rate,
+                   new_layer_learning_rate=config.train.new_layer_learning_rate,
+                   epochs_list=config.train.epochs_list,
+                   train_size_list=config.learncurve.train_size_list,
+                   batch_size=config.train.batch_size,
+                   random_seed=config.train.random_seed,
+                   model_save_path=config.train.model_save_path,
+                   test_results_save_path=config.test.test_results_save_path,
+                   dropout_rate=config.train.dropout_rate,
+                   val_size=config.data.val_size)
 
 
 def cli(command, configfile):

### 2019-05-29 src/searchnets/learncurve/__init__.py
learncurve/__init__.py imports learning_curve from .learncurve

@@ -1,2 +1,3 @@
 from .train import train
 from .test import test
+from .learncurve import learning_curve

### 2019-05-29 src/searchnets/learncurve/test.py
change absolute imports to relative in learncurve/test.py

@@ -6,8 +6,8 @@ import tensorflow as tf
 import joblib
 from tqdm import tqdm
 
-from searchnets.nets import AlexNet
-from searchnets.nets import VGG16
+from ..nets import AlexNet
+from ..nets import VGG16
 
 CSV_FIELDNAMES = ['setname', 'train_size', 'epochs', 'net_number', 'set_size', 'err']
 

### 2019-05-29 src/searchnets/config/parse.py
config.parse.parse_config checks if config_fname is a file
to avoid confusing non-error when ConfigParser gives a KeyError for a
key in a file that doesn't exist

@@ -1,6 +1,7 @@
 import ast
 import configparser
 from distutils.util import strtobool
+import os
 
 from .classes import TrainConfig, DataConfig, TestConfig, LearnCurveConfig, Config
 
@@ -20,6 +21,11 @@ def parse_config(config_fname):
         instance of searchstims.config.classes.Config,
         attrs-based class that represents all configuration parameters
     """
+    if not os.path.isfile(config_fname):
+        raise FileNotFoundError(
+            f'specified config.ini file not found: {configfile}'
+        )
+
     config = configparser.ConfigParser()
     config.read(config_fname)
 

### 2019-05-29 src/searchnets/train.py
change order of parameters in searchnets.train docstring to match function signature

@@ -84,6 +84,15 @@ def train(gz_filename,
         with 3 elements, (rows, columns, channels)
         should be (227, 227, 3) for AlexNet
         and (224, 224, 3) for VGG16
+    base_learning_rate : float
+        Applied to layers with weights loaded from training the
+        architecture on ImageNet. Should be a very small number
+        so the trained weights don't change much.
+    freeze_trained_weights : bool
+        if True, freeze weights in any layer not in "new_learn_rate_layers".
+        These are the layers that have weights pre-trained on ImageNet.
+        Default is False. Done by simply not applying gradients to these weights,
+        i.e. this will ignore a base_learning_rate if you set it to something besides zero.
     new_learn_rate_layers : list
         of layer names whose weights will be initialized randomly
         and then trained with the 'new_layer_learning_rate'.
@@ -103,15 +112,6 @@ def train(gz_filename,
         to seed random number generator
     model_save_path : str
         path to directory where model checkpoints should be saved
-    base_learning_rate : float
-        Applied to layers with weights loaded from training the
-        architecture on ImageNet. Should be a very small number
-        so the trained weights don't change much.
-    freeze_trained_weights : bool
-        if True, freeze weights in any layer not in "new_learn_rate_layers".
-        These are the layers that have weights pre-trained on ImageNet.
-        Default is False. Done by simply not applying gradients to these weights,
-        i.e. this will ignore a base_learning_rate if you set it to something besides zero.
     dropout_rate : float
         Probability that any unit in a layer will "drop out" during
         a training epoch, as a form of regularization. Default is 0.5.

### 2019-05-29 src/searchnets/learncurve/test.py
re-order parameters in signature of learncurve.test and rewrite docstring

@@ -12,10 +12,73 @@ from ..nets import VGG16
 CSV_FIELDNAMES = ['setname', 'train_size', 'epochs', 'net_number', 'set_size', 'err']
 
 
-def test(x_train, y_train, x_test, y_test, set_size_vec_train, set_size_vec_test,
-         net_name, number_nets_to_train, set_sizes, input_shape, new_learn_rate_layers, epochs_list, train_size_list,
-         batch_size, model_save_path, test_results_save_path):
-    """test networks trained for a learning curve"""
+def test(x_train,
+         y_train,
+         x_test,
+         y_test,
+         set_size_vec_train,
+         set_size_vec_test,
+         set_sizes,
+         net_name,
+         number_nets_to_train,
+         input_shape,
+         new_learn_rate_layers,
+         epochs_list,
+         train_size_list,
+         batch_size,
+         model_save_path,
+         test_results_save_path):
+    """test networks trained for a learning curve
+
+    Parameters
+    ----------
+    x_train : numpy.ndarray
+        training data, visual search stimuli
+    y_train. numpy.ndarray
+        expected outputs, vector of 0s and 1s (for 'target absent' and 'target present')
+    x_test : numpy.ndarray
+        test, visual search stimuli
+    y_test. numpy.ndarray
+        expected outputs for test set
+    set_size_vec_train : numpy.ndarray
+        vector whose length is equal to number of samples in x_train, each value indicates (visual search) set size for
+        sample at corresponding index in x_train
+    set_size_vec_test : numpy.ndarray
+        vector whose length is equal to number of samples in x_test, each value indicates (visual search) set size for
+        sample at corresponding index in x_test
+    set_sizes : list
+        of int, unique set of set sizes in data
+    net_name : str
+        name of convolutional neural net architecture to train.
+        One of {'alexnet', 'VGG16'}
+    number_nets_to_train : int
+        number of training "replicates"
+    input_shape : tuple
+        with 3 elements, (rows, columns, channels)
+        should be (227, 227, 3) for AlexNet
+        and (224, 224, 3) for VGG16
+    new_learn_rate_layers : list
+        of layer names whose weights will be initialized randomly
+        and then trained with the 'new_layer_learning_rate'.
+    epochs_list : list
+        of training epochs. Replicates will be trained for each
+        value in this list. Can also just be one value, but a list
+        is useful if you want to test whether effects depend on
+        number of training epochs.
+    train_size_list : list
+        of number of samples in training set. Used to generate a learning curve
[…]

### 2019-05-29 src/searchnets/learncurve/train.py
rewrite learncurve.train to match searchstims.train
- consists of following parameters:
  + base_learing_rate, freeze_trained_weights, new_layer_learning_rate
- not sure if I was using these when learncurve was a single function

@@ -5,14 +5,79 @@ import tensorflow as tf
 import joblib
 from tqdm import tqdm
 
-from searchnets.nets import AlexNet
-from searchnets.nets import VGG16
-from searchnets.train import batch_generator
+from ..nets import AlexNet
+from ..nets import VGG16
+from ..train import batch_generator
 
 
-def train(x_train, y_train, train_size_list, epochs_list, net_name, number_nets_to_train, input_shape,
-          new_learn_rate_layers, batch_size, dropout_rate, model_save_path):
-    """train networks for a learning curve"""
+def train(x_train,
+          y_train,
+          net_name,
+          number_nets_to_train,
+          input_shape,
+          base_learning_rate,
+          freeze_trained_weights,
+          new_learn_rate_layers,
+          new_layer_learning_rate,
+          train_size_list,
+          epochs_list,
+          batch_size,
+          dropout_rate,
+          model_save_path):
+    """train networks for a learning curve
+
+    Parameters
+    ----------
+    x_train : numpy.ndarray
+        training data, visual search stimuli
+    y_train. numpy.ndarray
+        expected outputs, vector of 0s and 1s (for 'target absent' and 'target present')
+    net_name : str
+        name of convolutional neural net architecture to train.
+        One of {'alexnet', 'VGG16'}
+    number_nets_to_train : int
+        number of training "replicates"
+    input_shape : tuple
+        with 3 elements, (rows, columns, channels)
+        should be (227, 227, 3) for AlexNet
+        and (224, 224, 3) for VGG16
+    base_learning_rate : float
+        Applied to layers with weights loaded from training the
+        architecture on ImageNet. Should be a very small number
+        so the trained weights don't change much.
+    freeze_trained_weights : bool
+        if True, freeze weights in any layer not in "new_learn_rate_layers".
+        These are the layers that have weights pre-trained on ImageNet.
+        Default is False. Done by simply not applying gradients to these weights,
+        i.e. this will ignore a base_learning_rate if you set it to something besides zero.
+    new_learn_rate_layers : list
+        of layer names whose weights will be initialized randomly
+        and then trained with the 'new_layer_learning_rate'.
+    new_layer_learning_rate : float
+        Applied to `new_learn_rate_layers'. Should be larger than
+        `base_learning_rate` but still smaller than the usual
+        learning rate for a deep net trained with SGD,
+        e.g. 0.001 instead of 0.01
+    train_size_list : list
+        of number of samples in training set. Used to generate a learning curve
[…]

### 2019-05-29 src/searchnets/learncurve/learncurve.py
fix learning_curve so it calls train and test with updated signatures

@@ -2,22 +2,25 @@ import numpy as np
 import tensorflow as tf
 import joblib
 
-from ._learncurve import train
-from ._learncurve import test
+from .train import train
+from .test import test
 
 
-def learncurve(gz_filename,
-               net_name,
-               number_nets_to_train,
-               input_shape,
-               new_learn_rate_layers,
-               epochs_list,
-               train_size_list,
-               batch_size,
-               random_seed,
-               model_save_path,
-               test_results_save_path,
-               dropout_rate=0.5):
+def learning_curve(gz_filename,
+                   net_name,
+                   number_nets_to_train,
+                   input_shape,
+                   base_learning_rate,
+                   freeze_trained_weights,
+                   new_learn_rate_layers,
+                   new_layer_learning_rate,
+                   epochs_list,
+                   train_size_list,
+                   batch_size,
+                   random_seed,
+                   model_save_path,
+                   test_results_save_path,
+                   dropout_rate=0.5):
     """generate a learning curve, training convolutional neural networks
     to perform visual search task with a range of sizes of training set.
 
@@ -41,6 +44,14 @@ def learncurve(gz_filename,
         Applied to layers with weights loaded from training the
         architecture on ImageNet. Should be a very small number
         so the trained weights don't change much.
+    freeze_trained_weights : bool
+        if True, freeze weights in any layer not in "new_learn_rate_layers".
+        These are the layers that have weights pre-trained on ImageNet.
+        Default is False. Done by simply not applying gradients to these weights,
+        i.e. this will ignore a base_learning_rate if you set it to something besides zero.
+    new_learn_rate_layers : list
+        of layer names whose weights will be initialized randomly
+        and then trained with the 'new_layer_learning_rate'.
     new_layer_learning_rate : float
         Applied to `new_learn_rate_layers'. Should be larger than
         `base_learning_rate` but still smaller than the usual
@@ -101,9 +112,35 @@ def learncurve(gz_filename,
                 continue
 
     # -------------------- first train a bunch of models ---------------------------------------------------------------
-    train(x_train, y_train, train_size_list, net_name, number_nets_to_train, input_shape,
-          new_learn_rate_layers, batch_size, dropout_rate, model_save_path)
+    train(x_train,
+          y_train,
+          net_name,
+          number_nets_to_train,
+          input_shape,
+          base_learning_rate,
+          freeze_trained_weights,
[…]

### 2019-05-29 setup.py
add 'pandas' and 'seaborn' to REQUIRED in setup.py

@@ -24,7 +24,8 @@ VERSION = '0.1.0'
 
 # What packages are required for this module to be executed?
 REQUIRED = [
-    'numpy', 'tensorflow', 'matplotlib', 'imageio', 'joblib', 'tqdm', 'attrs', 'searchstims>=1.2.0',
+    'numpy', 'tensorflow', 'matplotlib', 'pandas', 'seaborn', 'imageio', 'joblib', 'tqdm', 'attrs',
+    'searchstims>=1.2.0',
 ]
 
 # What packages are optional?

### 2019-05-29 src/searchnets/plot/figures.py
add learncurve function to plot/figures.py

@@ -2,11 +2,14 @@ import os
 import json
 from glob import glob
 
+import joblib
 import numpy as np
 import matplotlib as mpl
 import matplotlib.pyplot as plt
-import joblib
+import pandas as pd
 from scipy import stats
+import seaborn as sns
+
 
 mpl.style.use('bmh')
 
@@ -380,5 +383,24 @@ def train_history(acc_dir, set_sizes=(1, 2, 4, 8), save_as=None):
             ax[-ind].set_visible(False)
 
     fig.tight_layout()
+    if save_as:
+        plt.savefig(save_as)
+
+
+def learncurve(csv_fname, figsize=(15, 10), save_as=None):
+    """"""
+    df = pd.read_csv(csv_fname)
+    set_sizes = sorted(df.set_size.unique())
+
+    fig, ax = plt.subplots(1, len(set_sizes), figsize=figsize)
+    for ax_ind, set_size in enumerate(set_sizes):
+        df_this_ax = df.loc[df['set_size'] == set_size]
+        sns.scatterplot(x='train_size', y='err', hue='setname', data=df_this_ax, legend=False, ax=ax[ax_ind])
+        sns.lineplot(x='train_size', y='err', hue='setname', data=df_this_ax, ax=ax[ax_ind])
+        ax[ax_ind].set_ylim([0, 0.5])
+        ax[ax_ind].set_xticks(df.train_size.unique())
+        ax[ax_ind].set_xticklabels(df.train_size.unique(), rotation=45)
+        ax[ax_ind].set_title(f'set size: {set_size}')
+
     if save_as:
         plt.savefig(save_as)

### 2019-05-29 data/runall_alexnet_learncurve_2_v_5.sh
add other learncurve configs to runall_alexnet_learncurve_2_v_5.sh

@@ -1,5 +1,10 @@
 #!/usr/bin/env bash
 
-# prepare data sets for training neural nets, then train, then test accuracy of trained nets
+searchnets data ./configs/searchnets_alexnet_learncurve_RVvGV.ini
+searchnets learncurve ./configs/searchnets_alexnet_learncurve_RVvGV.ini
+
+searchnets data ./configs/searchnets_alexnet_learncurve_RVvRHGV.ini
+searchnets learncurve ./configs/searchnets_alexnet_learncurve_RVvRHGV.ini
+
 searchnets data ./configs/searchnets_alexnet_learncurve_2_v_5.ini
 searchnets learncurve ./configs/searchnets_alexnet_learncurve_2_v_5.ini

### 2019-05-30 src/searchnets/plot/figures.py
add ylim and suptitle parameters to plot.figures.learncurve

@@ -387,7 +387,7 @@ def train_history(acc_dir, set_sizes=(1, 2, 4, 8), save_as=None):
         plt.savefig(save_as)
 
 
-def learncurve(csv_fname, figsize=(15, 10), save_as=None):
+def learncurve(csv_fname, figsize=(15, 10), ylim=(0, 0.5), suptitle=None, save_as=None):
     """"""
     df = pd.read_csv(csv_fname)
     set_sizes = sorted(df.set_size.unique())
@@ -397,10 +397,18 @@ def learncurve(csv_fname, figsize=(15, 10), save_as=None):
         df_this_ax = df.loc[df['set_size'] == set_size]
         sns.scatterplot(x='train_size', y='err', hue='setname', data=df_this_ax, legend=False, ax=ax[ax_ind])
         sns.lineplot(x='train_size', y='err', hue='setname', data=df_this_ax, ax=ax[ax_ind])
-        ax[ax_ind].set_ylim([0, 0.5])
+        ax[ax_ind].set_ylim(ylim)
         ax[ax_ind].set_xticks(df.train_size.unique())
         ax[ax_ind].set_xticklabels(df.train_size.unique(), rotation=45)
         ax[ax_ind].set_title(f'set size: {set_size}')
+        if ax_ind == 0:
+            ax[ax_ind].set_ylabel('error')
+        else:
+            ax[ax_ind].set_ylabel('')
+        ax[ax_ind].set_xlabel('num. training samples')
+
+    if suptitle:
+        fig.suptitle(suptitle)
 
     if save_as:
         plt.savefig(save_as)

### 2019-05-30 src/searchnets/plot/figures.py
replace rcParrams block at top of figures.py with single call to sns.set_style

@@ -4,23 +4,12 @@ from glob import glob
 
 import joblib
 import numpy as np
-import matplotlib as mpl
 import matplotlib.pyplot as plt
 import pandas as pd
 from scipy import stats
 import seaborn as sns
 
-
-mpl.style.use('bmh')
-
-plt.rcParams['font.size'] = 18
-plt.rcParams['axes.labelsize'] = 18
-plt.rcParams['axes.labelweight'] = 'regular'
-plt.rcParams['axes.titlesize'] = 20
-plt.rcParams['xtick.labelsize'] = 16
-plt.rcParams['ytick.labelsize'] = 16
-plt.rcParams['legend.fontsize'] = 16
-plt.rcParams['figure.titlesize'] = 20
+sns.set_style("whitegrid")
 
 
 def acc_v_set_size(results, set_sizes=(1, 2, 4, 8), ax=None,

### 2019-05-30 setup.py
add 'scipy' to REQUIRED in setup.py

@@ -24,7 +24,7 @@ VERSION = '0.1.0'
 
 # What packages are required for this module to be executed?
 REQUIRED = [
-    'numpy', 'tensorflow', 'matplotlib', 'pandas', 'seaborn', 'imageio', 'joblib', 'tqdm', 'attrs',
+    'numpy', 'scipy', 'tensorflow', 'matplotlib', 'pandas', 'seaborn', 'imageio', 'joblib', 'tqdm', 'attrs',
     'searchstims>=1.2.0',
 ]
 

### 2019-05-30 src/searchnets/train.py
rewrite src/searchnets/train.py
- add parameters: use_val, val_step, patience
- do pre-processing of data sets within train function
- function now also saves training history for each net as a csv

@@ -1,4 +1,5 @@
 # Approach to training based on https://arxiv.org/pdf/1707.09775.pdf
+import csv
 import os
 
 import numpy as np
@@ -8,6 +9,8 @@ from tqdm import tqdm
 
 from .nets import AlexNet
 from .nets import VGG16
+
+IMAGENET_MEAN = [104.00698793, 116.66876762, 122.67891434]
 
 
 def batch_generator(X, y, batch_size=64,
@@ -67,7 +70,9 @@ def train(gz_filename,
           random_seed,
           model_save_path,
           dropout_rate=0.5,
-          val_size=None,
+          use_val=True,
+          val_step=None,
+          patience=None,
           save_acc_by_set_size_by_epoch=False):
     """train convolutional neural networks to perform visual search task.
 
@@ -121,24 +126,47 @@ def train(gz_filename,
         are saved in a matrix where rows are epochs and columns are set sizes.
         Useful for seeing whether accuracy converges for each individual
         set size. Default is False.
-    val_size : int
-        number of samples in validation set. Default is None.
+    use_val : bool
+        if True, use validation set.
+    val_step : int
+        if not None, accuracy on validation set will be measured every `val_step` steps
+    patience : int
+        if not None, training will stop if accuracy on validation set has not improved in `patience` steps
 
     Returns
     -------
     None
     """
+    if use_val and val_step is None or val_step < 1 or type(val_step) != int:
+        raise ValueError(
+            'invalid value for val_step: {val_step}. Validation step must be positive integer'
+        )
+
+    if use_val is False and patience is not None:
+        raise ValueError('patience argument only works with a validation set')
+
+    if patience is not None:
+        if type(val_step) != int or patience < 1:
+            raise TypeError('patience must be a positive integer')
+
     print('loading training data')
     data_dict = joblib.load(gz_filename)
     x_train = data_dict['x_train']
+    # pre-process images
+    x_train = np.asarray([img[:, :, [2, 1, 0]] - IMAGENET_MEAN for img in x_train])
     y_train = data_dict['y_train']
 
-    training_set = [x_train, y_train]
-    if val_size:
-        val_set = [data_dict['x_val'],
-                   data_dict['y_val']]
+    if use_val:
+        try:
+            x_val = data_dict['x_val']
+        except KeyError:
+            raise KeyError(
+                f'use_val set to True but x_val not found in data file: {gz_filename}'
[…]

### 2019-05-30 src/searchnets/train.py
fixup train

@@ -374,9 +374,11 @@ def train(gz_filename,
                         print(acc_set_size_str)
 
                 # --------------- done training, save checkpoint + training history info -------------------------------
-                print(f'Saving model in {savepath}')
-                ckpt_name = os.path.join(savepath, f'{net_name}-model-epoch-{epoch}.ckpt')
-                saver.save(sess, ckpt_name, global_step=epochs)
+                if patience is None:
+                    # only save at end if we haven't already been saving checkpoints
+                    print(f'Saving model in {savepath}')
+                    ckpt_name = os.path.join(savepath, f'{net_name}-model-epoch-{epoch}.ckpt')
+                    saver.save(sess, ckpt_name, global_step=epochs)
 
                 stem = f'{net_name}_trained_{epochs}_epochs_number_{net_number}'
 

### 2019-05-30 src/searchnets/config/classes.py
add new parameters for train function as attributes of TrainConfig

@@ -5,6 +5,17 @@ import attr
 from attr.validators import instance_of, optional
 
 
+def is_pos_int(instance, attribute, value):
+    if type(value) != int:
+        raise ValueError(
+            f'type of {attribute.name} must be an int'
+        )
+    if value < 1:
+        raise ValueError(
+            f'{attribute.name} must be a positive integer, but was: {value}'
+        )
+
+
 @attr.s
 class TrainConfig:
     """class to represent [TRAIN] section of config.ini file
@@ -57,6 +68,12 @@ class TrainConfig:
         are saved in a matrix where rows are epochs and columns are set sizes.
         Useful for seeing whether accuracy converges for each individual
         set size. Default is False.
+    use_val : bool
+        if True, use validation set.
+    val_step : int
+        if not None, accuracy on validation set will be measured every `val_step` steps
+    patience : int
+        if not None, training will stop if accuracy on validation set has not improved in `patience` steps
     """
     net_name = attr.ib(validator=instance_of(str))
     number_nets_to_train = attr.ib(validator=instance_of(int))
@@ -94,6 +111,9 @@ class TrainConfig:
     freeze_trained_weights = attr.ib(validator=instance_of(bool), default=False)
     dropout_rate = attr.ib(validator=instance_of(float), default=0.5)
     save_acc_by_set_size_by_epoch = attr.ib(validator=instance_of(bool), default=False)
+    use_val = attr.ib(validator=instance_of(bool), default=False)
+    val_step = attr.ib(validator=optional(is_pos_int), default=None)
+    patience = attr.ib(validator=optional(is_pos_int), default=None)
 
 
 @attr.s

### 2019-05-30 src/searchnets/config/parse.py
parse_config unpacks new parameters from .ini file and adds to TrainConfig

@@ -61,6 +61,21 @@ def parse_config(config_fname):
     else:
         save_acc_by_set_size_by_epoch = False
 
+    if config.has_option('TRAIN', 'USE_VAL'):
+        use_val = bool(strtobool(config['TRAIN']['USE_VAL']))
+    else:
+        use_val = True
+
+    if config.has_option('TRAIN', 'VAL_STEP'):
+        val_step = int(config['TRAIN']['VAL_STEP'])
+    else:
+        val_step = None
+
+    if config.has_option('TRAIN', 'PATIENCE'):
+        patience = int(config['TRAIN']['PATIENCE'])
+    else:
+        patience = None
+
     train_config = TrainConfig(net_name,
                                number_nets_to_train,
                                input_shape,
@@ -73,7 +88,10 @@ def parse_config(config_fname):
                                base_learning_rate,
                                freeze_trained_weights,
                                dropout_rate,
-                               save_acc_by_set_size_by_epoch)
+                               save_acc_by_set_size_by_epoch,
+                               use_val,
+                               val_step,
+                               patience)
 
     # ------------- unpack [DATA] section of config.ini file -----------------------------------------------------------
     train_dir = config['DATA']['TRAIN_DIR']

### 2019-06-03 src/searchnets/data.py
rewrite data.py to save filenames not images

@@ -4,9 +4,7 @@ from glob import glob
 import random
 
 import numpy as np
-import imageio
 import joblib
-from tqdm import tqdm
 
 
 def data(train_dir,
@@ -41,24 +39,18 @@ def data(train_dir,
     -----
     The dictionary saved has the following key-value pairs:
         x_train : np.ndarray
-            images used as input to train neural network
+            filenames corresponding to images used as input to train neural network
         y_train : np.ndarray
             labels, expected output of neural network.
             Either element is either 1, meaning "target present", or 0, "target absent".
         x_val : np.ndarray
-            images for validation set used during training
+            filenames corresponding to images for validation set used during training
         y_val : np.ndarray
             labels for validation set
         x_test : np.ndarray
-            images for set used to test accuracy of trained network
+            filenames corresponding to images for set used to test accuracy of trained network
         y_test : np.ndarray
             labels for set used to test accuracy of trained network
-        train_set_files : np.ndarray
-            filenames corresponding to images in x_train
-        val_set_files : np.ndarray
-            filenames corresponding to images in x_val
-        test_set_files : np.ndarray
-            filenames corresponding to images in x_test
         set_size_vec_train : np.ndarray
             "set size" of each image in x_train,
             total number of targets + distractors.
@@ -89,7 +81,7 @@ def data(train_dir,
     else:
         fname_json = fname_json[0]
 
-    train_set_files = []
+    x_train = []
     # initialize list to convert into a
     # vector that will indicates set size
     # for each image (number of items present)
@@ -97,13 +89,13 @@ def data(train_dir,
     set_size_vec_train = []
     stim_type_vec_train = []
     if val_size:
-        val_set_files = []
+        x_val = []
         set_size_vec_val = []
         stim_type_vec_val = []
     else:
-        val_set_files = None
+        x_val = None
         set_size_vec_val = None
-    test_set_files = []
+    x_test = []
     set_size_vec_test = []
     stim_type_vec_test = []
 
@@ -214,29 +206,29 @@ def data(train_dir,
                 train_inds = np.asarray(
                     [inds.pop() for _ in range(train_size_per_set_size)]
                 )
-                tmp_train_set_files = stim_fnames_arr[train_inds].tolist()
[…]

### 2019-06-03 src/searchnets/nets/alexnet.py
rename utils/data.py to utils/figshare.py

@@ -6,7 +6,7 @@ import os
 import tensorflow as tf
 import numpy as np
 
-from ..utils.data import fetch
+from ..utils.figshare import fetch
 from ..utils.figshare_urls import ALEXNET_WEIGHTS_URL
 from .layers import max_pool, lrn, dropout
 

### 2019-06-03 src/searchnets/utils/figshare.py

@@ -1,3 +1,4 @@
+"""utility functions for getting files off figshare"""
 # utility function for fetching files
 # Adapted from MNE-Python under BSD-3 License
 # https://github.com/mne-tools/mne-python/blob/master/mne/utils.py#L1884

### 2019-06-03 src/searchnets/data.py
data.py converts all image paths to absolute

@@ -230,6 +230,11 @@ def data(train_dir,
                 set_size_vec_test.extend([set_size] * len(tmp_x_test))
                 stim_type_vec_train.extend([stim_type] * len(tmp_x_test))
 
+    x_train = [os.path.join(os.path.abspath(train_dir), path) for path in x_train]
+    if x_val is not None:
+        x_val = [os.path.join(os.path.abspath(train_dir), path) for path in x_val]
+    x_test = [os.path.join(os.path.abspath(train_dir), path) for path in x_test]
+
     set_size_vec_train = np.asarray(set_size_vec_train)
     stim_type_vec_train = np.asarray(stim_type_vec_train)
     if val_size:

### 2019-06-03 src/searchnets/utils/tfdata.py
add utils/tfdata.py with functions for using tf.data.Dataset

@@ -1,0 +1,62 @@
+"""utility functions for tf.data.Datasets"""
+import numpy as np
+import tensorflow as tf
+
+IMAGENET_MEANS = np.asarray([123.68, 116.779, 103.939])  # in RGB order
+
+
+def load(x):
+    x = tf.read_file(x)
+    x = tf.image.decode_png(x)
+    x = tf.cast(x, tf.float32)
+    return x
+
+
+def preprocess(x, convert_bgr=True):
+    x = x - IMAGENET_MEANS  # do this before flipping channel order
+    if convert_bgr:
+        # make RGB into BGR
+        x = tf.reverse(x, axis=[-1])
+    return x
+
+
+def _generic_dataset(x, y, preprocess_func, batch_size, shuffle=True):
+    if type(x) == list:
+        x = np.asarray(x)
+    x_ds = tf.data.Dataset.from_tensor_slices(x)
+    x_ds = x_ds.map(load)
+    x_ds = x_ds.map(preprocess_func)
+    y_ds = tf.data.Dataset.from_tensor_slices(y)
+    ds = tf.data.Dataset.zip((x_ds, y_ds))
+    if shuffle:
+        ds = ds.shuffle(buffer_size=x.shape[-1])
+    ds = ds.batch(batch_size)
+    ds = ds.prefetch(buffer_size=x.shape[-1])
+    return ds
+
+
+def alexnet_preprocess():
+    def anonymous(x):
+        return preprocess(x, convert_bgr=True)
+    return anonymous
+
+
+def alexnet_dataset(x, y, batch_size, shuffle=True):
+    return _generic_dataset(x, y, alexnet_preprocess(), batch_size, shuffle)
+
+
+def vgg16_preprocess():
+    def anonymous(x):
+        return preprocess(x, convert_bgr=False)  # because weights we're using have channels in 1st layer switched
+    return anonymous
+
+
+def vgg16_dataset(x, y, batch_size, shuffle=True):
+    return _generic_dataset(x, y, vgg16_preprocess(), batch_size, shuffle)
+
+
+def get_dataset(x, y, net_name, batch_size, shuffle=True):
+    if net_name == 'alexnet':
+        return alexnet_dataset(x, y, batch_size, shuffle)
+    elif net_name == 'VGG16':
+        return vgg16_dataset(x, y, batch_size, shuffle)

### 2019-06-03 src/searchnets/train.py
rewrite searchnets/train.py to use tf.data.Dataset

@@ -9,53 +9,9 @@ from tqdm import tqdm
 
 from .nets import AlexNet
 from .nets import VGG16
+from .utils.tfdata import get_dataset
 
-IMAGENET_MEAN = [104.00698793, 116.66876762, 122.67891434]
 MOMENTUM = 0.9  # used for both Alexnet and VGG16
-
-
-def batch_generator(X, y, batch_size=64,
-                    shuffle=False, random_seed=None):
-    """generator that yields batches of training data and labels
-
-    Parameters
-    ----------
-    X : numpy.ndarray
-        training data, e.g., images with dimensions (number of samples, height, width, channels)
-    y : numpy.ndarray
-        labels for training data, e.g. vector of integers
-    batch_size : int
-        number of elements in each batch from X and y. Default is 64.
-    shuffle : bool
-        if True, shuffle data before creating generator. Default is False.
-    random_seed : int
-        integer to seed random number generator. Default is None.
-
-    Returns
-    -------
-    batch_x : numpy.ndarray
-        subset of X where first dimension is of size batch_size
-    batch_y : numpy.ndarray
-        subset of y where first dimension is of size batch_size
-
-    Notes
-    -----
-    adapted from code by Sebastian Raschka under MIT license
-    https://github.com/rasbt/python-machine-learning-book-2nd-edition/blob/master/LICENSE.txt
-    """
-    idx = np.arange(y.shape[0])
-
-    X_for_batch = np.copy(X)
-    y_for_batch = np.copy(y)
-
-    if shuffle:
-        rng = np.random.RandomState(random_seed)
-        rng.shuffle(idx)
-        X_for_batch = X_for_batch[idx]
-        y_for_batch = y_for_batch[idx]
-
-    for i in range(0, X.shape[0], batch_size):
-        yield (X_for_batch[i:i + batch_size, :], y_for_batch[i:i + batch_size])
 
 
 def train(gz_filename,
@@ -152,22 +108,17 @@ def train(gz_filename,
 
     print('loading training data')
     data_dict = joblib.load(gz_filename)
-    x_train = data_dict['x_train']
-    # pre-process images
-    x_train = np.asarray([img[:, :, [2, 1, 0]] - IMAGENET_MEAN for img in x_train])
-    y_train = data_dict['y_train']
+    train_ds = get_dataset(data_dict['x_train'], data_dict['y_train'], net_name, batch_size, shuffle=True)
 
     if use_val:
         try:
[…]

### 2019-06-03 src/searchnets/test.py
rewrite searchnets/test.py to use tf.data.Dataset

@@ -6,7 +6,7 @@ import joblib
 from tqdm import tqdm
 
 from .nets import AlexNet, VGG16
-from .train import IMAGENET_MEAN
+from .utils.tfdata import get_dataset
 
 
 def test(gz_filename,
@@ -65,10 +65,8 @@ def test(gz_filename,
     """
     print('loading testing data')
     data_dict = joblib.load(gz_filename)
-    x_test = data_dict['x_test']
-    # pre-process images
-    x_test = np.asarray([img[:, :, [2, 1, 0]] - IMAGENET_MEAN for img in x_test])
-    y_test = data_dict['y_test']
+    test_ds = get_dataset(data_dict['x_test'], data_dict['y_test'], net_name, batch_size, shuffle=False)
+
     set_sizes_by_stim_type = data_dict['set_sizes_by_stim_stype']
     set_sizes = []
     for stim_type, set_sizes_this_stim in set_sizes_by_stim_type.items():
@@ -93,8 +91,6 @@ def test(gz_filename,
             with tf.Session(graph=graph) as sess:
                 x = tf.placeholder(tf.float32, (None,) + input_shape, name='x')
                 y = tf.placeholder(tf.int32, shape=[None], name='y')
-                y_onehot = tf.one_hot(indices=y, depth=len(np.unique(y_test)),
-                                      dtype=tf.float32, name='y_onehot')
                 rate = tf.placeholder_with_default(tf.constant(1.0, dtype=tf.float32), shape=(), name='dropout_rate')
 
                 if net_name == 'alexnet':
@@ -118,16 +114,23 @@ def test(gz_filename,
                 ckpt_path = os.path.join(restore_path, f'{net_name}-model.ckpt-{epochs}')
                 saver.restore(sess, ckpt_path)
 
+                total = int(np.ceil(len(data_dict['x_test']) / batch_size))
+                total_loss = 0.0
+                iterator = test_ds.make_one_shot_iterator()
+                next_element = iterator.get_next()
+
+                y_test = []
                 y_pred_all = []
-                batch_inds = np.arange(0, x_test.shape[0], batch_size)
-                total = len(batch_inds)
-                pbar = tqdm(enumerate(batch_inds), total=total)
-                for count, ind in pbar:
-                    pbar.set_description(f'predicting target present/absent for batch {count} of {total}')
-                    feed = {x: x_test[ind:ind + batch_size, :, :]}
+                pbar = tqdm(range(total))
+                for i in pbar:
+                    batch_x, batch_y = sess.run(next_element)
[…]

### 2019-06-03 src/searchnets/train.py
fix ckpt_name in train.py so it matches what test.py expects

@@ -334,7 +334,7 @@ def train(gz_filename,
                 if patience is None:
                     # only save at end if we haven't already been saving checkpoints
                     print(f'Saving model in {savepath}')
-                    ckpt_name = os.path.join(savepath, f'{net_name}-model-epoch-{epoch}.ckpt')
+                    ckpt_name = os.path.join(savepath, f'{net_name}-model.ckpt')
                     saver.save(sess, ckpt_name, global_step=epochs)
 
                 stem = f'{net_name}_trained_{epochs}_epochs_number_{net_number}'

### 2019-06-03 src/searchnets/utils/tfdata.py
change buffer_size in call to prefetch to batch_size

@@ -31,7 +31,7 @@ def _generic_dataset(x, y, preprocess_func, batch_size, shuffle=True):
     if shuffle:
         ds = ds.shuffle(buffer_size=x.shape[-1])
     ds = ds.batch(batch_size)
-    ds = ds.prefetch(buffer_size=x.shape[-1])
+    ds = ds.prefetch(buffer_size=batch_size)
     return ds
 
 

### 2019-06-03 src/searchnets/nets/alexnet.py
add 'allow_pickle=True' in call to np.load in alexnet.py

@@ -53,7 +53,7 @@ class AlexNet:
         self.weights_path = weights_path
         with open(self.weights_path, "rb") as weights_fp:
             # use item to get dictionary saved in a numpy array
-            self.weights_dict = np.load(weights_fp, encoding="latin1").item()
+            self.weights_dict = np.load(weights_fp, encoding="latin1", allow_pickle=True).item()
 
         self.output = None  # will be set during call to self.create() below
 

### 2019-06-03 src/searchnets/nets/layers.py
change arg name in layers.py to fix deprecation warning

@@ -20,4 +20,4 @@ def lrn(x, radius, alpha, beta, name, bias=1.0):
 
 def dropout(x, dropout_rate):
     """Create a dropout layer."""
-    return tf.nn.dropout(x, dropout_rate)
+    return tf.nn.dropout(x, rate=dropout_rate)

### 2019-06-03 src/searchnets/train.py
move calls to get_dataset into graph context
to fix warning and hopefully fix weird bug where process just gets killed in the middle of training

@@ -106,20 +106,6 @@ def train(gz_filename,
         if type(val_step) != int or patience < 1:
             raise TypeError('patience must be a positive integer')
 
-    print('loading training data')
-    data_dict = joblib.load(gz_filename)
-    train_ds = get_dataset(data_dict['x_train'], data_dict['y_train'], net_name, batch_size, shuffle=True)
-
-    if use_val:
-        try:
-            val_ds = get_dataset(data_dict['x_val'], data_dict['y_val'], net_name, batch_size, shuffle=False)
-        except KeyError:
-            raise KeyError(
-                f'use_val set to True but x_val not found in data file: {gz_filename}'
-            )
-    else:
-        val_ds = None
-
     if type(epochs_list) is int:
         epochs_list = [epochs_list]
     elif type(epochs_list) is list:
@@ -127,6 +113,9 @@ def train(gz_filename,
     else:
         raise TypeError("'EPOCHS' option in 'TRAIN' section of config.ini file parsed "
                         f"as invalid type: {type(epochs_list)}")
+
+    print('loading training data')
+    data_dict = joblib.load(gz_filename)
 
     np.random.seed(random_seed)  # for shuffling in batch_generator
     tf.random.set_random_seed(random_seed)
@@ -136,8 +125,6 @@ def train(gz_filename,
         # in training loop
         set_size_vec_train = data_dict['set_size_vec_train']
         set_sizes = np.unique(set_size_vec_train)
-        if val_ds is not None:
-            set_size_vec_val = data_dict['set_size_vec_val']
 
         acc_savepath = os.path.join(model_save_path,
                                     f'acc_by_epoch_by_set_size')
@@ -151,6 +138,20 @@ def train(gz_filename,
             graph = tf.Graph()
             with tf.Session(graph=graph) as sess:
                 # --------------- do a bunch of graph set-up stuff -----------------------------------------------------
+                # apparently it matters if we make the tf.data.Dataset in the same graph as the network
+                train_ds = get_dataset(data_dict['x_train'], data_dict['y_train'], net_name, batch_size, shuffle=True)
+
+                if use_val:
+                    try:
+                        val_ds = get_dataset(data_dict['x_val'], data_dict['y_val'], net_name, batch_size,
+                                             shuffle=False)
+                    except KeyError:
+                        raise KeyError(
+                            f'use_val set to True but x_val not found in data file: {gz_filename}'
+                        )
+                else:
[…]

### 2019-06-03 src/searchnets/test.py
move call to get_datasets into "with graph" context in test.py as in train.py

@@ -65,7 +65,6 @@ def test(gz_filename,
     """
     print('loading testing data')
     data_dict = joblib.load(gz_filename)
-    test_ds = get_dataset(data_dict['x_test'], data_dict['y_test'], net_name, batch_size, shuffle=False)
 
     set_sizes_by_stim_type = data_dict['set_sizes_by_stim_stype']
     set_sizes = []
@@ -89,6 +88,8 @@ def test(gz_filename,
             tf.reset_default_graph()
             graph = tf.Graph()
             with tf.Session(graph=graph) as sess:
+                test_ds = get_dataset(data_dict['x_test'], data_dict['y_test'], net_name, batch_size, shuffle=False)
+
                 x = tf.placeholder(tf.float32, (None,) + input_shape, name='x')
                 y = tf.placeholder(tf.int32, shape=[None], name='y')
                 rate = tf.placeholder_with_default(tf.constant(1.0, dtype=tf.float32), shape=(), name='dropout_rate')

### 2019-06-03 src/searchnets/test.py
remove unused tf.placeholder in searchnets/test.py

@@ -91,7 +91,6 @@ def test(gz_filename,
                 test_ds = get_dataset(data_dict['x_test'], data_dict['y_test'], net_name, batch_size, shuffle=False)
 
                 x = tf.placeholder(tf.float32, (None,) + input_shape, name='x')
-                y = tf.placeholder(tf.int32, shape=[None], name='y')
                 rate = tf.placeholder_with_default(tf.constant(1.0, dtype=tf.float32), shape=(), name='dropout_rate')
 
                 if net_name == 'alexnet':

### 2019-06-03 src/searchnets/test.py
remove total_loss from searchstims/train.py (copied from test by accident)

@@ -115,7 +115,6 @@ def test(gz_filename,
                 saver.restore(sess, ckpt_path)
 
                 total = int(np.ceil(len(data_dict['x_test']) / batch_size))
-                total_loss = 0.0
                 iterator = test_ds.make_one_shot_iterator()
                 next_element = iterator.get_next()
 

### 2019-06-04 src/searchnets/nets/layers.py
go back to 'keep_prob' arg in call to tf.nn.dropout
- because we're back with tf 1.12.0 again

@@ -20,4 +20,4 @@ def lrn(x, radius, alpha, beta, name, bias=1.0):
 
 def dropout(x, dropout_rate):
     """Create a dropout layer."""
-    return tf.nn.dropout(x, rate=dropout_rate)
+    return tf.nn.dropout(x, keep_prob=dropout_rate)

### 2019-06-04 src/searchnets/utils/tfdata.py
add shuffle_size args in tfdata.py, increase num_parallel_calls to 24

@@ -20,16 +20,16 @@ def preprocess(x, convert_bgr=True):
     return x
 
 
-def _generic_dataset(x, y, preprocess_func, batch_size, shuffle=True):
+def _generic_dataset(x, y, preprocess_func, batch_size, shuffle=False, shuffle_size=None):
     if type(x) == list:
         x = np.asarray(x)
     x_ds = tf.data.Dataset.from_tensor_slices(x)
-    x_ds = x_ds.map(load)
-    x_ds = x_ds.map(preprocess_func)
+    x_ds = x_ds.map(load, num_parallel_calls=24)
+    x_ds = x_ds.map(preprocess_func, num_parallel_calls=24)
     y_ds = tf.data.Dataset.from_tensor_slices(y)
     ds = tf.data.Dataset.zip((x_ds, y_ds))
     if shuffle:
-        ds = ds.shuffle(buffer_size=x.shape[-1])
+        ds = ds.shuffle(buffer_size=shuffle_size)
     ds = ds.batch(batch_size)
     ds = ds.prefetch(buffer_size=batch_size)
     return ds
@@ -41,8 +41,8 @@ def alexnet_preprocess():
     return anonymous
 
 
-def alexnet_dataset(x, y, batch_size, shuffle=True):
-    return _generic_dataset(x, y, alexnet_preprocess(), batch_size, shuffle)
+def alexnet_dataset(x, y, batch_size, shuffle=True, shuffle_size=None):
+    return _generic_dataset(x, y, alexnet_preprocess(), batch_size, shuffle, shuffle_size)
 
 
 def vgg16_preprocess():
@@ -51,12 +51,12 @@ def vgg16_preprocess():
     return anonymous
 
 
-def vgg16_dataset(x, y, batch_size, shuffle=True):
-    return _generic_dataset(x, y, vgg16_preprocess(), batch_size, shuffle)
+def vgg16_dataset(x, y, batch_size, shuffle=True, shuffle_size=None):
+    return _generic_dataset(x, y, vgg16_preprocess(), batch_size, shuffle, shuffle_size)
 
 
-def get_dataset(x, y, net_name, batch_size, shuffle=True):
+def get_dataset(x, y, net_name, batch_size, shuffle=True, shuffle_size=None):
     if net_name == 'alexnet':
-        return alexnet_dataset(x, y, batch_size, shuffle)
+        return alexnet_dataset(x, y, batch_size, shuffle, shuffle_size)
[…]

### 2019-06-04 src/searchnets/train.py
rewrite searchnets/train to use initializable iterator
so that all of data is not embedded in graph; fed in through placeholders instead.
Without doing this, loading/iterating through data is incredibly slow
(because it's embedded in graph?)

@@ -139,12 +139,15 @@ def train(gz_filename,
             with tf.Session(graph=graph) as sess:
                 # --------------- do a bunch of graph set-up stuff -----------------------------------------------------
                 # apparently it matters if we make the tf.data.Dataset in the same graph as the network
-                train_ds = get_dataset(data_dict['x_train'], data_dict['y_train'], net_name, batch_size, shuffle=True)
+                filenames_placeholder = tf.placeholder(tf.string, shape=[None])
+                labels_placeholder = tf.placeholder(tf.int64, shape=[None])
+                train_ds = get_dataset(filenames_placeholder, labels_placeholder, net_name, batch_size,
+                                       shuffle=True, shuffle_size=len(data_dict['x_train']))
 
                 if use_val:
                     try:
-                        val_ds = get_dataset(data_dict['x_val'], data_dict['y_val'], net_name, batch_size,
-                                             shuffle=False)
+                        val_ds = get_dataset(filenames_placeholder, labels_placeholder, net_name, batch_size,
+                                             shuffle=False, shuffle_size=len(data_dict['x_val']))
                     except KeyError:
                         raise KeyError(
                             f'use_val set to True but x_val not found in data file: {gz_filename}'
@@ -236,7 +239,9 @@ def train(gz_filename,
                 for epoch in range(epochs):
                     total = int(np.ceil(len(data_dict['x_train']) / batch_size))
                     total_loss = 0.0
-                    iterator = train_ds.make_one_shot_iterator()
+                    iterator = train_ds.make_initializable_iterator()
+                    sess.run(iterator.initializer, feed_dict={filenames_placeholder: data_dict['x_train'],
+                                                              labels_placeholder: data_dict['y_train']})
                     next_element = iterator.get_next()
 
                     pbar = tqdm(range(total))
@@ -259,9 +264,10 @@ def train(gz_filename,
                     if val_ds is not None:
                         if epoch % val_step == 0:
                             total = int(np.ceil(len(data_dict['x_val']) / batch_size))
-                            iterator = val_ds.make_one_shot_iterator()
+                            iterator = val_ds.make_initializable_iterator()
+                            sess.run(iterator.initializer, feed_dict={filenames_placeholder: data_dict['x_val'],
+                                                                      labels_placeholder: data_dict['y_val']})
                             next_element = iterator.get_next()
-
                             pbar = tqdm(range(total))
                             val_acc_this_epoch = []
                             for i in pbar:
@@ -301,7 +307,9 @@ def train(gz_filename,
[…]

### 2019-06-04 src/searchnets/test.py
rewrite searchnets/test to use initializable iterator

@@ -88,7 +88,10 @@ def test(gz_filename,
             tf.reset_default_graph()
             graph = tf.Graph()
             with tf.Session(graph=graph) as sess:
-                test_ds = get_dataset(data_dict['x_test'], data_dict['y_test'], net_name, batch_size, shuffle=False)
+                filenames_placeholder = tf.placeholder(tf.string, shape=[None])
+                labels_placeholder = tf.placeholder(tf.int64, shape=[None])
+                test_ds = get_dataset(filenames_placeholder, labels_placeholder, net_name, batch_size,
+                                      shuffle=True, shuffle_size=len(data_dict['x_test']))
 
                 x = tf.placeholder(tf.float32, (None,) + input_shape, name='x')
                 rate = tf.placeholder_with_default(tf.constant(1.0, dtype=tf.float32), shape=(), name='dropout_rate')
@@ -115,7 +118,9 @@ def test(gz_filename,
                 saver.restore(sess, ckpt_path)
 
                 total = int(np.ceil(len(data_dict['x_test']) / batch_size))
-                iterator = test_ds.make_one_shot_iterator()
+                iterator = test_ds.make_initializable_iterator()
+                sess.run(iterator.initializer, feed_dict={filenames_placeholder: data_dict['x_test'],
+                                                          labels_placeholder: data_dict['y_test']})
                 next_element = iterator.get_next()
 
                 y_test = []

### 2019-06-04 src/searchnets/data.py
fix typo in variable name in searchnets/data.py

@@ -138,7 +138,7 @@ def data(train_dir,
         stim_info_by_set_size = {int(k): v for k, v in stim_info_by_set_size.items()}
 
         set_sizes = [k for k in stim_info_by_set_size.keys()]
-        set_sizes_by_stim_stype[stim_type] = set_sizes
+        set_sizes_by_stim_type[stim_type] = set_sizes
 
         train_size_per_set_size = (train_size_per_stim_type / len(set_sizes)) / 2
         if train_size_per_set_size.is_integer():
@@ -268,7 +268,7 @@ def data(train_dir,
                      set_size_vec_train=set_size_vec_train,
                      set_size_vec_val=set_size_vec_val,
                      set_size_vec_test=set_size_vec_test,
-                     set_sizes_by_stim_stype=set_sizes_by_stim_stype,
+                     set_sizes_by_stim_type=set_sizes_by_stim_type,
                      stim_type_vec_train=stim_type_vec_train,
                      stim_type_vec_val=stim_type_vec_val,
                      stim_type_vec_test=stim_type_vec_test,

### 2019-06-04 docs/ccneuro2019/paper.tex
rewrite title and abstract again

@@
 
-\title{Do convolutional neural networks exhibit human-like attentional effects? 
-Conflicting results from a visual search task}
+\title{Convolutional neural networks performing a visual search task show attention-like 
+limits on accuracy when trained to generalize across multiple search stimuli}
 
@@
 \bf
-Convolutional neural networks (CNNs) trained to perform tasks such as image classification
-with human-like accuracy acquire representations similar to those in the primate 
-visual system. Do CNNs that are pre-trained to use such representations conversely exhibit 
-human-like behavior? We test this with a visual search task that has long been used to 
-investigate attention mechanisms. We first show that CNNs using weights trained for image 
-classification can exhibit human-like attentional effects when performing this visual 
-search task. However, we then demonstrate that these effects can be 
-eliminated, without changing the lower-level learned representations, by modifying other 
-parameters of training and the balance of classes in the training data. One explanation 
-for this finding is that we are only training networks to discriminate single types of 
-visual search stimuli. We further test CNNs trained to perform multiple tasks.
-We go on to show that manipulating the 
-discriminability of the target and distractors in the task can produce attentional 
-effects in spite of the changes we made to training. Based on these results, we conclude 
-that CNNs do not show attentional effects.
+What limits our ability to find what we are looking for in the cluttered noisy world we 
+see around us? A now classic form of visual search task has long been used to investigate 
+proposed mechanisms of selective attention. In spite of nearly half a century of studies 
+using this task, it remains unclear how best to relate different effects found using this 
+task to computations that take place in the visual system. A separate thread of research 
+has studied the visual system of humans and other primates using convolutional neural 
+networks (CNNs) as models. While CNNs have proven quite useful as models of the visual 
+system, it also remains unclear the extent to which their behavior mirrors that of 
+humans when performing vision-based tasks. Here we investigate whether the accuracy of 
+CNNs is affected by the same factors that affect human performance on the classic form of 
+visual search task used to investigate attention mechanisms. We first replicate and extend
+a previous study which did find that CNNs trained for image classification show human-like
+attentional limitations. Then we demonstrate that, by changing how we train the networks, 
+these effects can largely be eliminated. Lastly we show that accuracy is impaired when 
+single networks are trained to discriminate multiple types of visual search stimuli. Based
+on these findings, we suggest that the need to generalize across multiple tasks may be one 
+source of limitations on visual search for both CNNs and the primate visual system.
 }

### 2019-06-05 docs/ccneuro2019/paper.tex
revamp first section of intro on visual search

@@
 \section{Introduction}
-\subsection{Attention}
-Multiple converging lines of evidence seem to suggest that training neural networks to 
-perform tasks with human-like abilities causes them to acquire brain-like representations.
+\subsection{What are the factors that limit visual search?}
+What limits our ability to find what we are looking for in the cluttered noisy world we 
+see around us? One of the principle tasks that has been used to investigate this question
+is a classic form of visual search task, in which subjects view a stimulus on each trial, 
+made up of discrete sets of items on a solid-colored background, and report whether a 
+target is present among distractors [@wolfeVisualSearch1998]. 
+Most studies experimentally manipulate factors such as features of the targets and 
+distractors, in order to identify those factors limit visual search 
+[@ecksteinVisualSearchRetrospective2011]. Importantly, these factors that *limit* visual 
+search performance form the basis of seminal theories of selective visual attention. 
+As we review briefly below for context, there are essentially two families of selective 
+attention mechanisms. The results supporting both mechanisms remain unreconciled, in spite
+of nearly half a century of studies based on this task. The core question is: to what 
+extent can the limitations on visual search be attributed to some computation that is 
+required to bind features into items, and to what extent can those limitations be 
+explained by other aspects of computation, such as noisy internal representations and 
+decision-making criteria? Another way of stating this problem is to ask: if there 
+was some algorithm that could produce an ideal observer capable of maximally utilizing 
+the available information to carry out the task, would that ideal observer still be 
+subject to some ceiling of performance, due simply to the constraints of the task?
 
-- If brain-like representations are both necessary and sufficient for human-like 
-intelligence, then shouldn't the converse also be true? Shouldn't neural networks that 
-have acquired brain-like representations behave like humans when asked to perform related 
-tasks? We test this
-
-In these fields, theories of attention mechanisms are often built on 
-results produced by experimental manipulations that impair behavior.
-Perhaps the pre-eminent example of this approach to understanding attention is a 
-a long-established form of visual search task [@wolfeVisualSearch1998]. In this behavioral
-paradigm, subjects view a set of discrete items on a flat background and report whether a 
-target is present or absent among distractors. Classically, results from this paradigm 
-have been used to support different theories of \bf{selective covert visual attention}, 
-a term that requires some defintion. The term "selective" is simply supposed to mean 
-some mechanism that selects some visual inputs for further processing while necessarily 
-discarding others. The word "covert" in this context means any mechanism that occurs 
-"centrally", i.e. inside the brain, without overt orientation, i.e. without eye or head 
-movements.
-- here we do not
-
-Our goal here is not to resolve the debate about whether selective covert visual 
-attention is serial or parallel. Rather our goal is to better understand attention as 
-defined in cognitive science and reconcile that with the definition used in artificial 
-intelligence.
-
-\subsection{visual search}
-\textbf{set size effects}
+For context, we briefly review the two competing theoretical mechanisms of selective 
+attention, and describe how they depend in part on how the visual search task is 
+performed. The first family consists of serial mechanisms. One of the most influential 
+early incarnations is known as Feature Integration Theory (FIT) 
+[@treismanFeatureintegrationAttention1980]. Essentially, FIT posits that the visual system
+can process single "features", such as color or orientation, in parallel, but when 
+required to utilize multiple features, it uses some sort of serial computation, i.e. one 
+that takes time. Evidence for this theory came from visual search experiments where 
+subjects were shown stimuli until responding, and their reaction time was measured. 
+Plotting reaction time as a function of set size revealed 
+lines whose slope were near zero when a single feature distinguished the target from 
+distractors (e.g., "red" v. "green"). The slope of reaction time v. set size increased 
+sharply, though, when distinguishing the target from distractors required finding a 
+conjunction of features (such as "red vertical" v. "red horizontal or green vertical"). 
+However, subsequent studies showed that slopes could not be cleanly segregated in a way 
+that supported distinct parallel and serial mechanisms [@wolfeWhatCanMillion1998]. This 
+led to a more general model of visual search which seeks to account for results from 
+multiple tasks, but still features at its core a serial processing step
+[@wolfeGuidedSearchRevised1994].
+The second family of search mechanisms are parallel, and arose in large part in reaction 
+to the way the visual search task was carried out when investigating
+serial mechanisms. Early studies that measured reaction times left several factors 
+uncontrolled, such as target-distractor similarity [@duncanVisualSearchStimulus1989], 
+drops in acuity outside the fovea, eye movements, and effects resulting from visual 
+crowding [@bergenRapidDiscriminationVisual1983; @ecksteinLowerVisualSearch1998]. 
+Hence researchers designed versions of the visual search task that controlled for such 
+factors[@palmerPsychophysicsVisualSearch2000]. Crucially, they showed subjects the 
+stimulus only briefly, to prevent eye movements, and measured accuracy instead of reaction
+time. These researchers argued that set size-dependent decreases in accuracy could be 
+explained by totally parallel mechanisms of attention, where performance was impaired by 
+noise. Computational models of parallel mechanisms succesfully explained results from 
+single feature and conjunction search stimuli, in terms of distributions and 
+discrimination thresholds borrowed from signal detection theory that psychophysicists 
+often employ.
 

### 2019-06-05 docs/ccneuro2019/paper.tex
another pass through intro

@@
 Most studies experimentally manipulate factors such as features of the targets and 
-distractors, in order to identify those factors limit visual search 
-[@ecksteinVisualSearchRetrospective2011]. Importantly, these factors that *limit* visual 
-search performance form the basis of seminal theories of selective visual attention. 
-As we review briefly below for context, there are essentially two families of selective 
-attention mechanisms. The results supporting both mechanisms remain unreconciled, in spite
-of nearly half a century of studies based on this task. The core question is: to what 
-extent can the limitations on visual search be attributed to some computation that is 
-required to bind features into items, and to what extent can those limitations be 
-explained by other aspects of computation, such as noisy internal representations and 
-decision-making criteria? Another way of stating this problem is to ask: if there 
-was some algorithm that could produce an ideal observer capable of maximally utilizing 
-the available information to carry out the task, would that ideal observer still be 
-subject to some ceiling of performance, due simply to the constraints of the task?
+distractors in order to identify the factors limit visual search 
+[@ecksteinVisualSearchRetrospective2011; @wolfeFiveFactorsThat2017].
+Importantly, it is these factors that *limit* visual search performance that form the 
+basis of seminal theories of selective visual attention, although authors differ in the 
+extent to which they identify limitations on visual search as synonymous with selective 
+attention. To remain neutral, we will use the term "capacity limitations". 
+As we review briefly below for context, there are essentially two competing proposed 
+sources of capacity limitations [@palmerSignalDetectionEvidence2011]. 
+The results supporting both remain unreconciled, in spite of nearly half a century of 
+studies based on this task. The core question is: to 
+what extent can limitations on visual search be attributed to an attention-like 
+computation, e.g. binding features into items, and to what extent can those limitations be 
+attributed to other computations, e.g. a decision-making process subject to  
+noisy internal representations?
+To foreshadow our approach below, we suggest that another way of posing this question is 
+to ask: if some algorithm could produce a statistical model capable of learning from data 
+to classify visual search stimuli as "target present" or "target absent" with high 
+accuracy, would the accuracy still be subject to some ceiling, due simply to the 
+constraints of the task? This way of posing the question bears some similarity to a 
+well-established framework in vision research known as ideal observer analsyis 
+[@geislerIdealObserverAnalysis2003]. However, few ideal observer models have take the 
+form of "pixel-in, behavior-out", and the few that have do not generalize to many 
+different types of visual search stimuli, so their predictive power is limited 
+[@geislerModelsOvertAttention2011].
 
-For context, we briefly review the two competing theoretical mechanisms of selective 
-attention, and describe how they depend in part on how the visual search task is 
-performed. The first family consists of serial mechanisms. One of the most influential 
-early incarnations is known as Feature Integration Theory (FIT) 
-[@treismanFeatureintegrationAttention1980]. Essentially, FIT posits that the visual system
-can process single "features", such as color or orientation, in parallel, but when 
-required to utilize multiple features, it uses some sort of serial computation, i.e. one 
-that takes time. Evidence for this theory came from visual search experiments where 
-subjects were shown stimuli until responding, and their reaction time was measured. 
-Plotting reaction time as a function of set size revealed 
+Before we discuss how to address this gap in models, we briefly review the two competing 
+proposed sources of capacity limitations, and describe how they depend in part on the 
+way the visual search task is performed. Adopting the terms of other researchers, we will 
+call the first family "attention-limited", and the second "noise-limited" 
+[@palmerSignalDetectionEvidence2011]. Both attention-limited and noise-limited models 
+depend on what are known as *set-size effects* seem when using visual search stimuli, 
+which we will define in terms of how the visual search task is carried out. 
+Attention-limited models posit two-stage theories of visual processing, a first 
+*preattentive* stage which can process single features, such as color or orientation, 
+in parallel, and a second *attentive* stage that binds features to objects that processes 
+serially, which results in a bottleneck [@treismanFeatureintegrationAttention1980; 
+@wolfeGuidedSearchRevised1994]. Evidence for this theory came from visual search 
+experiments where subjects were shown stimuli until responding, and their reaction time 
+was measured. Plotting reaction time as a function of set size revealed 
 lines whose slope were near zero when a single feature distinguished the target from 
@@
 conjunction of features (such as "red vertical" v. "red horizontal or green vertical"). 
-However, subsequent studies showed that slopes could not be cleanly segregated in a way 
-that supported distinct parallel and serial mechanisms [@wolfeWhatCanMillion1998]. This 
-led to a more general model of visual search which seeks to account for results from 
-multiple tasks, but still features at its core a serial processing step
-[@wolfeGuidedSearchRevised1994].
-The second family of search mechanisms are parallel, and arose in large part in reaction 
+This sort of set size effect has been replicated hundreds of times, although it remains 
+unclear why some search stimuli seem to be processed effectively in parallel while others 
+face a bottleneck [@wolfeWhatCanMillion1998; @wolfeFiveFactorsThat2017].
+The second family of capacity limitations can be conceived of as completely parallel but 
+noisy processing. These models arose in large part in reaction 
 to the way the visual search task was carried out when investigating
@@
 Hence researchers designed versions of the visual search task that controlled for such 
-factors[@palmerPsychophysicsVisualSearch2000]. Crucially, they showed subjects the 
+factors [@palmerPsychophysicsVisualSearch2000]. Crucially, they showed subjects the 
 stimulus only briefly, to prevent eye movements, and measured accuracy instead of reaction
-time. These researchers argued that set size-dependent decreases in accuracy could be 
-explained by totally parallel mechanisms of attention, where performance was impaired by 
-noise. Computational models of parallel mechanisms succesfully explained results from 
-single feature and conjunction search stimuli, in terms of distributions and 
-discrimination thresholds borrowed from signal detection theory that psychophysicists 
-often employ.
+time. Computational models of parallel mechanisms succesfully explained results from 
+single feature and conjunction search stimuli. These models borrowed ideas from signal 
+detection theory that psychophysicists often employ, where sensory representations are
+imagined as distributions, and decisions are seen as thresholds on those distributions.
+What both attention-limited and noise-limited models have in common is that they are 
+both highly abstracted models of the visual system, in which low-level features pass 
+through a hierarchy until reaching a final stage consisting of a simple decision rule.
+
+\subsection{CNNs as models of the visual system?}
+Similarly, the architecture of convolutional neural networks (CNNs) now routinely used for
+computer vision tasks represents a highly abstracted view of the visual system.
+Many researchers have drawn parallels between the architecture of CNNs and the 
+architecture of the visual system in the brain [@kriegeskorteDeepNeuralNetworks2015], 
+in particular in humans and other primates where this system has been most thoroughly 
+studied. Like CNNs, the visual system has a hierarchical structure, and is thought to 
+function in part by performing transformations at each level of this hierarchy so that 
+high-dimensional, low-level features are mapped into low-dimensional abstract 
+representations. Several recent studies in the realm of cognitive neuroscience 
+have measured representations in more directly tested the similarities between CNNs and 
+the visual system, by comparing representations in both [@khaligh-razaviDeepSupervisedNot2014; @yaminsPerformanceoptimizedHierarchicalModels2014; @yaminsUsingGoaldrivenDeep2016]. Representations 
+can be easily extracted from CNNs because they are "transparent box" models--i.e. we have 
+access to the "activity" of the individual units from which these models are built when 
+we "show" the CNNs a stimulus.
+A natural question to ask, for someone interested in visual search and models of the visual 
+system, is how CNNs perform this task. If CNNs are excellent models of the visual system, 
+one might expect them to be subject to the same limitations on visual search. Alternatively, 
+if CNNs are simply function approximators, one might expect that they would be able to 
+very easily learn to map the exclusive-or function of "target present or target absent", 
+especially when targets and distractors are easily discriminated by single feature 
+dimensions such as color. While there have been previous studies of neural networks 
+performing visual search tasks, we are aware of only one that employed architectures 
+which have been used in studies of the visual system referenced above.
 

### 2019-06-05 src/searchnets/config/classes.py
add attributes to DataConfig
- add train_size_per_set_size, val_size_per_set_size, and test_set_per_set_size, that
allow specifying for each visual search stimulus set size what the number of samples
in the training set should be
- add shard_train and shard_size; shard_train is bool that if True means that training
data should be split into "shards", i.e. subsets that will each fit into memory for
loading onto GPU / shuffling with a tf.data.Dataset pipeline (they suggest "sharding" in
the docs). The way its done here is to randomly shuffle the whole set and then split into
shards with the numb

@@ -116,6 +116,23 @@ class TrainConfig:
     patience = attr.ib(validator=optional(is_pos_int), default=None)
 
 
+def is_list_of_pos_int(instance, attribute, value):
+    if type(value) != list:
+        raise ValueError(
+            f'type of {attribute.name} must be a list'
+        )
+
+    for ind, item in enumerate(value):
+        if type(item) != int:
+            raise ValueError(
+                f'all elements in {attribute.name} must be int but item at index {ind} was: {type(item)}'
+            )
+        if item < 1:
+            raise ValueError(
+                f'all elements in {attribute.name} must be positive integer, but item at index {ind} was: {item}'
+            )
+
+
 @attr.s
 class DataConfig:
     """class to represent [DATA] section of config.ini file
@@ -158,6 +175,12 @@ class DataConfig:
             if type(set_size) != int:
                 raise TypeError('all values in set_sizes should be int but '
                                 f'got type {type(set_size)} for element {ind}')
+
+    train_size_per_set_size = attr.ib(validator=optional(is_list_of_pos_int), default=None)
+    val_size_per_set_size = attr.ib(validator=optional(is_list_of_pos_int), default=None)
+    test_size_per_set_size = attr.ib(validator=optional(is_list_of_pos_int), default=None)
+    shard_train = attr.ib(validator=instance_of(bool), default=True)
+    shard_size = attr.ib(validator=optional(instance_of(int)), default=None)
 
 
 @attr.s

### 2019-06-05 src/searchnets/config/parse.py
add logic to parse_config
that deals with .ini file options corresponding to attributes just added to DataConfig

@@ -111,12 +111,41 @@ def parse_config(config_fname):
         set_sizes = None
     gz_filename = config['DATA']['GZ_FILENAME']
 
+    if config.has_option('DATA', 'TRAIN_SIZE_PER_SET_SIZE'):
+        train_size_per_set_size = ast.literal_eval(config['DATA']['TRAIN_SIZE_PER_SET_SIZE'])
+    else:
+        train_size_per_set_size = None
+    if config.has_option('DATA', 'VAL_SIZE_PER_SET_SIZE'):
+        val_size_per_set_size = ast.literal_eval(config['DATA']['VAL_SIZE_PER_SET_SIZE'])
+    else:
+        val_size_per_set_size = None
+    if config.has_option('DATA', 'TEST_SIZE_PER_SET_SIZE'):
+        test_size_per_set_size = ast.literal_eval(config['DATA']['TEST_SIZE_PER_SET_SIZE'])
+    else:
+        test_size_per_set_size = None
+
+    if config.has_option('DATA', 'SHARD_TRAIN'):
+        shard_train = bool(strtobool(config['DATA']['SHARD_TRAIN']))
+    else:
+        shard_train = False
+    if config.has_option('DATA', 'SHARD_SIZE'):
+        shard_size = int(config['DATA']['SHARD_SIZE'])
+    else:
+        if shard_train:
+            raise ValueError('shard_train set to True inf config.ini file but shard_size not specified')
+        shard_size = None
+
     data_config = DataConfig(train_dir,
                              train_size,
                              gz_filename,
                              val_size,
                              test_size,
-                             set_sizes)
+                             set_sizes,
+                             train_size_per_set_size,
+                             val_size_per_set_size,
+                             test_size_per_set_size,
+                             shard_train,
+                             shard_size)
 
     # ------------- unpack [TEST] section of config.ini file -----------------------------------------------------------
     test_results_save_path = config['TEST']['TEST_RESULTS_SAVE_PATH']

### 2019-06-05 src/searchnets/data.py
add new arguments and logic to searchnets.data
function now uses train/val/test_size_per_set_size when specified, and
shard_train + shard_size if shard_train is True

@@ -1,6 +1,7 @@
 import os
 import json
 from glob import glob
+from math import ceil
 import random
 
 import numpy as np
@@ -11,7 +12,12 @@ def data(train_dir,
          train_size,
          gz_filename,
          val_size=None,
-         test_size=None):
+         test_size=None,
+         train_size_per_set_size=None,
+         val_size_per_set_size=None,
+         test_size_per_set_size=None,
+         shard_train=False,
+         shard_size=None):
     """prepare training, validation, and test datasets.
     Saves a compressed Python dictionary in the path specified
     by the GZ_FILENAME option in the DATA section of a config.ini file.
@@ -128,9 +134,10 @@ def data(train_dir,
             raise TypeError('test_size_per_set_size is not a whole number, adjust '
                             'total number of samples, or number of set sizes.')
     else:
+        # "-1" means "use the remaining samples for the test set"
         test_size_per_stim_type = -1
 
-    set_sizes_by_stim_stype = {}
+    set_sizes_by_stim_type = {}
     # below, stim type will be visual search stimulus type from searchstims package
     for stim_type, stim_info_by_set_size in fnames_dict.items():
         # and this will be set sizes declared by user for this stimulus (could be diff't for each stimulus type).
@@ -140,33 +147,55 @@ def data(train_dir,
         set_sizes = [k for k in stim_info_by_set_size.keys()]
         set_sizes_by_stim_type[stim_type] = set_sizes
 
-        train_size_per_set_size = (train_size_per_stim_type / len(set_sizes)) / 2
-        if train_size_per_set_size.is_integer():
-            train_size_per_set_size = int(train_size_per_set_size)
-        else:
-            raise TypeError(f'train_size_per_set_size, {train_size_per_set_size}, is is not a whole number.\n'
-                            'It is calculated as: (train_size_per_stim_type / len(set_sizes)) / 2\n'
-                            '(2 is for target present or absent).\n'
-                            'Adjust total number of samples, or number of set sizes.')
-        if val_size:
-            val_size_per_set_size = (val_size_per_stim_type / len(set_sizes)) / 2
-            if val_size_per_set_size.is_integer():
-                val_size_per_set_size = int(val_size_per_set_size)
+        if train_size_per_set_size is None:
+            train_size_per_set_size = (train_size_per_stim_type / len(set_sizes)) / 2
+            if train_size_per_set_size.is_integer():
+                train_size_per_set_size = int(train_size_per_set_size)
[…]

### 2019-06-05 src/searchnets/__main__.py
__main__ passes new arguments to searchstims.data

@@ -21,7 +21,12 @@ def _call_data(config):
          train_size=config.data.train_size,
          val_size=config.data.val_size,
          test_size=config.data.test_size,
-         gz_filename=config.data.gz_filename)
+         gz_filename=config.data.gz_filename,
+         train_size_per_set_size=config.data.train_size_per_set_size,
+         val_size_per_set_size=config.data.val_size_per_set_size,
+         test_size_per_set_size=config.data.test_size_per_set_size,
+         shard_train=config.data.shard_train,
+         shard_size=config.data.shard_size)
 
 
 def _call_train(config):

### 2019-06-05 src/searchnets/train.py
when dataset is split into shards, searchnets.train can use them

@@ -117,13 +117,22 @@ def train(gz_filename,
     print('loading training data')
     data_dict = joblib.load(gz_filename)
 
+    if 'shard_train' in data_dict:
+        shard_train = data_dict['shard_train']
+    else:
+        shard_train = False
+
     np.random.seed(random_seed)  # for shuffling in batch_generator
     tf.random.set_random_seed(random_seed)
 
     if save_acc_by_set_size_by_epoch:
         # get vecs for computing accuracy by set size below
         # in training loop
-        set_size_vec_train = data_dict['set_size_vec_train']
+        if shard_train:
+            set_size_vec_train = np.concatenate(data_dict['set_size_vec_train'])
+        else:
+            set_size_vec_train = data_dict['set_size_vec_train']
+
         set_sizes = np.unique(set_size_vec_train)
 
         acc_savepath = os.path.join(model_save_path,
@@ -139,26 +148,41 @@ def train(gz_filename,
             with tf.Session(graph=graph) as sess:
                 # --------------- do a bunch of graph set-up stuff -----------------------------------------------------
                 # apparently it matters if we make the tf.data.Dataset in the same graph as the network
-                filenames_placeholder = tf.placeholder(tf.string, shape=[None])
-                labels_placeholder = tf.placeholder(tf.int64, shape=[None])
-                train_ds = get_dataset(filenames_placeholder, labels_placeholder, net_name, batch_size,
-                                       shuffle=True, shuffle_size=len(data_dict['x_train']))
+                filenames_placeholder_tr = tf.placeholder(tf.string, shape=[None])
+                labels_placeholder_tr = tf.placeholder(tf.int64, shape=[None])
+                if shard_train:
+                    shuffle_size = data_dict['shard_size']
+                else:
+                    shuffle_size = len(data_dict['x_train'])
+                train_ds = get_dataset(filenames_placeholder_tr, labels_placeholder_tr, net_name, batch_size,
+                                       shuffle=True, shuffle_size=shuffle_size)
 
                 if use_val:
-                    try:
-                        val_ds = get_dataset(filenames_placeholder, labels_placeholder, net_name, batch_size,
-                                             shuffle=False, shuffle_size=len(data_dict['x_val']))
-                    except KeyError:
+                    if 'x_val' not in data_dict:
                         raise KeyError(
                             f'use_val set to True but x_val not found in data file: {gz_filename}'
                         )
+
[…]

### 2019-06-05 src/searchnets/test.py
fix typo in searchnets/train.py, 'stim_stype' -> 'stim_type'

@@ -66,7 +66,7 @@ def test(gz_filename,
     print('loading testing data')
     data_dict = joblib.load(gz_filename)
 
-    set_sizes_by_stim_type = data_dict['set_sizes_by_stim_stype']
+    set_sizes_by_stim_type = data_dict['set_sizes_by_stim_type']
     set_sizes = []
     for stim_type, set_sizes_this_stim in set_sizes_by_stim_type.items():
         if set_sizes == []:

### 2019-06-05 src/searchnets/test.py
fix bug in searchnets/test.py, shuffle should be False for test dataset

@@ -91,7 +91,7 @@ def test(gz_filename,
                 filenames_placeholder = tf.placeholder(tf.string, shape=[None])
                 labels_placeholder = tf.placeholder(tf.int64, shape=[None])
                 test_ds = get_dataset(filenames_placeholder, labels_placeholder, net_name, batch_size,
-                                      shuffle=True, shuffle_size=len(data_dict['x_test']))
+                                      shuffle=False, shuffle_size=None)
 
                 x = tf.placeholder(tf.float32, (None,) + input_shape, name='x')
                 rate = tf.placeholder_with_default(tf.constant(1.0, dtype=tf.float32), shape=(), name='dropout_rate')

### 2019-06-05 src/searchnets/test.py
searchnets/test.py looks for best-val-acc checkpoint if last not present

@@ -114,8 +114,12 @@ def test(gz_filename,
                                         f'trained_{epochs}_epochs',
                                         f'net_number_{net_number}')
                 print(f'Loading model from {restore_path}')
-                ckpt_path = os.path.join(restore_path, f'{net_name}-model.ckpt-{epochs}')
-                saver.restore(sess, ckpt_path)
+                try:
+                    ckpt_path = os.path.join(restore_path, f'{net_name}-model.ckpt-{epochs}')
+                    saver.restore(sess, ckpt_path)
+                except ValueError:
+                    ckpt_path = os.path.join(restore_path, f'{net_name}-model-best-val-acc.ckpt-{epochs}')
+                    saver.restore(sess, ckpt_path)
 
                 total = int(np.ceil(len(data_dict['x_test']) / batch_size))
                 iterator = test_ds.make_initializable_iterator()

### 2019-06-06 docs/ccneuro2019/paper.tex
Updates from Overleaf

@@
 see around us? One of the principle tasks that has been used to investigate this question
-is a classic form of visual search task, in which subjects view a stimulus on each trial, 
+is a classic form of visual search task (Figure~\ref{fig1}), in which subjects view a stimulus on each trial, 
 made up of discrete sets of items on a solid-colored background, and report whether a 
@@
 stimulus only briefly, to prevent eye movements, and measured accuracy instead of reaction
-time. Computational models of parallel mechanisms succesfully explained results from 
+time. Computational models of parallel mechanisms successfully explained results from 
 single feature and conjunction search stimuli. These models borrowed ideas from signal 
@@
 
-\subsection{CNNs as models of the visual system?}
+\subsection{CNNs as models of the visual system}
 Similarly, the architecture of convolutional neural networks (CNNs) now routinely used for
@@
 \caption{A classic form of visual search task.} 
-\label{sample-figure}
+\label{fig1}
 \end{figure}

### 2019-06-06 docs/ccneuro2019/paper.tex
Updates from Overleaf

@@
 see around us? A now classic form of visual search task has long been used to investigate 
-proposed mechanisms of selective attention. In spite of nearly half a century of studies 
+this question. In spite of nearly half a century of studies 
 using this task, it remains unclear how best to relate different effects found using this 
@@
 CNNs is affected by the same factors that affect human performance on the classic form of 
-visual search task used to investigate attention mechanisms. We first replicate and extend
+a task used to investigate capacity limitations in visual search. We first replicate and extend
 a previous study which did find that CNNs trained for image classification show human-like
 attentional limitations. Then we demonstrate that, by changing how we train the networks, 
-these effects can largely be eliminated. Lastly we show that accuracy is impaired when 
+these effects can be almost completely eliminated. Lastly we show that accuracy is impaired when 
 single networks are trained to discriminate multiple types of visual search stimuli. Based
@@
 made up of discrete sets of items on a solid-colored background, and report whether a 
-target is present among distractors [@wolfeVisualSearch1998]. 
+target is present among distractors \cite{wolfeVisualSearch1998}.
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=\columnwidth]{fig1/fig1.eps}
+\end{center}
+\caption{A classic form of visual search task.} 
+\label{fig1}
+\end{figure}
+
 Most studies experimentally manipulate factors such as features of the targets and 
 distractors in order to identify the factors limit visual search 
-[@ecksteinVisualSearchRetrospective2011; @wolfeFiveFactorsThat2017].
+\cite{ecksteinVisualSearchRetrospective2011,wolfeFiveFactorsThat2017}.
 Importantly, it is these factors that *limit* visual search performance that form the 
@@
 As we review briefly below for context, there are essentially two competing proposed 
-sources of capacity limitations [@palmerSignalDetectionEvidence2011]. 
+sources of capacity limitations \cite{palmerSignalDetectionEvidence2011}. 
 The results supporting both remain unreconciled, in spite of nearly half a century of 
@@
 constraints of the task? This way of posing the question bears some similarity to a 
-well-established framework in vision research known as ideal observer analsyis 
-[@geislerIdealObserverAnalysis2003]. However, few ideal observer models have take the 
+well-established framework in vision research known as ideal observer analysis 
+\cite{geislerIdealObserverAnalysis2003}. However, few ideal observer models have take the 
 form of "pixel-in, behavior-out", and the few that have do not generalize to many 
 different types of visual search stimuli, so their predictive power is limited 
-[@geislerModelsOvertAttention2011].
+\cite{geislerModelsOvertAttention2011}.
 
@@
 call the first family "attention-limited", and the second "noise-limited" 
-[@palmerSignalDetectionEvidence2011]. Both attention-limited and noise-limited models 
+\cite{palmerSignalDetectionEvidence2011}. Both attention-limited and noise-limited models 
 depend on what are known as *set-size effects* seem when using visual search stimuli, 
@@
 in parallel, and a second *attentive* stage that binds features to objects that processes 
-serially, which results in a bottleneck [@treismanFeatureintegrationAttention1980; 
-@wolfeGuidedSearchRevised1994]. Evidence for this theory came from visual search 
+serially, which results in a bottleneck 
+\cite{treismanFeatureintegrationTheoryAttention1980,wolfeGuidedSearchAlternative1989,wolfeGuidedSearchRevised1994}. Evidence for this theory came from visual search 
 experiments where subjects were shown stimuli until responding, and their reaction time 
@@
 unclear why some search stimuli seem to be processed effectively in parallel while others 
-face a bottleneck [@wolfeWhatCanMillion1998; @wolfeFiveFactorsThat2017].
+face a bottleneck \cite{wolfeWhatCanMillion1998,wolfeFiveFactorsThat2017}.
 The second family of capacity limitations can be conceived of as completely parallel but 
@@
 drops in acuity outside the fovea, eye movements, and effects resulting from visual 
-crowding [@bergenRapidDiscriminationVisual1983; @ecksteinLowerVisualSearch1998]. 
+crowding \cite{bergenRapidDiscriminationVisual1983, ecksteinLowerVisualSearch1998}. 
 Hence researchers designed versions of the visual search task that controlled for such 
-factors [@palmerPsychophysicsVisualSearch2000]. Crucially, they showed subjects the 
+factors \cite{palmerPsychophysicsVisualSearch2000}. Crucially, they showed subjects the 
 stimulus only briefly, to prevent eye movements, and measured accuracy instead of reaction
@@
 Many researchers have drawn parallels between the architecture of CNNs and the 
-architecture of the visual system in the brain [@kriegeskorteDeepNeuralNetworks2015], 
+architecture of the visual system in the brain \cite{kriegeskorteDeepNeuralNetworks2015}, 
 in particular in humans and other primates where this system has been most thoroughly 
@@
 have measured representations in more directly tested the similarities between CNNs and 
-the visual system, by comparing representations in both [@khaligh-razaviDeepSupervisedNot2014; @yaminsPerformanceoptimizedHierarchicalModels2014; @yaminsUsingGoaldrivenDeep2016]. Representations 
+the visual system, by comparing representations in both 
+\cite{khaligh-razaviDeepSupervisedNot2014, yaminsUsingGoaldrivenDeep2016, yaminsPerformanceoptimizedHierarchicalModels2014}. Representations 
 can be easily extracted from CNNs because they are "transparent box" models--i.e. we have 
@@
 which have been used in studies of the visual system referenced above.
-
-\begin{figure}[ht]
-\begin{center}
-\includegraphics[width=\columnwidth]{fig1/fig1.eps}
-\end{center}
-\caption{A classic form of visual search task.} 
-\label{fig1}
-\end{figure}
 
@@
 
-\nocite{ChalnickBillman1988a}
-\nocite{Feigenbaum1963a}
-\nocite{Hill1983a}
-\nocite{OhlssonLangley1985a}
-% \nocite{Lewis1978a}
-\nocite{Matlock2001}
-\nocite{NewellSimon1972a}
-\nocite{ShragerLangley1990a}
-
 \bibliographystyle{apacite}
@@
 
-\bibliography{ccn_style}
+\bibliography{ccneuro2019}
 

### 2019-06-06 src/searchnets/config/classes.py
add stim_types attribute to DataConfig

@@ -133,6 +133,19 @@ def is_list_of_pos_int(instance, attribute, value):
             )
 
 
+def is_list_of_str(instance, attribute, value):
+    if type(value) != list:
+        raise ValueError(
+            f'type of {attribute.name} must be a list'
+        )
+
+    for ind, item in enumerate(value):
+        if type(item) != str:
+            raise ValueError(
+                f'all elements in {attribute.name} must be str but item at index {ind} was: {type(item)}'
+            )
+
+
 @attr.s
 class DataConfig:
     """class to represent [DATA] section of config.ini file
@@ -165,6 +178,7 @@ class DataConfig:
 
     train_size = attr.ib(validator=instance_of(int))
     gz_filename = attr.ib(validator=instance_of(str))
+    stim_types = attr.ib(validator=optional(is_list_of_str), default=None)
     val_size = attr.ib(validator=optional(instance_of(int)), default=None)
     test_size = attr.ib(validator=optional(instance_of(int)), default=None)
     set_sizes = attr.ib(validator=instance_of(list), default=None)
@@ -181,7 +195,6 @@ class DataConfig:
     test_size_per_set_size = attr.ib(validator=optional(is_list_of_pos_int), default=None)
     shard_train = attr.ib(validator=instance_of(bool), default=True)
     shard_size = attr.ib(validator=optional(instance_of(int)), default=None)
-
 
 @attr.s
 class TestConfig:

### 2019-06-06 src/searchnets/config/parse.py
add logic to parse_config to handle stim_types option

@@ -96,6 +96,12 @@ def parse_config(config_fname):
     # ------------- unpack [DATA] section of config.ini file -----------------------------------------------------------
     train_dir = config['DATA']['TRAIN_DIR']
     train_size = int(config['DATA']['TRAIN_SIZE'])
+
+    if config.has_option('DATA', 'STIM_TYPES'):
+        stim_types = ast.literal_eval(config['DATA']['STIM_TYPES'])
+    else:
+        stim_types = None
+
     if config.has_option('DATA', 'VALIDATION_SIZE'):
         val_size = int(config['DATA']['VALIDATION_SIZE'])
     else:
@@ -138,6 +144,7 @@ def parse_config(config_fname):
     data_config = DataConfig(train_dir,
                              train_size,
                              gz_filename,
+                             stim_types,
                              val_size,
                              test_size,
                              set_sizes,

### 2019-06-06 src/searchnets/__main__.py
__main__.py passes stim_types argument to searchnets.data

@@ -19,6 +19,7 @@ def _call_data(config):
     """helper function to call searchstims.data"""
     data(train_dir=config.data.train_dir,
          train_size=config.data.train_size,
+         stim_types=config.data.stim_types,
          val_size=config.data.val_size,
          test_size=config.data.test_size,
          gz_filename=config.data.gz_filename,

### 2019-06-06 src/searchnets/data.py
add ability to select search stimulus types to searchstims.data

@@ -11,6 +11,7 @@ import joblib
 def data(train_dir,
          train_size,
          gz_filename,
+         stim_types=None,
          val_size=None,
          test_size=None,
          train_size_per_set_size=None,
@@ -27,15 +28,24 @@ def data(train_dir,
     train_dir : str
         path to directory where training data is saved
     train_size : int
-        number of samples to put in training data set
+        number of samples to put in training data set.
+        Note that if there is more than on stimulus type in the source data set, then the number of stimuli from
+        each type will be train_size / number of types.
     gz_filename : str
         name of .gz file in which to save dataset
+    stim_types : list
+        of strings; specifies which visual search stimulus types to use when creating dataset. Strings must be keys in
+        .json file within train_dir. Default is None, in which case all types found in .json file will be used.
     val_size : int
         number of samples to put in validation data set.
         Default is None, in which case there will be no validation set.
+        Note that if there is more than on stimulus type in the source data set, then the number of stimuli from
+        each type will be val_size / number of types.
     test_size : int
         number of samples to put in test data set.
         Default is None, in which case all samples not used in training and validation sets are used for test set.
+        Note that if there is more than on stimulus type in the source data set, then the number of stimuli from
+        each type will be test_size / number of types.
 
     Returns
     -------
@@ -108,7 +118,20 @@ def data(train_dir,
     with open(fname_json) as f:
         fnames_dict = json.load(f)
 
-    num_stim_types = len(fnames_dict)  # number of keys in fnames_dict will be number of stim
+    if stim_types:
+        if type(stim_types) != list or not all([type(stim_type) == str for stim_type in stim_types]):
+            raise TypeError('stim_types must be a list of strings')
+        if not all([stim_type in fnames_dict for stim_type in stim_types]):
+            not_in_fnames_dict = [stim_type for stim_type in stim_types if stim_type not in fnames_dict]
+            raise ValueError(
+                f'the following stimulus types were not found in {fname_json}: {not_in_fnames_dict}'
+            )
+        # just keep the stimuli specified in stim_types
+        fnames_dict = {k: v for k, v in fnames_dict.items() if k in stim_types}
+    else:
+        stim_types = list(fnames_dict.keys())  # number of keys in fnames_dict will be number of stim
+    num_stim_types = len(stim_types)
+
     train_size_per_stim_type = train_size / num_stim_types
[…]

### 2019-06-06 src/bin/searchstims_multiple_stims_script.py
add script for making multiple visual search stimuli

@@ -1,7 +1,7 @@
 from pathlib import Path
 
 import searchstims.make
-from searchstims.stim_makers import RVvGVStimMaker, RVvRHGVStimMaker, Two_v_Five_StimMaker
+from searchstims.stim_makers import RVvGVStimMaker, RVvRHGVStimMaker, Two_v_Five_StimMaker, TLStimMaker, xoStimMaker, TStimMaker
 
 ALEXNET_SIZE = (227, 227)
 VGG16_SIZE = (224, 224)
@@ -11,7 +11,7 @@ ITEM_BBOX_SIZE = (30, 30)
 JITTER = 12
 
 for window_size in ALEXNET_SIZE, VGG16_SIZE:
-    keys = ['RVvGV', 'RVvRHGV', '2_v_5']
+    keys = ['RVvGV', 'RVvRHGV', '2_v_5', 'YT_v_BTYL', 'YT_v_BTBL', 'Bx_v_RxBo', 'Bx_v_RxRo', 'TvT']
     vals = [
         RVvGVStimMaker(target_color='red',
                        distractor_color='green',
@@ -35,7 +35,39 @@ for window_size in ALEXNET_SIZE, VGG16_SIZE:
                              item_bbox_size=ITEM_BBOX_SIZE,
                              jitter=JITTER,
                              target_number=2,
-                             distractor_number=5)
+                             distractor_number=5),
+        TLStimMaker(window_size=window_size,
+                    border_size=BORDER_SIZE,
+                    grid_size=GRID_SIZE,
+                    item_bbox_size=ITEM_BBOX_SIZE,
+                    jitter=JITTER),
+        TLStimMaker(distractor_L_color=(100, 149, 237),
+                    window_size=window_size,
+                    border_size=BORDER_SIZE,
+                    grid_size=GRID_SIZE,
+                    item_bbox_size=ITEM_BBOX_SIZE,
+                    jitter=JITTER),
+        xoStimMaker(target_x_color='blue',
+                    distractor_x_color='red',
+                    distractor_o_color='blue',
+                    window_size=window_size,
+                    border_size=BORDER_SIZE,
+                    grid_size=GRID_SIZE,
+                    item_bbox_size=ITEM_BBOX_SIZE,
+                    jitter=JITTER),
+        xoStimMaker(target_x_color='blue',
+                    distractor_x_color='red',
[…]

### 2019-06-06 src/bin/searchstims_train_script.py
add src/bin/searchstims_train_script.py

@@ -1,0 +1,68 @@
+from pathlib import Path
+
+import searchstims.make
+from searchstims.stim_makers import RVvGVStimMaker, RVvRHGVStimMaker, Two_v_Five_StimMaker
+
+ALEXNET_SIZE = (227, 227)
+VGG16_SIZE = (224, 224)
+BORDER_SIZE = (30, 30)
+GRID_SIZE = (5, 5)
+ITEM_BBOX_SIZE = (30, 30)
+JITTER = 12
+
+for window_size in ALEXNET_SIZE, VGG16_SIZE:
+    keys = ['RVvGV', 'RVvRHGV', '2_v_5']
+    vals = [
+        RVvGVStimMaker(target_color='red',
+                       distractor_color='green',
+                       window_size=window_size,
+                       border_size=BORDER_SIZE,
+                       grid_size=GRID_SIZE,
+                       item_bbox_size=ITEM_BBOX_SIZE,
+                       jitter=JITTER),
+        RVvRHGVStimMaker(target_color='red',
+                         distractor_color='green',
+                         window_size=window_size,
+                         border_size=BORDER_SIZE,
+                         grid_size=GRID_SIZE,
+                         item_bbox_size=ITEM_BBOX_SIZE,
+                         jitter=JITTER),
+        Two_v_Five_StimMaker(target_color='white',
+                             distractor_color='white',
+                             window_size=window_size,
+                             border_size=BORDER_SIZE,
+                             grid_size=GRID_SIZE,
+                             item_bbox_size=ITEM_BBOX_SIZE,
+                             jitter=JITTER,
+                             target_number=2,
+                             distractor_number=5)
+    ]
+    if window_size == ALEXNET_SIZE:
+        alexnet_zip = zip(keys, vals)
+    elif window_size == VGG16_SIZE:
+        vgg16_zip = zip(keys, vals)
+
+OUTPUT_DIR = Path('data/visual_search_stimuli')
+TARGET_PRESENT = [3600, 7200, 14400, 28800]
+TARGET_ABSENT = [3600, 7200, 14400, 28800]
+SET_SIZES = [1, 2, 4, 8]
+
+
+def main():
+    for cnn, zipped in zip(
+        ['alexnet', 'VGG16'],
+            [alexnet_zip, vgg16_zip]):
+        for key, val in zipped:
+            json_filename = f'{cnn}_train_{key}.json'
+            stim_dict = {key: val}
+            output_dir = OUTPUT_DIR.joinpath(f'{cnn}_train_{key}')
[…]

### 2019-06-06 src/searchnets/utils/general.py
add results_csv function to utils.general

@@ -1,6 +1,10 @@
 import tarfile
 import os
 
+import joblib
+import numpy as np
+import pandas as pd
+
 
 def make_targz(output_filename, source_dir):
     """make a .tar.gz file from a directory, i.e. a gzip-compressed archive"""
@@ -30,3 +34,51 @@ def targz_dirs(dir_names=('checkpoints', 'configs', 'data_prepd_for_nets', 'resu
         output_filename = dir_name + '.tar.gz'
         print(f'making {output_filename}')
         make_targz(output_filename=output_filename, source_dir=os.path.join(path, dir_name))
+
+
+HEADER = ['net_name', 'train_type', 'net_number', 'stimulus', 'target_condition', 'set_size', 'accuracy']
+
+
+def results_csv(results_dir,
+                data_prep_dir,
+                test_csv_path='./test.csv',
+                nets = ('alexnet', 'VGG16'),
+                train_types = ('finetune', 'train'),
+                stims = ('2_v_5', 'RVvGV', 'RVvRHGV'),
+                target_condition = ('present', 'absent')):
+    """make csv from results directory
+
+    creates Pandas dataframe from results that is then saved to a .csv file.
+    The resulting dataframe can be uesd with searchstims.plot.figures.acc_v_set_size_df
+    """
+    rows = []
+    for train_type in train_types:
+        for net in nets:
+            for stim in stims:
+                gz_fname = data_prep_dir.joinpath(f'{net}_{train_type}_{stim}_data.gz')
+                data_dict = joblib.load(gz_fname)
+                y_test = data_dict['y_test']
+                set_size_vec_test = data_dict['set_size_vec_test']
+                set_sizes = data_dict['set_sizes_by_stim_type'][stim]
+                results_path = list(results_dir.joinpath(f'{net}_{train_type}_{stim}').glob(f'test_{net}*.gz'))
+                assert len(results_path) == 1
+                results_dict = joblib.load(results_path[0])
+                for net_num in range(4):
+                    ppm = results_dict['predictions_per_model_dict']
+                    key = [key for key in ppm.keys() if f'net_number_{net_num}' in key][0]
+                    y_pred = ppm[key]
+                    for target_cond in target_condition:
+                        for set_size in set_sizes:
+                            if target_cond == 'present':
+                                inds_this_cond = np.where(
+                                    np.logical_and(y_test == 1, set_size_vec_test == set_size))
+                            elif target_cond == 'absent':
[…]

### 2019-06-06 src/searchnets/plot/figures.py
add acc_v_set_size_df function to plot.figures

@@ -91,6 +91,115 @@ def acc_v_set_size(results, set_sizes=(1, 2, 4, 8), ax=None,
         plt.savefig(save_as)
 
 
+MN_ACC_TARGET_PRESENT_LABEL = 'mean acc.\ntarget present'
+MN_ACC_TARGET_ABSENT_LABEL = 'mean acc.\ntarget absent'
+
+
+def acc_v_set_size_df(df, net_name, train_type, stimulus, ax=None,
+                      title=None, save_as=None, figsize=(10, 5),
+                      set_xlabel=False, set_ylabel=False, set_ylim=True,
+                      ylim=(0, 1.1), plot_mean=True, add_legend=False):
+    """plot accuracy as a function of visual search task set size
+    for models trained on a single task or dataset
+
+    Accepts a Pandas dataframe and column names that determine what to plot.
+    Dataframe is produces by searchstims.utils.general.results_csv function.
+
+    Parameters
+    ----------
+    df : pandas.Dataframe
+        path to results.gz file saved after measuring accuracy of trained networks
+        on test set of visual search stimuli
+    set_sizes : list
+        of int, set sizes of visual search stimuli. Default is [1, 2, 4, 8].
+    ax : matplotlib.Axis
+        axis on which to plot figure. Default is None, in which case a new figure with
+        a single axis is created for the plot.
+    title : str
+        string to use as title of figure. Default is None.
+    save_as : str
+        path to directory where figure should be saved. Default is None, in which
+        case figure is not saved.
+    figsize : tuple
+        (width, height) in inches. Default is (10, 5). Only used if ax is None and a new
+        figure is created.
+    set_xlabel : bool
+        if True, set the value of xlabel to "set size". Default is False.
+    set_ylabel : bool
+        if True, set the value of ylabel to "accuracy". Default is False.
+    set_ylim : bool
+        if True, set the y-axis limits to the value of ylim.
+    ylim : tuple
+        with two elements, limits for y-axis. Default is (0, 1.1).
+    plot_mean : bool
+        if True, find mean accuracy and plot as a separate solid line. Default is True.
+    add_legend : bool
+        if True, add legend to axis. Default is False.
+    task_name : str
+
+
+    Returns
+    -------
+    None
+    """
+    if ax is None:
[…]

### 2019-06-06 src/searchnets/plot/figures.py
add plt.rcParams section back to plot.figures
need to do futzy stuff like set font.size on axis so removed
call to seaborn.style and did it all by hand instead

@@ -4,12 +4,24 @@ from glob import glob
 
 import joblib
 import numpy as np
+import matplotlib as mpl
 import matplotlib.pyplot as plt
 import pandas as pd
 from scipy import stats
 import seaborn as sns
 
-sns.set_style("whitegrid")
+
+mpl.style.use('seaborn-dark')
+
+plt.rcParams['font.size'] = 18
+plt.rcParams['axes.labelsize'] = 18
+plt.rcParams['axes.labelweight'] = 'regular'
+plt.rcParams['axes.titlesize'] = 20
+plt.rcParams['xtick.labelsize'] = 16
+plt.rcParams['ytick.labelsize'] = 16
+plt.rcParams['legend.fontsize'] = 16
+plt.rcParams['figure.titlesize'] = 20
+plt.rcParams['axes.axisbelow'] = True
 
 
 def acc_v_set_size(results, set_sizes=(1, 2, 4, 8), ax=None,

### 2019-06-06 docs/ccneuro2019/paper.tex
Updates from Overleaf

@@
 
-\author{{\large \bf David Nicholson (dnicho4@emory.edu)} \\
-  A Department, 1234 Example Street\\
-A City, State 12345 A country
-  \AND {\large \bf Astrid Prinz (AnotherPerson@this.planet.edu)} \\
-  A Department, 1234 Example Street\\
-A City, State 12345 A country}
+\author{{\large \bf David A. Nicholson (dnicho4@emory.edu)} \\
+  \AND {\large \bf Astrid A. Prinz (astrid.prinz@emory.edu)} \\
+  Emory University Department Of Biology, 1510 Clifton Road, Atlanta, GA 30322}
 
@@
 \bf
-What limits our ability to find what we are looking for in the cluttered noisy world we 
-see around us? A now classic form of visual search task has long been used to investigate 
-this question. In spite of nearly half a century of studies 
-using this task, it remains unclear how best to relate different effects found using this 
-task to computations that take place in the visual system. A separate thread of research 
+What limits our ability to find what we are looking for in the cluttered noisy world? 
+To investigate this, cognitive scientists have long used a 
+visual search task in which subjects view a display of discrete items on each trial, 
+and report whether a target is present among distractors. 
+In spite of hundreds of studies, it remains unclear how to relate 
+effects found using the discrete item display search task to computations 
+in the visual system. A separate thread of research 
 has studied the visual system of humans and other primates using convolutional neural 
-networks (CNNs) as models. While CNNs have proven quite useful as models of the visual 
-system, it also remains unclear the extent to which their behavior mirrors that of 
-humans when performing vision-based tasks. Here we investigate whether the accuracy of 
-CNNs is affected by the same factors that affect human performance on the classic form of 
-a task used to investigate capacity limitations in visual search. We first replicate and extend
-a previous study which did find that CNNs trained for image classification show human-like
-attentional limitations. Then we demonstrate that, by changing how we train the networks, 
-these effects can be almost completely eliminated. Lastly we show that accuracy is impaired when 
-single networks are trained to discriminate multiple types of visual search stimuli. Based
-on these findings, we suggest that the need to generalize across multiple tasks may be one 
-source of limitations on visual search for both CNNs and the primate visual system.
+networks (CNNs) as models. CNNs have a hierarchical structure similar to 
+the visual system, and multiple lines of evidence suggest that training CNNs to 
+perform tasks such as image classification causes them to learn  
+representations similar to those used by the visual system.
+These studies raise the question of whether CNNs that have learned such 
+representations behave similarly to humans performing other vision-based tasks. 
+Here we address this by measuring the behavior of CNNs trained for image classification 
+while they perform the discrete item display search task.
+We first show how a fine-tuning approach often used to adapt pre-trained CNNs to new 
+tasks can produce models that show human-like limitations on this task. However we 
+then demonstrate that we can greatly reduce these effects by changing training,
+without changing the learned representations. Lastly we show that accuracy is 
+impaired when single networks are trained to discriminate multiple types of visual 
+search stimuli. Based on these findings, we suggest that the need to generalize across
+multiple tasks may be one source of limitations on visual search for both CNNs and 
+the primate visual system.
 }
@@
 \section{Introduction}
-\subsection{What are the factors that limit visual search?}
-What limits our ability to find what we are looking for in the cluttered noisy world we 
-see around us? One of the principle tasks that has been used to investigate this question
-is a classic form of visual search task (Figure~\ref{fig1}), in which subjects view a stimulus on each trial, 
-made up of discrete sets of items on a solid-colored background, and report whether a 
-target is present among distractors \cite{wolfeVisualSearch1998}.
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
 \begin{center}
-\includegraphics[width=\columnwidth]{fig1/fig1.eps}
+\includegraphics[width=\columnwidth]{fig1/fig1.png}
 \end{center}
-\caption{A classic form of visual search task.} 
+\caption{
+\textbf{The discrete item display search task}. On each trial, subjects view  
+ a display (example images, panel a) of discrete items on a flat background. Subjects 
+respond whether the target is present or absent (condition varies across rows 
+of panel a). Another condition that varies across trials is the set size, that is, 
+the total number of targets and distractors (varies across rows in a). Many studies 
+focus on set size effects. Typical effects are an increase in reaction time 
+or a decrease in accuracy as set size increases (depicted schematically in panel b, 
+redrawn from \cite{wolfeReactionTimeDistributions2010} and \cite{ecksteinLowerVisualSearch1998a}). 
+Effects vary based on the features that distinguish targets from distractors 
+(shown in columns). Details in introduction. Accuracy for spatial configuration-type 
+stimuli not shown in panel b because this has been less studied (but see 
+\cite{palmerSignalDetectionEvidence2011}).
+}
 \label{fig1}
@@
 
-Most studies experimentally manipulate factors such as features of the targets and 
-distractors in order to identify the factors limit visual search 
-\cite{ecksteinVisualSearchRetrospective2011,wolfeFiveFactorsThat2017}.
-Importantly, it is these factors that *limit* visual search performance that form the 
-basis of seminal theories of selective visual attention, although authors differ in the 
-extent to which they identify limitations on visual search as synonymous with selective 
-attention. To remain neutral, we will use the term "capacity limitations". 
-As we review briefly below for context, there are essentially two competing proposed 
-sources of capacity limitations \cite{palmerSignalDetectionEvidence2011}. 
-The results supporting both remain unreconciled, in spite of nearly half a century of 
-studies based on this task. The core question is: to 
-what extent can limitations on visual search be attributed to an attention-like 
+What limits our ability to find what we are looking for in the cluttered noisy world we 
+see around us? One of the principle tasks that has been used to investigate this question 
+is a visual search task (Figure~\ref{fig1}) \cite{wolfeVisualSearch1998}
+we will refer to as the \emph{discrete item display search} task. 
+Most studies using this task experimentally manipulate factors such as features of the targets and 
+distractors in order to identify those factors that limit visual search 
+\cite{ecksteinVisualSearchRetrospective2011,wolfeFiveFactorsThat2017}, known as 
+\emph{capacity limitations}. There are essentially two models of capacity limitations: 
+\emph{attention-limited} and \emph{noise-limited} models \cite{palmerSignalDetectionEvidence2011}. 
+Briefly, we review these and describe how they depend in part on the 
+way the visual search task is performed. Both models 
+depend on what are known as \emph{set-size effects} seen when using visual search stimuli, 
+depicted schematically in \ref{fig1}\textbf{b}).
+Attention-limited models posit two-stage theories of visual processing, a first 
+\emph{preattentive} stage which can process single features such as color or orientation 
+in parallel, and a second \emph{attentive} stage which requires a serial computation to 
+binds features, resulting in a bottleneck 
+\cite{treismanFeatureintegrationTheoryAttention1980,wolfeGuidedSearchAlternative1989,wolfeGuidedSearchRevised1994}. 
+Evidence for this theory came from visual search 
+experiments where subjects were shown stimuli until responding, and their reaction time 
+was measured. Plotting reaction time as a function of set size revealed 
+lines whose slope were near zero (\ref{fig1}\textbf{b}, top row, left plot) 
+when a single feature distinguished the target from 
+distractors (\ref{fig1}, panel a, left column). 
+The slope of reaction time versus set size increased though (\ref{fig1}\textbf{b}, top row, middle plot), 
+when distinguishing the target from distractors required finding a 
+conjunction of features (\ref{fig1}, panel a, middle column). Slopes became 
+even steeper (\ref{fig1}\textbf{b}, top row, right plot) 
+when targets and distractors shared features and could only be 
+distinguished by their spatial configuration (\ref{fig1}\textbf{a}, right column). 
+This sort of set size effect has been replicated hundreds of times, although it remains 
+unclear why some search stimuli are processed effectively in parallel while others 
+face a bottleneck \cite{wolfeWhatCanMillion1998,wolfeFiveFactorsThat2017}.
+The second family of capacity limitations conceives of visual search as completely parallel but 
+noisy. These models arose in large part in reaction 
+to the way the visual search task was carried out when investigating
+serial mechanisms. Early studies that measured reaction times left several factors 
+uncontrolled, such as target-distractor similarity \cite{duncanVisualSearchStimulus1989}, 
+drops in acuity outside the fovea, eye movements, and effects resulting from visual 
+crowding \cite{bergenRapidDiscriminationVisual1983,ecksteinLowerVisualSearch1998a}. 
+Hence researchers designed versions of the visual search task that controlled for such 
+factors. Crucially, they showed subjects the 
+stimulus only briefly, to prevent eye movements, and measured accuracy instead of reaction
+time (shown schematically in \ref{fig1}\textbf{b}, bottom row). 
+Computational models of parallel mechanisms, based on signal detection theory, 
+successfully explained results from feature and conjunction search stimuli 
+(\ref{fig1}, panel a, left and middle columns) 
+\cite{palmerPsychophysicsVisualSearch2000,ecksteinVisualSearchRetrospective2011}.
+
+In spite of nearly half a century of studies based on the discrete item display search task,
+this core question remains unresolved: to what extent can limitations be attributed to an attention-like 
 computation, e.g. binding features into items, and to what extent can those limitations be 
-attributed to other computations, e.g. a decision-making process subject to  
-noisy internal representations?
-To foreshadow our approach below, we suggest that another way of posing this question is 
-to ask: if some algorithm could produce a statistical model capable of learning from data 
-to classify visual search stimuli as "target present" or "target absent" with high 
-accuracy, would the accuracy still be subject to some ceiling, due simply to the 
-constraints of the task? This way of posing the question bears some similarity to a 
+attributed to other computations, e.g. a decision-making process subject to noisy internal representations?
+To foreshadow our approach, we suggest another way of posing this question: 
+if some algorithm could produce a statistical model capable of learning from data 
+to classify visual search stimuli as target present or absent with high 
+accuracy, would that accuracy still be subject to some ceiling, due simply to the 
+constraints of the task? This framing bears some similarity to a 
 well-established framework in vision research known as ideal observer analysis 
-\cite{geislerIdealObserverAnalysis2003}. However, few ideal observer models have take the 
-form of "pixel-in, behavior-out", and the few that have do not generalize to many 
+\cite{geislerIdealObserverAnalysis2003}. However, few ideal observer models have taken the 
+form of "pixel-in, behavior-out", and therefore do not generalize to many 
 different types of visual search stimuli, so their predictive power is limited 
@@
 
-Before we discuss how to address this gap in models, we briefly review the two competing 
-proposed sources of capacity limitations, and describe how they depend in part on the 
-way the visual search task is performed. Adopting the terms of other researchers, we will 
-call the first family "attention-limited", and the second "noise-limited" 
-\cite{palmerSignalDetectionEvidence2011}. Both attention-limited and noise-limited models 
-depend on what are known as *set-size effects* seem when using visual search stimuli, 
-which we will define in terms of how the visual search task is carried out. 
-Attention-limited models posit two-stage theories of visual processing, a first 
-*preattentive* stage which can process single features, such as color or orientation, 
-in parallel, and a second *attentive* stage that binds features to objects that processes 
-serially, which results in a bottleneck 
-\cite{treismanFeatureintegrationTheoryAttention1980,wolfeGuidedSearchAlternative1989,wolfeGuidedSearchRevised1994}. Evidence for this theory came from visual search 
-experiments where subjects were shown stimuli until responding, and their reaction time 
-was measured. Plotting reaction time as a function of set size revealed 
-lines whose slope were near zero when a single feature distinguished the target from 
-distractors (e.g., "red" v. "green"). The slope of reaction time v. set size increased 
-sharply, though, when distinguishing the target from distractors required finding a 
-conjunction of features (such as "red vertical" v. "red horizontal or green vertical"). 
-This sort of set size effect has been replicated hundreds of times, although it remains 
-unclear why some search stimuli seem to be processed effectively in parallel while others 
-face a bottleneck \cite{wolfeWhatCanMillion1998,wolfeFiveFactorsThat2017}.
-The second family of capacity limitations can be conceived of as completely parallel but 
-noisy processing. These models arose in large part in reaction 
-to the way the visual search task was carried out when investigating
-serial mechanisms. Early studies that measured reaction times left several factors 
-uncontrolled, such as target-distractor similarity [@duncanVisualSearchStimulus1989], 
-drops in acuity outside the fovea, eye movements, and effects resulting from visual 
-crowding \cite{bergenRapidDiscriminationVisual1983, ecksteinLowerVisualSearch1998}. 
-Hence researchers designed versions of the visual search task that controlled for such 
-factors \cite{palmerPsychophysicsVisualSearch2000}. Crucially, they showed subjects the 
-stimulus only briefly, to prevent eye movements, and measured accuracy instead of reaction
-time. Computational models of parallel mechanisms successfully explained results from 
-single feature and conjunction search stimuli. These models borrowed ideas from signal 
-detection theory that psychophysicists often employ, where sensory representations are
-imagined as distributions, and decisions are seen as thresholds on those distributions.
-What both attention-limited and noise-limited models have in common is that they are 
-both highly abstracted models of the visual system, in which low-level features pass 
-through a hierarchy until reaching a final stage consisting of a simple decision rule.
-
-\subsection{CNNs as models of the visual system}
+Both attention-limited and noise-limited models are highly abstracted models of the visual system,
+in which low-level features pass through a hierarchy until reaching a final stage consisting of
+a simple decision rule.
 Similarly, the architecture of convolutional neural networks (CNNs) now routinely used for
@@
 high-dimensional, low-level features are mapped into low-dimensional abstract 
-representations. Several recent studies in the realm of cognitive neuroscience 
-have measured representations in more directly tested the similarities between CNNs and 
-the visual system, by comparing representations in both 
-\cite{khaligh-razaviDeepSupervisedNot2014, yaminsUsingGoaldrivenDeep2016, yaminsPerformanceoptimizedHierarchicalModels2014}. Representations 
-can be easily extracted from CNNs because they are "transparent box" models--i.e. we have 
-access to the "activity" of the individual units from which these models are built when 
-we "show" the CNNs a stimulus.
-A natural question to ask, for someone interested in visual search and models of the visual 
-system, is how CNNs perform this task. If CNNs are excellent models of the visual system, 
+representations. Several recent studies have found that that, when optimized to perform 
+tasks such as image classification, CNNs learn representations similar to those 
+observed in the visual system
+\cite{khaligh-razaviDeepSupervisedNot2014, yaminsUsingGoaldrivenDeep2016, 
+yaminsPerformanceoptimizedHierarchicalModels2014}. 
+These studies raise the question of whether CNNs that have learned such 
+representations behave similarly to humans performing other vision-based tasks. 
+Here we address this by measuring the behavior of CNNs trained for image classification 
+while they perform the discrete item display search task.
+If CNNs function similarly to the visual system, 
 one might expect them to be subject to the same limitations on visual search. Alternatively, 
@@
 dimensions such as color. While there have been previous studies of neural networks 
-performing visual search tasks, we are aware of only one that employed architectures 
-which have been used in studies of the visual system referenced above.
+performing visual search tasks, we are aware of only one study \cite{poderCapacityLimitationsVisual2017} 
+that employed the sort of CNN architectures used in studies of the visual system referenced above.
+We replicate the methods from that study, and extend that author's results.
+In the interest of replicability, we have made the code and summary results \footnote{https://github.com/NickleDave/visual-search-nets} 
+available, and will release the raw data upon publication \footnote{at https://figshare.com/articles/visual-search-nets/7688840}.
 
@@
 
-\subsection{CNNs trained for image classification show set size effects}
+\subsection{Fine-tuning CNNs trained for image classification results in human-like limits on accuracy in the discrete item display search task}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=\columnwidth]{fig2/fig2.png}
+\end{center}
+\caption{
+\textbf{Human-like limits on accuracy of CNNs performing the discrete item display search task}
+As described in the text, when using weights pre-trained on ImageNet in convolutional layers, 
+and fine-tuning weights in fully-connected layers so AlexNet could perform the task, 
+this CNN showed decreases in accuracy as set size increased (\textbf{a}). This decrease was 
+smallest for feature search stimuli (left plot), intermediate for conjunction stimuli (middle plot), 
+and largest for spatial configuration stimuli (right plot). The same approach produced similar 
+results with the VGG16 architecture (\textbf{b}).
+} 
+\label{fig2}
+\end{figure}
+
+As referenced above, previous work suggests that optimizing CNNs to perform image 
+classification causes them to acquire representations which resemble those that can 
+be identified in the brain. We first tested whether the CNN architecture AlexNet 
+\cite{krizhevskyImagenetClassificationDeep2012}
+would be subject to human-like limits on the discrete item display search task, 
+when utilizing representations learned by optimizing the weights for image classification 
+on the ImageNet dataset \cite{dengImagenetLargescaleHierarchical2009}. 
+To do so, we used a \emph{fine-tuning} approach, in which we used pre-trained weights in earlier 
+convolutional layers of AlexNet, but randomly initialized weights in the later 
+fully-connected layers \cite{yosinskiHowTransferableAre2014}.
+More specifically, we replicated the training method described in
+\cite{poderCapacityLimitationsVisual2017}, where we trained each network that we tested with 
+6400 samples of \emph{one} of the visual search stimuli (e.g. feature search), and then 
+measured accuracy on a separate test set of 800 samples. Stimuli were generated with a 
+small Python package \footnote{https://github.com/NickleDave/searchstims} 
+which produced images the same size as the images used to train on 
+ImageNet, and were pre-processed in the same way as those images, 
+except that no re-sizing was done.
+This training method produced AlexNet models whose accuracy showed set-size dependence 
+similar to human subjects (Figure \ref{fig2}\textbf{a}). 
+For the three types of visual search stimuli that we used, there was a set size effect,
+where accuracy decreased as the set size increased. 
+This effect was smallest for feature search (left column), slightly 
+larger for conjunction search (middle column), and largest for spatial configuration 
+search. To test whether this effect was unique to AlexNet, we also used the same approach 
+with the VGG16 architecture, and produced similar results (Figure \ref{fig2}\textbf{b}). 
+In addition to finding set size effects that were qualitatively 
+similar to those seen in human subjects, we also noted that accuracy was always higher 
+for the "target absent" condition. We do not find reports of similar differences between 
+target present and target absent conditions in human subjects.
+
+\subsection{Changing training greatly reduces set size effects}
 
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{fig2/fig2.png}
+\includegraphics[width=\columnwidth]{fig3/fig3.png}
 \end{center}
-\caption{CNNs trained for image classification show set size effects.} 
-\label{sample-figure}
+\caption{
+\textbf{Changing training greatly reduces set size effects.}
+} 
+\label{fig3}
 \end{figure}
 
-\subsection{Changing hyperparameters and balancing the dataset largely 
-eliminates set size effects}
+Because accuracy of CNNs depends in part on training, it could be the case that 
+the results just described are an artifact of how we trained the networks. 
+To gain insight into how our results depended on training, 
+we plotted training histories where we measured accuracy on the training set 
+at each epoch \emph{separately for each set size in the visual search stimuli}.
+These plots revealed (1) that accuracy had not yet approached some asymptotic 
+value by the end of training, and (2) that there was an inverse relationship between 
+the set size of a visual search stimulus and the rate that its accuracy increased, 
+e.g. accuracy on set size 1 reached its highest value within a few epochs, while accuracy on 
+set size 8 never converged on an asymptotic value (figure \ref{fig3}, panel a).
+Because these plots showed accuracy had not converged when we stopped training, we 
+used random search to find an optimal learning rate (a key hyperparameter).
+We did find we were able to improve accuracy and decrease training time by 
+using a typical learning rate on the fully-connected layers, while simply 
+freezing the pre-trained weights in earlier convolutional layers.
+Based on the the observation that different set sizes converged at different rates, we 
+also augmented the number of samples for larger set sizes.
+It may seem counterintuitive to "unbalance" the dataset this way, until one considers 
+that there are many more combinations of displays of set size 8 then of set size 1. 
+For example, if items are located on a 5-by-5 grid, then for set size 1 there are 25-pick-1
+combinations, i.e. 25,  while for set size 8 there are 25-pick-8 combinations (approximately 100k).
+(Note that the library we used to generate images ensures that there are no repeats; for set size 1, 
+jitter is added to produce more than 25 possible displays; this is typically done in experiments 
+with human subjects, and acts as a form of data augmentation for neural networks.)
+We simply multiplied the number of samples by the set size, since scaling 
+by the number of possible combinations would have produced prohibitively large datasets.
+After making these changes to the learning rate and the statistics of the dataset,
 
-Because we defined set size with effects in terms of accuracy, it could 
-be the case that our results do not arise because of a similarity between 
-how CNNs and brains process images. Instead they could be an artifact of 
-how we trained networks. As discussed below, we see this as a general issue 
-for studies that compare cognitive science and artificial intelligence, and 
-so here we report results that might otherwise be considered methodological 
-troubleshooting.
-To determine whether our results were an artifact of training, we considered aspects of 
-training that can impair accuracy: the amount of training data, the hyperparameters 
-used to train the network, and the statistics of the dataset. As shown in 
-figure two, we found that all three factors contributed to the effect we saw.
-We generated learning curves where we plotted accuracy on the training and 
-test set as a function of the training set size. These learning curves revealed 
-that the original size we chose for the training set was far from where 
-accuracy on training and test set converged. This result indicates that the 
-network can improve accuracy with increased training data. To gain insight 
-about how accuracy depended on the learning rate, a key hyperparameter, and 
-the balance of the data set, we plotted accuracy on the training set 
-\emph{separately for each set size in the visual search stimuli}. As shown in 
-figure 2b, these plots revealed (1) that accuracy had not yet reached some asymptotic 
-value, and (2) that there was an inverse relationship between the set size of 
-a visual search stimulus and the rate that its accuracy increased.
-To test whether we could improve the learning rate, we used random search, and 
-did find we were able to improve accuracy and decrease training time by 
-abandoning the fine-tuning approach and instead using a typical learning rate 
-on the fully-connected layers (and simply freezing the pre-trained weights in other layers).
-
-Use standard APA citation format. Citations within the text should
-include the author's last name and year. If the authors' names are
-included in the sentence, place only the year in parentheses, as in
-\citeA{NewellSimon1972a}, but otherwise place the entire reference in
-parentheses with the authors and year separated by a comma
-\cite{NewellSimon1972a}. List multiple references alphabetically and
-separate them by semicolons
-\cite{ChalnickBillman1988a,NewellSimon1972a}. Use the
-``et~al.'' construction only after listing all the authors to a
-publication in an earlier reference and for citations with four or
-more authors.
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=\columnwidth]{fig4/fig4.png}
+\end{center}
+\caption{
+\textbf{Human-like limits on accuracy of CNNs performing the discrete item display search task}
+As described in the text, when using weights pre-trained on ImageNet in convolutional layers, 
+and fine-tuning weights in fully-connected layers so AlexNet could perform the task, 
+this CNN showed decreases in accuracy as set size increased (\textbf{a}). This decrease was 
+smallest for feature search stimuli (left plot), intermediate for conjunction stimuli (middle plot), 
+and largest for spatial configuration stimuli (right plot). The same approach produced similar 
+results with the VGG16 architecture (\textbf{b}).
+} 
+\label{fig2}
+\end{figure}
 
@@
 - problem for any 
+we see this as a general issue 
+for studies that compare cognitive science and artificial intelligence, and 
+so here we report results that might otherwise be considered methodological 
+troubleshooting.
 
@@
 
-\subsection{Footnotes}
-
-Indicate footnotes with a number\footnote{Sample of the first
-footnote.} in the text. Place the footnotes in 9~point type at the
-bottom of the column on which they appear. Precede the footnote block
-with a horizontal rule.\footnote{Sample of the second footnote.}
-
-\subsection{Tables}
-
-Number tables consecutively. Place the table number and title (in
-10~point) above the table with one line space above the caption and
-one line space below it, as in Table~\ref{sample-table}. You may float
-tables to the top or bottom of a column, or set wide tables across
-both columns.
-
-\begin{table}[!ht]
-\begin{center} 
-\caption{Sample table title.} 
-\label{sample-table} 
-\vskip 0.12in
-\begin{tabular}{ll} 
-\hline
-Error type    &  Example \\
-\hline
-Take smaller        &   63 - 44 = 21 \\
-Always borrow~~~~   &   96 - 42 = 34 \\
-0 - N = N           &   70 - 47 = 37 \\
-0 - N = 0           &   70 - 47 = 30 \\
-\hline
-\end{tabular} 
-\end{center} 
-\end{table}
-
-\subsection{Figures}
-
-Make sure that the artwork can be printed well (e.g. dark colors) and that 
-the figures make understanding the paper easy.
- Number figures sequentially, placing the figure
-number and caption, in 10~point, after the figure with one line space
-above the caption and one line space below it, as in
-Figure~\ref{sample-figure}. If necessary, leave extra white space at
-the bottom of the page to avoid splitting the figure and figure
-caption. You may float figures to the top or bottom of a column, or
-set wide figures across both columns.
-
-\begin{figure}[ht]
-\begin{center}
-\includegraphics{fig1}
-\end{center}
-\caption{This is a figure.} 
-\label{sample-figure}
-\end{figure}
-
 \section{Acknowledgments}
-
-Place acknowledgments (including funding information) in a section at
-the end of the paper.
-
-\section{References Instructions}
-
-Follow the APA Publication Manual for citation format, both within the
-text and in the reference list, with the following exceptions: (a) do
-not cite the page numbers of any book, including chapters in edited
-volumes; (b) use the same format for unpublished references as for
-published ones. Alphabetize references by the surnames of the authors,
-with single author entries preceding multiple author entries. Order
-references by the same authors by the year of publication, with the
-earliest first.
-
-Use a first level section heading, ``{\bf References}'', as shown
-below. Use a hanging indent style, with the first line of the
-reference flush against the left margin and subsequent lines indented
-by 1/8~inch. Below are example references for a conference paper, book
-chapter, journal article, dissertation, book, technical report, and
-edited volume, respectively.
+Research funded by the Lifelong Learning Machines program, 
+DARPA/Microsystems Technology Office, DARPA cooperative agreement HR0011-18-2-0019.
+David Nicholson was partially supported by the 2017 William K. and Katherine W. Estes Fund
+to F. Pestilli, R. Goldstone and L. Smith, Indiana University Bloomington.
+Thank you to Zsolt Kira and Yen-Chang Hsu for feedback on results,
+and for suggesting changes to training methods.
+Thank you also to Constantine Dovrolis and Sarah Pallas for feedback on earlier versions 
+of this work.
 

### 2019-06-08 src/searchnets/plot/figures.py
change plotting parameters in acc_v_set_size_df
- add zorders and alphas
- change linewidth parameters

@@ -181,16 +181,18 @@ def acc_v_set_size_df(df, net_name, train_type, stimulus, ax=None,
                 assert np.array_equal(set_sizes, set_size)
 
     for arr_acc_present, arr_acc_absent in zip(acc_target_present, acc_target_absent):
-        ax.plot(set_sizes, arr_acc_present, color='violet', linestyle='--', marker='o', label=None)
-        ax.plot(set_sizes, arr_acc_absent, color='lightgreen', linestyle='--', marker='o', label=None)
+        ax.plot(set_sizes, arr_acc_present, color='violet', linewidth=2, linestyle='--',
+                marker='o', zorder=1, alpha=0.85, label=None)
+        ax.plot(set_sizes, arr_acc_absent, color='lightgreen', linewidth=2, linestyle='--',
+                marker='o', zorder=1, alpha=0.85, label=None)
 
     if plot_mean:
         mn_acc_present = np.asarray(acc_target_present).mean(axis=0)
-        mn_acc_present_line, = ax.plot(set_sizes, mn_acc_present, color='magenta', linewidth=2,
-                                       label=MN_ACC_TARGET_PRESENT_LABEL)
+        mn_acc_present_line, = ax.plot(set_sizes, mn_acc_present, color='magenta', linewidth=4,
+                                       zorder=0, label=MN_ACC_TARGET_PRESENT_LABEL)
         mn_acc_absent = np.asarray(acc_target_absent).mean(axis=0)
-        mn_acc_absent_line, = ax.plot(set_sizes, mn_acc_absent, color='lawngreen', linewidth=2,
-                                      label=MN_ACC_TARGET_ABSENT_LABEL)
+        mn_acc_absent_line, = ax.plot(set_sizes, mn_acc_absent, color='lawngreen', linewidth=4,
+                                      zorder=0, label=MN_ACC_TARGET_ABSENT_LABEL)
 
     ax.set_xticks(set_sizes)
 

### 2019-06-08 docs/ccneuro2019/paper.tex
push submitted version of paper from Overleaf

@@
 What limits our ability to find what we are looking for in the cluttered noisy world?
-To investigate this, cognitive scientists have long used a 
-visual search task in which subjects view a display of discrete items on each trial, 
-and report whether a target is present among distractors. 
-In spite of hundreds of studies, it remains unclear how to relate 
-effects found using the discrete item display search task to computations 
-in the visual system. A separate thread of research 
-has studied the visual system of humans and other primates using convolutional neural 
-networks (CNNs) as models. CNNs have a hierarchical structure similar to 
-the visual system, and multiple lines of evidence suggest that training CNNs to 
-perform tasks such as image classification causes them to learn  
-representations similar to those used by the visual system.
-These studies raise the question of whether CNNs that have learned such 
-representations behave similarly to humans performing other vision-based tasks. 
-Here we address this by measuring the behavior of CNNs trained for image classification 
-while they perform the discrete item display search task.
-We first show how a fine-tuning approach often used to adapt pre-trained CNNs to new 
-tasks can produce models that show human-like limitations on this task. However we 
-then demonstrate that we can greatly reduce these effects by changing training,
-without changing the learned representations. Lastly we show that accuracy is 
-impaired when single networks are trained to discriminate multiple types of visual 
-search stimuli. Based on these findings, we suggest that the need to generalize across
-multiple tasks may be one source of limitations on visual search for both CNNs and 
-the primate visual system.
+To investigate this, cognitive scientists have long used visual search.
+In spite of hundreds of studies, it remains unclear how to relate effects found using the discrete item display search task to computations in the visual system.
+A separate thread of research has studied the visual system of humans and other primates using convolutional neural networks (CNNs) as models.
+Multiple lines of evidence suggest that training CNNs to perform tasks such as image classification causes them to learn  representations similar to those used by the visual system.
+These studies raise the question of whether CNNs that have learned such representations behave similarly to humans performing other vision-based tasks.
+Here we address this by measuring the behavior of CNNs trained for image classification while they perform the discrete item display search task.
+We first show how a fine-tuning approach often used to adapt pre-trained CNNs to new tasks can produce models that show human-like limitations on this task.
+However we then demonstrate that we can greatly reduce these effects by changing training,without changing the learned representations.
+Lastly we show that accuracy is not impaired when single networks are trained to discriminate multiple types of visual search stimuli.
+Based on these findings, we suggest that CNNs are not necessarily subject to the same limitations as the primate visual system.
 }
@@
 \textbf{The discrete item display search task}. On each trial, subjects view  
- a display (example images, panel a) of discrete items on a flat background. Subjects 
+ a display (example images, \textbf{a}) of discrete items on a flat background. Subjects 
 respond whether the target is present or absent (condition varies across rows 
-of panel a). Another condition that varies across trials is the set size, that is, 
-the total number of targets and distractors (varies across rows in a). Many studies 
+of \textbf{a}). Another condition that varies across trials is the set size, that is, 
+the total number of targets and distractors (varies across rows in \textbf{a}). Many studies 
 focus on set size effects. Typical effects are an increase in reaction time 
-or a decrease in accuracy as set size increases (depicted schematically in panel b, 
+or a decrease in accuracy as set size increases (depicted schematically in \textbf{b}, 
 redrawn from \cite{wolfeReactionTimeDistributions2010} and \cite{ecksteinLowerVisualSearch1998a}). 
@@
 (shown in columns). Details in introduction. Accuracy for spatial configuration-type 
-stimuli not shown in panel b because this has been less studied (but see 
+stimuli not shown in \textbf{b} because this has been less studied (but see 
 \cite{palmerSignalDetectionEvidence2011}).
@@
 when a single feature distinguished the target from 
-distractors (\ref{fig1}, panel a, left column). 
+distractors (\ref{fig1}, \textbf{a}, left column). 
 The slope of reaction time versus set size increased though (\ref{fig1}\textbf{b}, top row, middle plot), 
 when distinguishing the target from distractors required finding a 
-conjunction of features (\ref{fig1}, panel a, middle column). Slopes became 
+conjunction of features (\ref{fig1}\textbf{a}, middle column). Slopes became 
 even steeper (\ref{fig1}\textbf{b}, top row, right plot) 
@@
 drops in acuity outside the fovea, eye movements, and effects resulting from visual 
-crowding \cite{bergenRapidDiscriminationVisual1983,ecksteinLowerVisualSearch1998a}. 
+crowding \cite{ecksteinLowerVisualSearch1998a}. 
 Hence researchers designed versions of the visual search task that controlled for such 
@@
 successfully explained results from feature and conjunction search stimuli 
-(\ref{fig1}, panel a, left and middle columns) 
+(\ref{fig1}\textbf{a}, left and middle columns) 
 \cite{palmerPsychophysicsVisualSearch2000,ecksteinVisualSearchRetrospective2011}.
@@
 observed in the visual system
-\cite{khaligh-razaviDeepSupervisedNot2014, yaminsUsingGoaldrivenDeep2016, 
-yaminsPerformanceoptimizedHierarchicalModels2014}. 
+\cite{yaminsUsingGoaldrivenDeep2016, yaminsPerformanceoptimizedHierarchicalModels2014}. 
 These studies raise the question of whether CNNs that have learned such 
@@
 while they perform the discrete item display search task.
-If CNNs function similarly to the visual system, 
-one might expect them to be subject to the same limitations on visual search. Alternatively, 
-if CNNs are simply function approximators, one might expect that they would be able to 
-very easily learn to map the exclusive-or function of "target present or target absent", 
-especially when targets and distractors are easily discriminated by single feature 
-dimensions such as color. While there have been previous studies of neural networks 
+While there have been previous studies of neural networks 
 performing visual search tasks, we are aware of only one study \cite{poderCapacityLimitationsVisual2017} 
@@
 \section{Results}
-
-\subsection{Fine-tuning CNNs trained for image classification results in human-like limits on accuracy in the discrete item display search task}
 
@@
 
-\subsection{Changing training greatly reduces set size effects}
-
 \begin{figure}[ht]
@@
 \textbf{Changing training greatly reduces set size effects.}
+Training histories showed that the accuracy of models trained 
+with the fine tuning approach did not converge
+on some asymptotic value, and 
+varied depending on the set size of the search stimuli (\textbf{a}, left plot). 
+Increasing the learning rate and including more examples of stimuli with 
+larger set sizes greatly sped up convergence (\textbf{a}, right plot).
+AlexNet models trained with this higher learning rate and larger training set
+showed reduced set size effects (\textbf{b}).
 } 
@@
 e.g. accuracy on set size 1 reached its highest value within a few epochs, while accuracy on 
-set size 8 never converged on an asymptotic value (figure \ref{fig3}, panel a).
+set size 8 never converged on an asymptotic value (figure \ref{fig3}\textbf{a}).
 Because these plots showed accuracy had not converged when we stopped training, we 
-used random search to find an optimal learning rate (a key hyperparameter).
+used random search to find an optimal learning rate, a key hyperparameter.
 We did find we were able to improve accuracy and decrease training time by 
@@
 After making these changes to the learning rate and the statistics of the dataset, 
+we saw that the set size effects were greatly reduced \ref{fig3}\textbf{b}. This indicates 
+that these effects are due in least at part to how we trained the networks.
+
+Because neural networks, including CNNs, act as function approximators, it could also be the case 
+that they were able to perform the discrete item display search task with relatively high 
+accuracy simply by learning an exclusive-or function for the single search stimulus that we trained 
+them to classify. A more rigorous test would be to train single networks to classify multiple 
+stimuli, in the same way that humans do in discrete item display search experiments. 
+As a final test, we trained two instances of Alexnet on very large datasets containing 
+all three stimuli used in this study \ref{fig4}. We found similar accuracy as shown in \ref{fig3}, 
+even when single networks were challenged to perform this task with multiple stimuli.
 
@@
 \caption{
-\textbf{Human-like limits on accuracy of CNNs performing the discrete item display search task}
-As described in the text, when using weights pre-trained on ImageNet in convolutional layers, 
-and fine-tuning weights in fully-connected layers so AlexNet could perform the task, 
-this CNN showed decreases in accuracy as set size increased (\textbf{a}). This decrease was 
-smallest for feature search stimuli (left plot), intermediate for conjunction stimuli (middle plot), 
-and largest for spatial configuration stimuli (right plot). The same approach produced similar 
-results with the VGG16 architecture (\textbf{b}).
+\textbf{Training single networks on multiple stimuli does not impair accuracy}
+Two instances of AlexNet trained on datasets containing all three types of visual search 
+stimuli used in this study still attained high accuracy.
 } 
-\label{fig2}
+\label{fig4}
 \end{figure}
@@
 \section{Discussion}
-
-\subsection{Implications for comparative studies of neuroscience and 
-artificial intelligence}
-- problem for any 
-we see this as a general issue 
-for studies that compare cognitive science and artificial intelligence, and 
-so here we report results that might otherwise be considered methodological 
-troubleshooting.
-
-- e.g. Bengio study of gestalt fx
-
-\subsection{Implications for the study of attention in the brain}
-- if the task is possible to perform perfectly, who needs attention?
-- brain optimized for something else besides classifying static images
-- in this sense, consistent with findings that recurrent
+We tested whether CNNs using representations learned from image classification tasks would be subject 
+to human-like capacity limitations when performing the discrete item display search task. While we 
+did not find that CNNs can always perform the discrete item display search task with perfect accuracy, 
+we did show that these models are not necessarily limited by the same factors as the primate visual system.
 

### 2019-06-08 data/runall_alexnet_train.sh
add data/runall_alexnet_train.sh

@@ -1,0 +1,6 @@
+#!/usr/bin/env bash
+
+# prepare data sets for training neural nets, then train, then test accuracy of trained nets
+searchnets all ./configs/searchnets_alexnet_train_RVvGV.ini
+searchnets all ./configs/searchnets_alexnet_train_RVvRHGV.ini
+searchnets all ./configs/searchnets_alexnet_train_2_v_5.ini

### 2019-06-08 src/searchnets/utils/figshare_urls.py
fix URLs to neural net weights on figshare

@@ -1,2 +1,2 @@
-ALEXNET_WEIGHTS_URL = 'https://ndownloader.figshare.com/files/14299136'
-VGG16_WEIGHTS_URL = 'https://ndownloader.figshare.com/files/14449532'
+ALEXNET_WEIGHTS_URL = 'https://ndownloader.figshare.com/files/15375197'
+VGG16_WEIGHTS_URL = 'https://ndownloader.figshare.com/files/15375212'

### 2019-06-08 src/searchnets/test.py
change how searchnets/test.py saves results from tests
- add 'configfile' parameter to test function
- '.ini' extension gets removed from 'configfile' and what is left is
used as "stem" for the test results output filename
- also add list of set sizes from data file to dictionary that gets saved
in test results file that is output

@@ -17,7 +17,8 @@ def test(gz_filename,
          epochs_list,
          batch_size,
          model_save_path,
-         test_results_save_path):
+         test_results_save_path,
+         configfile):
     """measure accuracy of trained convolutional neural networks on test set of visual search stimuli
 
     Parameters
@@ -47,6 +48,9 @@ def test(gz_filename,
         path to directory where model checkpoints should be saved
     test_results_save_path : str
         path to directory where results from measuring accuracy on test set should be saved
+    configfile : str
+        filename of config.ini file. Used (without .ini extension) as name for output file
+        that is saved in test_results_save_path.
 
     Returns
     -------
@@ -166,9 +170,11 @@ def test(gz_filename,
 
         if not os.path.isdir(test_results_save_path):
             os.makedirs(test_results_save_path)
+        results_fname_stem = configfile.replace('.ini', '')  # remove .ini extension
         results_fname = os.path.join(test_results_save_path,
-                                     f'test_{net_name}_trained_{epochs}_epochs.gz')
+                                     f'{results_fname_stem}_trained_{epochs}_epochs_test_results.gz')
         results_dict = dict(acc_per_set_size_per_model=acc_per_set_size_per_model,
                             acc_per_set_size_model_dict=acc_per_set_size_model_dict,
-                            predictions_per_model_dict=predictions_per_model_dict)
+                            predictions_per_model_dict=predictions_per_model_dict,
+                            set_sizes=set_sizes)
         joblib.dump(results_dict, results_fname)

### 2020-02-16 docs/notebooks/ccneuro2019/results.Rmd
create ccneuro2019/ in ./docs/notebooks to clean up

---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.0'
      jupytext_version: 1.0.1
  kernelspec:
    display_name: Python 3
    language: python
    name: python3
---

# Lack of evidence for attention-like limits on accuracy of convolutional neural networks performing a visual search task

## Abstract
What limits our ability to find what we are looking for in the cluttered noisy world? To investigate this, cognitive scientists have long used visual search. In spite of hundreds of studies, it remains unclear how to relate effects found using the discrete item display search task to computations in the visual system. A separate thread of research has studied the visual system of humans and other primates using convolutional neural networks (CNNs) as models. Multiple lines of evidence suggest that training CNNs to perform tasks such as image classification causes them to learn  representations similar to those used by the visual system. These studies raise the question of whether CNNs that have learned such representations behave similarly to humans performing other vision-based tasks. Here we address this by measuring the behavior of CNNs trained for image classification while they perform the discrete item display search task. We first show how a fine-tuning approach often used to adapt pre-trained CNNs to new tasks can produce models that show human-like limitations on this task. However we then demonstrate that we can greatly reduce these effects by changing training,without changing the learned representations. Lastly we show that accuracy is not impaired when single networks are trained to discriminate multiple types of visual search stimuli. Based on these findings, we suggest that CNNs are not necessarily subject to the same limitations as the primate visual system.

## Introduction

### The discrete item display search task
On each trial, subjects view a display (example images, $\mathbf{a}$) of discrete items on a flat background. Subjects respond whether the target is present or absent (condition varies across rows of $\mathbf{a}$). Another condition that varies across trials is the set size, that is, the total number of targets and distractors (varies across rows in $\mathbf{a}$). Many studies focus on *set size effects*. Typical effects are an increase in reaction time or a decrease in accuracy as set size increases (depicted schematically in $\mathbf{b}$, redrawn from (Wolfe et al., 2010) and (Eckstein, 1998). Effects vary based on the features that distinguish targets from distractors (shown in columns). Details in introduction. Accuracy for spatial configuration-type stimuli not shown in $\mathbf{b}$ because this has been less studied (but see E. M. Palmer et al., 2011)

![figure 1](../static/fig1/fig1.png)

## Results
### Human-like limits on accuracy of convolutional neural networks (CNNs) performing the discrete item display search task

When using weights pre-trained on ImageNet in convolutional layers, and fine-tuning weights in fully-connected layers so AlexNet could perform the task, this CNN showed decreases in accuracy as set size increased ($\mathbf{a}$). This decrease was smallest for feature search stimuli (left plot), intermediate for conjunction stimuli (middle plot), and largest for spatial configuration stimuli (right plot). The same approach produced similar results with the VGG16 architecture ($\mathbf{b}$)

![Figure 2](../static/fig2/fig2.png)

### Changing training greatly reduces set size effects
Training histories showed that the accuracy of models trained with the fine tuning approach did not converge
on some asymptotic value, and varied depending on the set size of the search stimuli ($\textbf{a}$, left plot). Increasing the learning rate and including more examples of stimuli with larger set sizes greatly sped up convergence ($\textbf{a}$, right plot). AlexNet models trained with this higher learning rate and larger training set showed reduced set size effects ($\textbf{b}$).
![figure 3](../static/fig3/fig3.png)

### Training single networks on multiple stimuli does not impair accuracy
Two instances of AlexNet trained on datasets containing all three types of visual search stimuli used in this study still attained high accuracy.
![figure 4](../static/fig4/fig4.png)

## Methods
Fine-tuning Alexnet and VGG16:
* based on Põder 2017: https://arxiv.org/pdf/1707.09775.pdf
* Stochastic gradient descent
* Base learning rate: 1e-20
* Rate for fully-connected layers: 0.0001
* epochs: 200

Training  
* Freeze convolutional layers
* Rate for fully-connected layers: 0.001
* Epochs: until validation accuracy did not improve for more than 20 epochs

## References

## Acknowledgments
- Research funded by the Lifelong Learning Machines program, DARPA/Microsystems Technology Office,  DARPA cooperative agreement HR0011-18-2-0019
- David Nicholson was partially supported by the 2017 William K. and Katherine W. Estes Fund to F. Pestilli, R. Goldstone and L. Smith, Indiana University Bloomington.
